%%
%% $Id: ag_encounters.tex,v 1.1 2004/10/07 18:26:39 goltz20707 Exp $
%%
\externaldocument[GM-]{gm_guide}

\chapter{Encounters}
\label{cha:encounters}
% X

The following creatures, non-player characters, spaceships, and
accidents are used by the GM when the appropriate encounter occurs
(see GM guide \ref{GM-sec:creating-encounters}). These lists should
not serve as the sole source of encounters; after the GM has run a
number of adventures, the supply will dwindle and the GM should start
to develop his own encounters (although he may continue to get many
ideas from these encounters).  Federal encounters are not described in
this chapter; they occur in accordance with GM guide
\ref{GM-sec:federal-encounters}.

\section{Creatures}
\label{sec:creatures}
% [47.0]

The GM should choose one of the following 40 creatures when a creature
encounter occurs. He consults the appropriate column (common, rare, or
unique) of the \emph{Creature Location} section of the Terrain Effects
Chart (GM guide table \ref{GM-tab:terrain-effects}) to quickly find
which creatures (by identity number) are eligible to appear in the
party's environ, and makes the selection from that group. An
item-by-item explanation of the creature descriptions follows and
detailed \textbf{Explanations of Creature Powers} are included after
these.

If a creature description indicates that more than one creature is
present, the GM should roll one die, two dice or percentile dice (as
seems appropriate to the situation) to determine how many creatures
there are.

Each creature description provides the GM with a complete and variable
explanation of the creature's attributes and is organized as follows:

\begin{description}
\item[Identity Nr.] Identifies the creature when referring to the
  Creature Location Chart.
\item[Restrictions.] Aside from a creature's environ location,
  additional restrictions may be placed on the ranges of temperature,
  gravity, and water availability in which the creature may exist.
  Specific sites for the creature may also be suggested (see the list
  of sites in GM guide \ref{GM-sec:environ-hex-map}).
\item[Combat Rating.] A quantification, ranging from \textbf{1} to
  \textbf{15}, of the creature's brute strength and endurance. Certain
  creatures are incapable of combat (represented by \textbf{na}). The
  creature's strength when attacking (or defending in close combat) is
  equal to its Combat Rating times \emph{one half} of its Agility
  Rating (rounded down). When a creature's Combat Rating is reduced to
  \textbf{0}, it is dead.
\item[Agility Rating.] A quantification, ranging from \textbf{1} to
  \textbf{15}, of the creature's speed and nimbleness. Certain
  creatures are immobile (represented by \textbf{na}). A creature may
  move a number of hexes equal to its current Agility Rating each
  Action Round.
\item[Aggression Rating.] A quantification, ranging from \textbf{1} to
  \textbf{12}, of the creature's self-preserving and aggressive
  instincts. A creature's Aggression Rating is used to determine the
  creature's reaction to the party (see
  GM guide \ref{GM-sec:creature-reaction}).
  
  \textbf{NOTE:} The preceding four ratings are listed as a range of
  three or four possible ratings for each creature, to allow variety
  when using the same creature type more than once. The GM should
  choose specific ratings from these ranges for each creature he uses.
  
\item[Initiative Percentage.] A number, ranging from \textbf{1} to
  \textbf{20}, representing the creature's awareness and ability to
  get the jump on its adversaries. The percentage is multiplied by the
  \emph{Terrain Value} to determine the base awareness chance at the
  beginning of an encounter.
\item[Composition.] The basic stuff of life that the creature is
  composed of. Types include humanoid, mammalian, Terran-like,
  protein, carbon, and non-carbon. Each creature has a primary
  composition and a secondary composition (followed by a percentage
  chance).  The GM may roll percentile dice to see if the creature
  consists of the secondary composition; otherwise, the primary
  composition is used. Composition determines how a creature will be
  affected by drugs and toxins (see
  GM guide \ref{GM-sec:toxin-effects-matrix}).
\item[Size.] A category matching the target size categories in
  GM guide \ref{GM-sec:sizes-of-individuals}. These include very small, small,
  man-sized, large, one-hex, two-hex, and so on. An additional
  category, \emph{minuscule}, is used for bugs, micro organisms, and
  other creatures too small to fire a weapon at or see.
\item[Powers.] A list of one to five powers the creature may possibly
  possess, with a percentage chance for each. Upon choosing a
  creature, the GM rolls percentile dice for each listed power to see
  if the creature has it. The powers and their variable availability
  give each creature a unique and ever-changing \emph{modus operandi}
  (see \textbf{Explanations of Creature Powers}). The GM may ``mix and
  match'' powers as much as he desires.
\item[Warning.] A visual or sensual description to be read to the
  characters if they and the creature are aware of each other at the
  moment of encounter. A warning allows the characters to choose a
  strategy while knowing very little about the appearance and nature
  of the creature.
\item[Sight.] A visual description to be read to the characters if
  they are aware of a creature that is itself not aware at the moment
  of encounter, or when they get their first good look at the
  creature later in the encounter. Sight describes a creature in
  objective visual terms, but reveals nothing that would not be
  immediately obvious. A character may not fire at a creature until he
  has heard this description.
\item[Perception.] An analytical description read to the party if a
  character successfully perceives. Once sight has been established,
  the party may attempt the perception tasks listed in the life sense
  skill (GM guide \ref{GM-sec:psionicskills}) and/or the perception tasks
  listed in the biology skill (GM guide \ref{GM-sec:scientific-skills}). If
  the creature has a non-organic composition, the GM may allow a
  perception task using a scientific skill other than biology.
  Perception may reveal vulnerable areas of the creature, its
  intentions towards the party, its method of propulsion, attack,
  defense, and ingestion, and other information that will aid the
  party in dealing with the creature.
\item[Examination.] A physiological description read to the party if a
  character successfully conducts examination (see the biology skill,
  GM guide \ref{GM-sec:scientific-skills}). Examination may only be attempted
  if the creature is dead, unconscious or safely restrained.
  Examination may reveal strange fluids in the creature's interior
  systems, entirely unique interior systems, valuable chemicals, drugs
  or poisons in the creature, and whether or not it is edible.
\end{description}

\textbf{NOTE:} Warning, sight, perception, and examination may also
reveal any \emph{powers} possessed by the creature. Each power in
Explanation of Creature Powers lists one of these four categories and
should be revealed to the party when they are read that category
description (if they have not been made aware of the power by the
creature's actions). The GM may always alter or elaborate upon the
wording of a description to better fit the situation. Many creatures
do not possess all four description categories; such a creature
possesses no attributes of interest that would be revealed in the
category.

Certain creatures with bizarre attributes have special procedures for
their use listed at the end of their description.

None of the creatures is given an actual name. If the characters are
encountering a creature that they already know exists in the area, the
GM should provide a name. If they are encountering a previously
undiscovered creature, the GM should allow the party to agree upon a
name.

\newcounter{creature}
\newenvironment{creature}[1]{%
\refstepcounter{creature}\paragraph{CREATURE Nr. \thecreature:}

#1\\
}{%

\bigskip

}

\pagebreak

%%\begin{multicols}{2}
  \begin{creature}{Man size}
    \label{sec:creatures-1}
    \textbf{Restrictions:} Light vegetation; woods; forest.  Non-corrosive
    atmosphere.  Gravity of \textbf{0.6} to \textbf{1.6}. Temperature
    between \textbf{30\textdegree} and \textbf{100\textdegree}. \\ 
    \textbf{Combat:} 2--3\hspace{2ex}\textbf{AY:} 3--6 \textbf{AG:} 1--3 \textbf{IN:}
    na \textbf{Initiative:} 2\%  \\
    \textbf{Composition:} Mammalian; Terran 10\% \\
    \textbf{Powers:} Edible 5\%; camouflage 5\%; heightened smell 30\% \\
    \begin{itemize}
    \item \textbf{Warning:} You see a group of large four-legged entities
      which seem to be busy grazing on the local flora.  \\
    \item \textbf{Sight:} Each entity is two meters long and stands almost
      two meters high at the shoulder. They look like Terran swine, but
      have elongated pointed snouts (about 10 to 16 centimeters long).
      They appear stupid and harmless.
    \end{itemize}
  \end{creature}

\hrulefill

  \begin{creature}{Small}
    \textbf{Restrictions:} Light vegetation; woods; forest; marsh. Must
    have non-corrosive atmosphere. Gravity between \textbf{0.4} and \textbf{1.4}.
    Temperature
    between \textbf{20\textdegree} and \textbf{100\textdegree}. Near water. \\
    \textbf{Combat:} 1--3 \textbf{AY:} 1--2 \textbf{AG:} 2--3 \textbf{IN:} na \textbf{Initiative:} 2\% \\
    \textbf{Composition:} Mammalian; Terran 40\% \\
    \textbf{Powers:} Dart shooting 100\%; edible 10\%; heightened hearing
    30\%; poison attack 35\%; camouflage 5\%
    \begin{itemize}
    \item \textbf{Warning:} You hear a crashing and rooting in the nearby
      undergrowth.
    \item\textbf{Sight:} A meter-long hedgehog-like entity covered in
      15--20cm long quills. Locomotion is by slithering on belly since
      creature is legless.
    \item\textbf{Perception:} Entity has anteater-type mouth and single
      eye. Shoots quills at threatening objects or entities if frightened.
    \item\textbf{Examination:} Belly is not coated with quills. Entity is
      helpless if turned on back since powerful locomotive muscles are on
      its underside only.
    \end{itemize}
  \end{creature}

\hrulefill

  \begin{creature}{Minuscule}
    \textbf{Restrictions:} Marsh; jungle. Must have non-corrosive
    atmosphere. Gravity between \textbf{0.4} and \textbf{2.0}. Temperature above
    \textbf{50\textdegree}.\\
    \textbf{Combat:na}
    \textbf{AY:} 8--10 \textbf{AG:} 10--12 \textbf{IN:} na \textbf{Initiative:} 15\% \\
    \textbf{Composition:} Terran; protein 20\% \textbf{Powers:} Camouflage 100\%; flight 100\%; poison attack 100\%; radar sense 60\%; swarm mind 30\% 
    \begin{itemize}
    \item\textbf{Warning:} You notice that members of your party not wearing body armor are acting strangely, as if they are suffering from some 
      illness, though the exact nature of the problem is not clear. 
    \item\textbf{Perception:} The illness will be diagnosed as a type of poison (virulent, viral, paralytic, or psychotic, but not a spore) and will be 
      connected with a one-inch long insect present in great numbers. 
    \item\textbf{Examination:} Insect has powerful quarter-inch sting in tail containing poison. 
    \end{itemize}
  \end{creature}

\hrulefill

  \begin{creature}{2-hex}
    \textbf{Restrictions:} Light vegetation; woods; marsh; forest; jungle. Normal atmosphere only. Gravity of \textbf{0.6} to \textbf{2.5}. Temperature between \textbf{50\textdegree} 
    and \textbf{150\textdegree}. \\\\
    \textbf{Combat:} 8--10 \textbf{AY:} 6--8 \textbf{AG:} 9--12 \textbf{IN:} na \textbf{Initiative:} 15\% \\
    \textbf{Composition:} Terran; protein 10\% \\
    \textbf{Powers:} Chitinous carapace 90\%; tunnelling 65\%; entrapment 35\%; poison attack 80\%; regeneration 20\% 
    \begin{itemize}
    \item\textbf{Warning:} You spot segmented centipede-like entity 8 meters or more in length and 2 meters thick. 
    \item\textbf{Sight:} The entity has 8 to 10 body segments, a score of triple-jointed legs (with broad softly padded feet) and a set of mandibles. 
    \item\textbf{Examination:} The padded feet can be used as scoops to assist in rapid tunnelling. Mandibles are the prime weapons. 
    \end{itemize}
  \end{creature}

\hrulefill

  \begin{creature}{Small}
    \textbf{Restrictions:} Light vegetation; woods; marsh; forest; jungle. Non-poisonous, non-corrosive atmosphere. Gravity of \textbf{0.6} to \textbf{1.6}. 
    Temperature between \textbf{100\textdegree} and \textbf{175\textdegree}. \\\\
    \textbf{Combat:} 4--6 \textbf{AY:} na \textbf{AG:} 3--5 \textbf{IN:} na \textbf{Initiative:} 3\% \\
    \textbf{Composition:} Protein; Terran 5\% \\
    \textbf{Powers:} Adhesive secretion 100\%; poison 100\%; radar sense 20\%; regeneration 5\%; edible 60\% 
    \begin{itemize}
    \item\textbf{Warning:} You observe an odd, cactus-like form of vegetation. 
    \item\textbf{Sight:} Vegetation is less than a meter tall and 20 centimeters in diameter and appears to be entirely smooth-skinned. 
    \item\textbf{Examination:} Plant has sweet center favoured by local fauna who are induced to steer clear of it only by its defenses, a sticky 
      substance that makes mouths adhere to ii and a poison which it then secretes. 
    \end{itemize}
  \end{creature}

\hrulefill

  \begin{creature}{Man-sized}
    \textbf{Restrictions:} Light vegetation; woods. Any atmospheres other than thin. Gravity of \textbf{0.4} to \textbf{1.2}. Temperature between \textbf{75\textdegree} and \textbf{150\textdegree}. 
    Water must be readily available. \\\\
    \textbf{Combat:} 4--7 \textbf{AY:} 8--10 \textbf{AG:} 3--5 \textbf{IN:} na \textbf{Initiative:} 2\% \\
    \textbf{Composition:} Mammalian; Terran 20\% \\
    \textbf{Powers:} Swarm mind 100\%; edible 10\%; poison attack 80\% 
    \begin{itemize}
    \item\textbf{Warning:} You observe a small herd (6--18) of what appear to be four-legged lizards. They are grazing on local flora. 
    \item\textbf{Sight:} Two additional legs (total of 6) partially retractable and tipped with sharp claws as main defense. Stands one meter tall and 
      about 2 meters in length. Forked tail. 
    \item\textbf{Perception:} Entity is very fast for herbivore. 
    \item\textbf{Examination:} Entity is omnivore (not herbivore) with retracting fangs for tearing meat. Herd group moves and acts as one. Usually 
      eat only flora and carrion, but will attack for meat if starving. 
    \end{itemize}
  \end{creature}

\hrulefill

  \begin{creature}{Very small}
    \textbf{Restrictions:} Light vegetation; woods; forest; marsh; jungle. Normal or poisonous atmosphere. Gravity of \textbf{0.6} to \textbf{1.2}. Temperature 
    between \textbf{50\textdegree} and \textbf{200\textdegree}. \\\\
    \textbf{Combat:} 1--2 \textbf{AY:} 10--12 \textbf{AG:} 4--6 \textbf{IN:} na \textbf{Initiative:} 5\% \\
    \textbf{Composition:} Terran; protein 30\% \\
    \textbf{Powers:} Edible 5\%; swarm mind 60\%; poison attack 5\%; flying 100\% 
    \begin{itemize}
    \item\textbf{Warning:} You hear a low droning buzz. 
    \item\textbf{Sight:} A score of 20-centimeter long insectile creatures much like dragonflies. 
    \item\textbf{Perception:} Entities are slightly territorial, but herbivorous. 
    \end{itemize}
  \end{creature}

\hrulefill

  \begin{creature}{Large}
    \textbf{Restrictions:} Woods; forest. Non-poisonous atmosphere. Gravity of \textbf{0.6} to \textbf{1.6}. Temperature of \textbf{50\textdegree--100\textdegree}. Must be near water 
    source. \\\\
    \textbf{Combat:} 7--12 \textbf{AY:} 8--10 \textbf{AG:} 10--12 \textbf{IN:} na \textbf{Initiative:} 15\% \\
    \textbf{Composition:} Mammalian; Terran 30\% \\
    \textbf{Powers:} Heightened vision 90\%; heightened smell 100\%; camouflage 80\% 
    \begin{itemize}
    \item\textbf{Warning:} You have a feeling that you are being watched. 
    \item\textbf{Sight:} Hairless, otherwise wolf-like 3-meter long entity with prehensile tail. Has long ``whiskers'' or feelers around eyes. 
    \item\textbf{Perception:} Mouth contains large sharp fangs. Entity is carnivore. Will not flee or otherwise leave. 
    \item\textbf{Examination:} ``Whiskers'' are sense organs used to smell
      prey. Entity has no ``nose'' in Terran sense\ldots just bare organs. 
    \end{itemize}
  \end{creature}

\hrulefill

  \begin{creature}{Man sized}
    \textbf{Restrictions:} Volcanic; crater; barren; light vegetation. Thin atmosphere only. Gravity of \textbf{0.2} to \textbf{0.8}. Temperature less than \textbf{200\textdegree}. \\\\
    \textbf{Combat:} 6--8 \textbf{AY:} 6--8 \textbf{AG:} 7--10 \textbf{IN:} na \textbf{Initiative:} 20\% 
    Composition; Mammalian; Terran 15\% \\
    \textbf{Powers:} Radar sense 100\%; poison attack 100\%; poison 10\% 
    \begin{itemize}
    \item\textbf{Warning:} You hear a horrible sucking sound originating nearby. 
    \item\textbf{Sight:} Ashort six-legged, hairless dog-like entity. Black skin. Headless, with neck ending in muscular mouth-like opening. Entity 
      stands 1 Vi meters at the shoulder and is 2 meters long. 
    \item\textbf{Examination:} Entity is toothless. Instead, strong muscles crush foreign matter in throat while poisonous secretion breaks down fibres. 
      Carnivorous. Retractable claws for pinning prey on front and rear legs. 
    \end{itemize}
  \end{creature}

\hrulefill

  \begin{creature}{Small}
    \textbf{Restrictions:} Water. Lake or pond only. Must have atmosphere. Gravity between \textbf{0.4} and \textbf{1.6}. Temperature between \textbf{75\textdegree} and \textbf{150\textdegree}. \\\\
    \textbf{Combat:} 5--8 \textbf{AY:} 8--10 \textbf{AG:} 3--5 \textbf{IN:} na \textbf{Initiative:} 1\% \\
    \textbf{Composition:} Terran; protein 20\% \\
    \textbf{Powers:} Heightened hearing 100\%; poison 90\%; reflective carapace 10\% 
    \begin{itemize}
    \item\textbf{Warning:} You observe a small blur of red out of the corner of your eye. 
    \item\textbf{Sight:} Small amphibious saurian about a meter long with 8 legs, tail, and feathery ridges along back. 
      Covered in iridescent red scales. 
    \item\textbf{Perception:} Entity is omnivorous, timid, and reclusive. 
    \item\textbf{Examination:} Feathery ridges are sense organs. Feet are webbed with clear membranes. 
    \end{itemize}
  \end{creature}

\hrulefill

\begin{creature}{Large}
\textbf{Restrictions:} Crater; barren; light vegetation. Non-corrosive, non-poisonous atmosphere. Gravity of 0.4 to 1.6. Temperature between 
75\textdegree\ and 200\textdegree. Usually found near water. \\\\
\textbf{Combat:} 2--4 \textbf{AY:} 1--3 \textbf{AG:} 8--10 \textbf{IN:} na \textbf{Initiative:} 5\% \\
\textbf{Composition:} Terran; protein 20\% \\
\textbf{Powers:} Camouflage 50\%; edible 30\%; heightened vision 40\%;
poison discharge 80\% 
\begin{itemize}
\item \textbf{Warning:} You observe a large, bird-like entity approximately 4 meters high. 
\item \textbf{Sight:} Entity appears to be watching party. Has short wings, feather-like skin, four legs, and large comb. 
\item \textbf{Perception:} Size of wings indicates that entity is flightless. Teeth in beak-like mouth indicate carnivore or omnivore. 
\item \textbf{Examination:} Reveals poison discharge as gaseous cloud from anal orifice (where appropriate). 
\end{itemize}
\end{creature}

\hrulefill

\begin{creature}{19 hex}
\textbf{Restrictions:} Water. Temperature between 50\textdegree\ and 150\textdegree. Large body of water only. May be submerged or on surface. \\\\
\textbf{Combat:} 12--15 \textbf{AY:} 12--15 \textbf{AG:} 10--12 \textbf{IN:} na \textbf{Initiative:} 20\% \\
\textbf{Composition:} Protein; Terran 20\% \\
\textbf{Powers:} Adhesive secretion 30\%; radar sense 50\%; regeneration 40\% (not center hex); inky discharge 30\% 
\begin{itemize}
\item \textbf{Warning:} A large undulating mass under your party is rising rapidly toward you. 
\item \textbf{Sight:} A collection of squid-like tentacles joined at the center and about 25 meters in diameter. 
\item \textbf{Perception:} Has 30 12-meter long suckered tentacles. No body at center point, but lidless eyes and sharp beaks are at the tip of 
each tentacle. Presents man-sized target in each hex occupied (due to open structure). Obviously very dangerous. 
\item \textbf{Examination:} Small, powerful heart-like organ at center. Nerve collections at end of each tentacle form brain of sorts. 
\end{itemize}
\end{creature}

\hrulefill

\begin{creature}{Small}
\textbf{Restrictions:} Air. Requires non-corrosive atmosphere and moisture. Gravity of 0.2 to 1.0. Temperature between 50\textdegree\ and 200\textdegree. \\\\
\textbf{Combat:} 1--2 \textbf{AY:} 1--2 \textbf{AG:} 1--2 \textbf{IN:} na \textbf{Initiative:} 10\% \\
\textbf{Composition:} Protein; carbon 40\% \\
\textbf{Powers:} Poison discharge 100\%; multiple attack 100\%; kinetic absorption 10\%; unknown propulsion 5\%; poison 15\% 
\begin{itemize}
\item \textbf{Warning:} What appear to be windblown puff-balls are headed in your direction. 
\item \textbf{Sight:} Each ball is 50 centimeters in diameter. Balls have fuzzy appearance. 
\item \textbf{Perception:} Balls are loose aggregations of seedpods which are discharged when in presence of energy source (including body 
heat). Entity itself is a floating plant. 
\end{itemize}
\end{creature}

\hrulefill

\begin{creature}{Large}
\textbf{Restrictions:} Barren; crater; light vegetation. Must have atmosphere. Gravity of 0.4 to 2.0. Any temperature. Sunlight. \\\\
\textbf{Combat:} na \textbf{AY:} na \textbf{AG:} na \textbf{IN:} na \textbf{Initiative:} na \\
\textbf{Composition:} Protein; carbon 20\% \\
\textbf{Powers:} Edible 10\% 
\begin{itemize}
\item \textbf{Sight:} An artichoke-like plant 3 meters tall and 4 meters in diameter. Leaves will be open and flat on ground during daylight, but close 
up to form perfectly sealed chamber at night. 
\end{itemize}
\end{creature}

\hrulefill

\begin{creature}{Large}
\textbf{Restrictions:} Air. Normal atmosphere only. Gravity of 0.2 to 1.0. Temperature above 100\textdegree. \\\\
\textbf{Combat:} 4--7 \textbf{AY:} 10--11 \textbf{AG:} 5--7 \textbf{IN:} na \textbf{Initiative:} 15\% \\
\textbf{Composition:} Protein; Terran 45\% \\
\textbf{Powers:} Unknown propulsion 100\%; multiple attack 100\%; radar sense 60\%; balance disruption 10\% 
\begin{itemize}
\item \textbf{Warning:} You hear a rushing noise and a flapping of wings. 
\item \textbf{Sight:} A 15-meter long "flying" reptile-like entity. Wings are used mostly to guide entity, seldom flapped. Span of 5 meters. 
\item \textbf{Perception:} Two orifices on leading edge of wings near juncture with torso. Two retracting talons on underside. 
\item \textbf{Examination:} Orifices are actually a biological "supercharger" used to maintain speed and altitude. Entity is a carnivore with large 
needs since energy for the "supercharger" is provided by it's body. 
\end{itemize}
\end{creature}

\hrulefill

\begin{creature}{Large}
\textbf{Restrictions:} Woods flat; woods hills; light vegetation flat; light vegetation hills; ice flat; ice hills. Any non-poisonous atmosphere. 
Gravity of 0.6 to 1.6. Temperature between 0\textdegree\ and 90\textdegree. \\\\
\textbf{Combat:} 3--6 \textbf{AY:} 3--5 \textbf{AG:} 3--5 \textbf{IN:} na \textbf{Initiative:} 5\% \\
\textbf{Composition:} Protein; Terran 40\% \\
\textbf{Powers:} Edible 80\%; heightened smell 90\%; invisibility 100\% 
\begin{itemize}
\item \textbf{Warning:} You observe a group of herd animals (between 10 and 200) grazing on the local vegetation. Each entity weighs about one 
ton apparently and measures close to 3 meters in length. They appear to be antelope-like, but are six-legged and not horned. 
\item \textbf{Sight:} Entities will always become invisible when frightened. 
\end{itemize}
\end{creature}

\hrulefill

\begin{creature}{Man-sized}
\textbf{Restrictions:} Any environ. Gravity between 0.2 and 1.2. Must have non-poisonous, non-corrosive atmosphere. 
Temperature between 0\textdegree\ and 75\textdegree. \\\\
\textbf{Combat:} 7--9 \textbf{AY:} 6--9 \textbf{AG:} 9--12 \textbf{IN:} na \textbf{Initiative:} 20\% \\
\textbf{Composition:} Protein; Terran 40\% \\
\textbf{Powers:} Entrapment 100\%; heightened vision 60\%; poison attack 70\%; reflective carapace 10\%; adhesive secretion 80\% 
\begin{itemize}
\item \textbf{Warning:} You notice a camouflaged pit 5 meters deep and covered with a sticky web-like film. If unaware, one random member of 
the party falls into the pit. 
\item \textbf{Sight:} The inside of the pit is covered with a sticky substance. Bottom contains three tunnels (each one meter in diameter). Sound of 
skittering in darkness. 
\item \textbf{Perception:} Occupant is a 2-meter long insectile-appearing entity much like a wingless wasp or hornet. 
\item \textbf{Examination:} Creature inhabits tunnels created by some departed life-form. Uses as both home and trapping device. Will leave 
tunnels if it thinks it can make a kill. 
\end{itemize}
\end{creature}

\hrulefill

\begin{creature}{Man sized}
\textbf{Restrictions:} Any environ. Found only in poisonous or corrosive atmosphere. Gravity of 0.2 to 0.8. Temperature between 50\textdegree\ and 
150\textdegree. \\\\
\textbf{Combat:} 6--8 \textbf{AY:} 2--4 \textbf{AG:} 5--7 \textbf{IN:} na \textbf{Initiative:} 2\% \\
\textbf{Composition:} Carbon; non-carbon 40\% \\
\textbf{Powers:} Radar sense 20\%; acid 100\%; adhesive secretion 40\%; unknown propulsion 20\%. 
\begin{itemize}
\item \textbf{Warning:} You see a jellyfish-type creature 2 meters in diameter. It is transparent and the internal organs are clearly visible. 
\item \textbf{Sight:} Entity's spherical shape is broken by two pseudopods, each about 2 meters long. Each pseudopod has a ring of black dots on 
it. Entity glistens as if wet. 
\item \textbf{Perception:} Black dots appear to be sense organs of some sort. They are coated with some sort of secretion. 
\end{itemize}
\end{creature}

\hrulefill

\begin{creature}{Very small}
\textbf{Restrictions:} Woods; forest; marsh. Non-poisonous, non-corrosive atmosphere. Gravity of 0.4 to 1.2. Temperature between 50\textdegree\ and 
100\textdegree. Only found near water (large pond or small lake). \\\\
\textbf{Combat:} 1--2 \textbf{AY:} 7--9 \textbf{AG:} 4--5 \textbf{IN:} 5--7 \textbf{Initiative:} 15\% \\
\textbf{Composition:} Protein; Terran 50\% \\
\textbf{Powers:} Camouflage 100\%; radar sense 10\%; edible 20\%; mind control 90\%; psionic communication 5\% 
\begin{itemize}
\item \textbf{Warning:} A character with a Mental Power of 4 or greater feels that he is being observed. 
\item \textbf{Sight:} The source of the emanations appears to be a small (1-foot long) toad-like entity which is barely discernible. Entity has three 
legs, skin which changes colour to blend with surroundings. Appears to have tufted antennae along backbone. 
\end{itemize}
\end{creature}

\hrulefill

\begin{creature}{Small}
\textbf{Restrictions:} Woods; forest; marsh. Non-corrosive
 atmosphere. Gravity of 0.4 to 1.0. Temperature between 75\textdegree\
 and 125\textdegree.\\\\
\textbf{Combat:} 7--11 \textbf{AY:} 9--12 \textbf{AG:} 10--12 \textbf{IN:} 1--3 \textbf{Initiative:} 20\% \\
\textbf{Composition:} Mammalian; Terran 30\% \\
\textbf{Powers:} Entrapment 85\%; camouflage 100\%; heightened smell 90\% 
\begin{itemize}
\item \textbf{Warning:} You hear a rustling noise from the trees above. 
\item \textbf{Sight:} A 1.5 meter tall monkey-like entity with two long tentacles (2 meters each) instead of arms, a circular, sharp-toothed mouth, 
single central eye. Hairless. Skin appears to change colour. 
\end{itemize}
\end{creature}

\hrulefill

\begin{creature}{4 hex}
\textbf{Restrictions:} Volcanic hills; volcanic mountain; crater hills; crater mountain; crater flat; barren flat; barren hills; barren mountain. Must 
have atmosphere. Gravity 0.2 to 0.8. Temperature between 60\textdegree\ and 200\textdegree. \\\\
\textbf{Combat:} 12--15 \textbf{AY:} 9--12 \textbf{AG:} 9--12 \textbf{IN:} na \textbf{Initiative:} 20\% \\
\textbf{Composition:} Protein; carbon 5\% \\
\textbf{Powers:} Tunnelling 100\%; entrapment 40\%; poison attack 100\%; camouflage 50\%; regeneration 30\% 
\begin{itemize}
\item \textbf{Warning:} The earth begins to rumble and lose solidity. 
\item \textbf{Sight:} A gigantic (20-meter long, 7-meter thick) worm-like entity. Segmented. 5-meter maw at one end contains ridged muscles for 
crushing prey or defense. 
\item \textbf{Perception:} Maw also has poisonous digestive juices secreted from sacs inside muscle ridges. Heavily-lidded visual sensor located 
inside maw (protected by muscle). 
\end{itemize}
\end{creature}

\hrulefill

\begin{creature}{2 hex}
\textbf{Restrictions:} Water. Must have some atmosphere. Gravity of 0.5 to 2.5. Temperature between 50\textdegree\ and 125\textdegree. \\\\
\textbf{Combat:} 6--8 \textbf{AY:} 7--9 \textbf{AG:} 4--6 \textbf{IN:} na \textbf{Initiative:} 15\% \\
\textbf{Composition:} Protein; Terran 20\% \\
\textbf{Powers:} Radar sense 80\%; energy ingestion 100\%; balance disruption 100\%; acidic circulatory fluid 70\%; camouflage 15\% 
\begin{itemize}
\item \textbf{Warning:} You experience a sense of disorientation and your sensors begin to give off ``jangled'' readings indicating a large power 
source nearby. 
\item \textbf{Sight:} A 9-meter long by 2-meter thick eel-like entity without apparent head or tail. Glows slightly. 
\item \textbf{Perception:} Entity ``eats'' and stores energy (like a large battery). Retractable eye stalks at both ends used for guidance. 
\end{itemize}
\end{creature}

\hrulefill

\begin{creature}{Large}
 \textbf{Restrictions:} Barren flat; barren hills; light vegetation flat; light vegetation hills; crater flat. Non-poisonous, 
non-corrosive atmosphere. Gravity of 0.4 to 1.0. Temperature between 20\textdegree\ and 80\textdegree. \\\\
\textbf{Combat:} 4--7 \textbf{AY:} 7--10 \textbf{AG:} 4--6 \textbf{IN:} na \textbf{Initiative:} 6\% \\
\textbf{Composition:} Protein; Terran 5\% \\
\textbf{Powers:} Radar sense 80\%; poison 100\%; camouflage 5\%; psionic communication 60\%; web shooting 5\% 
\begin{itemize}
\item \textbf{Warning:} You observe a large, many-legged entity skittering away from you. 
\item \textbf{Sight:} Entity appears to be a giant spider-like being, with a dozen legs meeting in a smooth, round body. Each leg is about 3 meters 
in length. 
\item \textbf{Perception:} Entity is actually two species. Spider-like entity has low-level nervous system. Higher level system is provided by 
symbiotic moss growing on body. Moss is poisonous. ``Spider'' is herbivore. 
\end{itemize}
\end{creature}

\hrulefill

\begin{creature}{Man sized}
\textbf{Restrictions:} Any environ containing a large body of water nearby (ocean, lake, or large pond). Must have atmosphere. Gravity of 0.4 
to 1.4. Temperature between 50\textdegree\ and 150\textdegree. \\\\
\textbf{Combat:} 4--6 \textbf{AY:} 5--8 \textbf{AG:} 3--5 \textbf{IN:} 1--3 \textbf{Initiative:} 10\% \\
\textbf{Composition:} Protein; Terran 5\% \\
\textbf{Powers:} Capture 80\%; inky discharge 40\%; poison attack 10\%; camouflage 100\% 
\begin{itemize}
\item \textbf{Warning:} A scaled humanoid appears before you. 
\item \textbf{Sight:} Entity has webbed feet, gills, bluish scales, and is 3 meters tall. 
\item \textbf{Perception:} Entity is adapted to both surface atmosphere and water. Hands have retractable claws. Teeth indicate entity is a 
carnivore. 
\end{itemize}
\end{creature}

\hrulefill

\begin{creature}{Large}
\textbf{Restrictions:} Found only in the absence of atmosphere. \\\\
\textbf{Combat:} 10--12 \textbf{AY:} 8--10\footnote{Agility is ``na'' if entity does not have unknown propulsion.} \textbf{AG:} 6--8 \textbf{IN:} 5--7 \textbf{Initiative:} 20\% \\
\textbf{Composition:} Carbon; non-carbon 40\% \\
\textbf{Powers:} Unknown propulsion 75\%; energy drain 100\%; energy blast 40\% 
\begin{itemize}
\item \textbf{Warning:} You see a floating red, pulsing light with a radius of about 3 meters. It gives off a constant discharge of heat (350\textdegree) and 
radiation. 
\item \textbf{Perception:} All equipment being drained of power by entity. 
\item \textbf{Examination:} Entity views energy sources as ``life.'' Draws no distinction between man and machine. Will neither attack nor flee. 
\end{itemize}
\end{creature}

\hrulefill

\begin{creature}{Small}
\textbf{Restrictions:} Air. Any atmosphere. Gravity between 0.2 and 0.8. Temperature less than 200\textdegree. \\\\
\textbf{Combat:} na \textbf{AY:} 4--7 \textbf{AG:} 2--4 \textbf{IN:} 3--5 \textbf{Initiative:} 5\% \\
\textbf{Composition:} Carbon; protein 30\% \\
\textbf{Powers:} Unknown propulsion 100\%; radar sense 30\%; psionic communication 20\%; poison discharge 50\%; heightened hearing 70\% 
\begin{itemize}
\item \textbf{Warning:} You see a balloon-like object in the distance. 
\item \textbf{Sight:} Object is living entity composed of gasbag body and half a dozen 2-foot trailing tendrils grouped on underside. 
\item \textbf{Perception:} Being maintains altitude via gases produced by body and stored inside entity. Same gases can be vented out underside 
as means of protection having same effects as poison. 
\item \textbf{Examination:} Tiny tendrils are sensors and manipulative appendages at the same time. Entity is simple-minded floating plant which 
converts air and sunlight into gases to feed itself. 
\end{itemize}
\end{creature}

\hrulefill

\begin{creature}{Large}
\textbf{Restrictions:} Volcanic; crater; barren. Must have some atmosphere. Gravity of 0.6 to 1.6. Temperature between 50\textdegree\ and 250\textdegree. \\\\
\textbf{Combat:} 8--10 \textbf{AY:} 4--6 \textbf{AG:} 9--12 \textbf{IN:} 2--4 \textbf{Initiative:} 20\% \\
\textbf{Composition:} Non-carbon; carbon 20\% \\
\textbf{Powers:} Camouflage 100\%; entrapment 40\%; armored carapace 65\%; tunnelling 100\%; acid 100\% 
\begin{itemize}
\item \textbf{Warning:} You catch a movement out of the corner of your eye. 
\item \textbf{Sight:} Movement generated by nearby rock. Appears to be somehow animated. Rock is 3 meters in diameter by 1 meter high. 
\item \textbf{Perception:} Rock is living entity capable of disappearing by burrowing into and eating rock. 
\item \textbf{Examination:} Emily secretes acid which ``digests'' rock. Entity is apparent creator of nearby tunnel system. 
\end{itemize}
\end{creature}

\hrulefill

\begin{creature}{Small}
\textbf{Restrictions:} Air. Some atmosphere required. Any gravity. Any temperature. \\\\
\textbf{Combat:} na \textbf{AY:} na \textbf{AG:} na \textbf{IN:} na \textbf{Initiative:} 0\% \\
\textbf{Composition:} Carbon; non-carbon 30\% \\
\textbf{Powers:} Swarm mind 100\%; energy absorption 100\%; energy ingestion 5\%; energy drain 10\%; kinetic absorption 100\% 
\begin{itemize}
\item \textbf{Warning:} The air is increasingly filled with sticky scraps of some diaphanous membrane. 
\item \textbf{Sight:} Membrane is crossed with tiny silver filaments that catch sunlight. 
\item \textbf{Perception:} Scraps are sometimes blown apart or together by the wind. Larger scraps pulse with light. Appear to absorb energy. 
\item \textbf{Examination:} The membranes are cells which when separate are harmless, if annoying. When the number which blow together 
reaches a critical mass (several thousand), they form a single entity which may even develop a rudimentary intelligence. 
\end{itemize}
\end{creature}

\hrulefill

\begin{creature}{Man-sized}
\textbf{Restrictions:} Any land environ. Gravity of 0.2 to 0.6. Non-corrosive atmosphere. \\\\
\textbf{Combat:} 6--8 \textbf{AY:} 6--9 \textbf{AG:} 8--10 \textbf{IN:} 2--4 \textbf{Initiative:} 10\% \\
\textbf{Composition:} Terran; mammalian 20\% \\
\textbf{Powers:} Heightened visual sense 60\%; adhesive secretion 70\%; flying 100\% 
\begin{itemize}
\item \textbf{Warning:} You observe a large bat-like creature about 2 meters tall with a wingspan of 4 meters. 
\item \textbf{Sight:} The entity has a somewhat humanoid torso and face, but no arms and long (2 meter) tendrils instead of legs. The initial bat-like 
appearance is a result of the large leathery wings. 
\item \textbf{Perception:} Entity has compound eyes, is carnivorous, aggressive, and seemingly intelligent. 
\end{itemize}
\end{creature}

\hrulefill

\begin{creature}{Minuscule}
\textbf{Restrictions:} Any environ except water. Must have atmosphere. Only encountered if party are not wearing expedition suits or one or 
more such suits have been punctured. \\\\
\textbf{Combat:} na \textbf{AY:} na \textbf{AG:} 7--8 \textbf{IN:} 8--10\footnote{Entity will possess intelligence only if psionic communication is possessed.} \textbf{Initiative:} 20\% \\
\textbf{Composition:} Protein; Terran 10\% \\
\textbf{Powers:} Unknown propulsion 100\%; radar sense 80\%; psionic communication 40\%; mental attack 10\%; mind control 100\% 
\begin{itemize}
\item \textbf{Warning:} One or more party members begin to behave erratically, picking fights with others in the party, breaking equipment, and so 
forth. 
\item \textbf{Perception:} Disturbed party members infected by foreign organism. 
\item \textbf{Examination:} Attacking organism identified as almost microscopic tick-like entity of malevolent intent. Never attacks. Uses mind 
control instead. 
\end{itemize}
\end{creature}

\hrulefill

\begin{creature}{Man sized}
\textbf{Restrictions:} Any environ. Only thin atmosphere or no atmosphere. Any gravity. Any temperature. \\\\
\textbf{Combat:} 10--13 \textbf{AY:} 9--12 \textbf{AG:} 5--8 \textbf{IN:} na \textbf{Initiative:} 5\% \\
\textbf{Composition:} Carbon; non-carbon 50\% \\
\textbf{Powers:} Unknown propulsion 100\%; energy ingestion 100\%; energy blast 20\%; energy drain 60\%; radar sense 100\% 
\begin{itemize}
\item \textbf{Warning:} A sphere of plasma appears. It glows intensely from inner light. 
\item \textbf{Sight:} The sphere is perfectly rounded and undulates slightly as it moves. It has a diameter of 2 meters. Floats several centimetres 
from earth. Rotates slowly clockwise. 
\item\textbf{Perception:} Area within 5 meters of entity is very hot (in terms of both thermal and radiation effects). 
\item \textbf{Examination:} Radiation lingers over area after passage. 
\end{itemize}
\end{creature}

\hrulefill

\begin{creature}{Small (pod)}
\textbf{Restrictions:} Light vegetation; woods. Non-poisonous, non-corrosive atmosphere. Gravity of 0.6 to 1.8. Temperature less than 200\textdegree. \\\\
\textbf{Combat:} 6--8 \textbf{AY:} 5--7\footnote{Agility refers to ``seeds'' only.} \textbf{AG:} 9--12 \textbf{IN:} 5--7 \textbf{Initiative:} 10\% \\
\textbf{Composition:} Protein; carbon 40\% \\
\textbf{Powers:} Unknown propulsion 10\%; swarm mind 80\%; mind control 100\%; psionic communication 90\%; multiple attack 100\% 
\begin{itemize}
\item \textbf{Warning:} Character with Mental Power of 4 or greater will experience a premonition of evil so strong as to be a physical force. 
\item \textbf{Sight:} A patch of tangled plants containing watermelon-like pods about a meter long. Plants are knee high. 
\item \textbf{Perception:} Psions will perceive this as the source of the ``evil'' emanations. 
\item \textbf{Examination:} Pods weigh 35 pounds. Clear skin reveals seed-like forms inside. Pods attack by bursting and squirting seeds 3 meters 
in all directions. ``Seeds'' fold out upon leaving pod and take on a manta ray form. Attach to spinal column from whence they can control 
the entity to which they are affixed. 
\end{itemize}
\end{creature}

\hrulefill

\begin{creature}{Large}
\textbf{Restrictions:} Woods; forest; jungle. Must have
 atmosphere. Gravity of 0.6 to 1.2. Temperature between 50\textdegree\
 and 100\textdegree.\\\\
\textbf{Combat:} 9--13 
\textbf{AY:} 9--13\footnote{Refers to neck orifice only; the plant itself is not mobile.} \textbf{AG:} 11--12 \textbf{IN:} 6--8 \textbf{Initiative:} 20\% \\
\textbf{Composition:} Protein; Terran 20\% \\
\textbf{Powers:} Heightened smell 100\%; adhesive secretion 80\%; edible 10\%; psionic communication 20\%; vegetation control 5\% 
\begin{itemize}
\item \textbf{Warning:} You see what appears to be a giant mushroom about 3 meters tall. A character with a Mental Power of 4 or greater will 
experience a feeling of being watched. 
\item \textbf{Sight:} Object pulses slightly and is covered with secretion. 
\item \textbf{Perception:} Tentacle or appendage hidden within broad head. 
\item \textbf{Examination:} Appendage is eating orifice on neck-like muscle which leaps out to devour prey. Appendage range is 5 meters (one 
hex). 
\end{itemize}
\end{creature}

\hrulefill

\begin{creature}{Small}
\textbf{Restrictions:} Light vegetation; woods; forest. Non-corrosive, non-poisonous atmosphere. Temperature between 30\textdegree\ and 100\textdegree. \\\\
\textbf{Combat:} 5--8 \textbf{AY:} 4--7 \textbf{AG:} 6--8 \textbf{IN:} 8--12 \textbf{Initiative:} 20\% \\
\textbf{Composition:} Humanoid; mammalian 5\% \\
\textbf{Powers:} Invisibility 100\%; illusions 90\%; mind control 60\%; psionic communication 90\%; mental attack 30\% 
\begin{itemize}
\item \textbf{Warning:} A character with Mental Power of 4 or greater feels that he is being watched. 
\item \textbf{Sight:} When visible, appears to be a tiny humanoid one meter tall with large (8-inch long) pointed ears. Has two tentacles instead of 
arms. 
\end{itemize}
\end{creature}

\hrulefill

\begin{creature}{Large}
Restrictions; Woods; forest; jungle; marsh. Non-poisonous, non-corrosive atmosphere. Gravity of 0.6 to 2.6. Temperature between 
50\textdegree\ and 100\textdegree. \\\\
\textbf{Combat:} 9--12 \textbf{AY:} 4--6 \textbf{AG:} 6--8 \textbf{IN:} 2--4 \textbf{Initiative:} 10\% \\
\textbf{Composition:} Mammalian; Terran 20\% \\
\textbf{Powers:} Heightened smell 60\%; tunnelling 80\%; radar sense 50\% 
\begin{itemize}
\item \textbf{Warning:} You feel as though you are being watched. 
\item \textbf{Sight:} A gigantic, 3-meter tall shaggy teddy bear-like creature with 6 arm-like appendages (also serving as legs). Has large, toothy 
maw. Eyeless. 
\item \textbf{Perception:} Shape of teeth indicates that entity is omnivorous. 
\end{itemize}
\end{creature}

\hrulefill

\begin{creature}{Small}
\textbf{Restrictions:} Air. Must have atmosphere. Gravity 0.2 to 1.2 . \\\\
\textbf{Combat:} na \textbf{AY:} 12--15 \textbf{AG:} 6 \textbf{IN:} 10--12 \textbf{Initiative:} 10\% \\
\textbf{Composition:} Carbon; non-carbon 20\% \\
\textbf{Powers:} Unknown propulsion 100\%; Psionic communication 60\%; heightened hearing 30\%; immaterial 100\% 
\begin{itemize}
\item \textbf{Warning:} One character cannot get an unknown tune out of his head. 
\item \textbf{Sight:} You see a murky gas cloud from which emanates a strange, almost musical sound. 
\item \textbf{Perception:} Gas cloud is non-hostile entity composed of energy. 
\item \textbf{Examination:} Entity dissipates if stunned. 
\end{itemize}
\end{creature}

\hrulefill

\begin{creature}{Man sized}
\textbf{Restrictions:} Non-volcanic land environ. Non-corrosive, non-poisonous atmosphere. Gravity of 0.4 to 1.6. Temperature between 40\textdegree\ 
and 120\textdegree. \\\\
\textbf{Combat:} 8--12 \textbf{AY:} 10--12 \textbf{AG:} 10--12 \textbf{IN:} 2--4 \textbf{Initiative:} 5\% \\
\textbf{Composition:} Humanoid; mammalian 40\% \\
\textbf{Powers:} Chameleon 100\%; illusions 80\%; heightened smell 90\%; mind control 10\%; Psionic communication 20\% 
\begin{itemize}
\item \textbf{Warning:} One or more members of your party have strangely disappeared from sight. 
\item \textbf{Sight:} In own shape, entity is 2-meter tall humanoid covered with long silken fur. Head features pointed ears, muzzle (with rows of 
sharp teeth) and whisker-like sensors much like cat whiskers. Long tail emanates from lumbar region. Hands and feet are partly 
tentacle, partly clawed (retractable). 
\item \textbf{Examination:} Entity is carnivore which will stalk party, adopting forms of members already killed until all are dead. Entity will then 
return to kills, eating some and stashing others. Entity is unable to communicate unless psionic communication is possessed and will 
have only the physical appearance of an object it is simulating, not its mental powers. 
\end{itemize}
\end{creature}

\hrulefill

\begin{creature}{2 hex}
\textbf{Restrictions:} Land environs only. Non-corrosive atmosphere. Gravity 1.2 to 2.6. Temperature 100\textdegree\ to 250\textdegree. \\\\
\textbf{Combat:} 11--13 \textbf{AY:} 1--3 \textbf{AG:} 6--9 \textbf{IN:} 9--12 \textbf{Initiative:} 5\% \\
\textbf{Composition:} Terran; protein 30\% \textbf{Powers:} Armored carapace 40\%; mind control 5\%; heightened vision 5\% 
\begin{itemize}
\item \textbf{Warning:} A large shape moves at the edge of the party's sight. 
\item \textbf{Sight:} You observe a large humanoid approximately 12 meters high. Though the torso, head, and arms seem Terran enough, the 
entity has four legs radiating from its torso in a spider-like manner. Primitive clothing covers the torso. 
\item \textbf{Perception:} Clothing hides the fact that torso structure is somewhat arachnid and features carapace. 
Creature will always communicate on an ``ignore'' result. 
\end{itemize}
\end{creature}

\hrulefill

\begin{creature}{2 hex}
\textbf{Restrictions:} Any land environ. Must have atmosphere. Gravity of 0.4 to 0.8. Temperature between 0\textdegree\ and 150\textdegree. \\\\
\textbf{Combat:} 12--15 \textbf{AY:} 10--12 \textbf{AG:} 10--12 \textbf{IN:} 10--12 \textbf{Initiative:} 20\% \\
\textbf{Composition:} Terran; protein 20\% \\
\textbf{Powers:} Capture 90\%; regeneration 60\%; heightened vision 50\%; psionic communication 20\%; Chitinous carapace 100\% 
\begin{itemize}
\item \textbf{Warning:} You hear an oddly pitched humming sound. 
\item \textbf{Sight:} Entity is a 3-meter tall, 8-meter long insectile being much like a praying mantis. Has large, strong mandibles as primary 
weapon. Front legs appear to have almost humanoid gripping members. Has compound eye. 
\item \textbf{Examination:} Entity is carnivorous and will either attempt to capture party members for food or will rip open their bodies and use 
them as repositories for larvae. 
\end{itemize}
\end{creature}

\hrulefill

\begin{creature}{Man size}
\textbf{Restrictions:} Any land environ. Any type of atmosphere. Gravity of 0.4 to 1.0. Temperature between 50\textdegree\ and 250\textdegree. \textbf{Combat:} 2--3 \textbf{AY:} 
11--12 \textbf{AG:} 5--8 \textbf{IN:} 7--10 \textbf{Initiative:} 15\% \\
\textbf{Composition:} Protein; carbon 5\% \\
\textbf{Powers:} Swarm mind 100\% psionic communication 60\%; mental attack 25\%; mind control 80\%; poison discharge 100\% 
\begin{itemize}
\item \textbf{Warning:} You see a glowing dust cloud hanging in the air. 
\item \textbf{Sight:} Cloud is filled with sparkling lights. Appears gaseous. 
\item \textbf{Perception:} Lights are actually tiny firefly sized insectile entities, which flit about within a cloud of gases that they discharge in flight. 
Gases are poisonous. 
\item \textbf{Examination:} If a character with a Mental Power of 4 or greater is within 5 meters (same hex) of the cloud, one or more of the entities 
which make up the cloud will land on him and lay its eggs. Treat eggs as acidic. 

\end{itemize}
\end{creature}

%%\end{multicols}

\hrulefill



\subsection*{EXPLANATIONS OF CREATURE POWERS}


A power improves or specializes a creature's attributes, giving it
advantages when interacting with characters.  The GM secretly
determines which powers a creature will possess by rolling percentile
dice once for each power listed in the creature's description. The GM
may use any of these powers in a logical manner when designing his own
creatures and may introduce new powers of his own devising to the
already existing creatures. Certain powers provide a creature with
psionic skills, like those available to the characters. A creature
that uses such powers may also suffer psionic backlash (see GM guide
\ref{GM-sec:psibacklash}).  Explanations of all creatures powers
follow, in alphabetical order.  Each power lists one of the four
creature description categories (warning, sight, etc.,). Unless the
power is used in such a way that it is immediately obvious to the
players, it is revealed to them when they are read the listed
description.

\begin{description}
\item[Acid] (Perception). Substance on some part of creature's
  exterior will eat through expedition suit in one Action Round and, at
  the GM's discretion, through armor at a slow rate. May be combined
  with poison (see below).

\item[Acidic Circulatory Fluid] (Examination). Identical to acid;
  released if creature is punctured.

\item[Adhesive Secretion] (Perception). Skin, suckers, or web
  strands contain sticky substance which easily restrains any individual
  the creature attacks. Hits received by a character attacked by a
  creature with adhesive secretion and \emph{capture} are not reduced as
  listed in the capture description.

\item[Armored Carapace] (Perception). Armor with beam projectile
  Defense Strengths ranging from \textbf{3} to \textbf{5}.

\item[Balance Disruption] (Examination). Ambient energy emissions
  that scrambles the human sense of balance. Roll one die; every
  character with a Mental Power Rating less than the die result
  immediately suffers the effects of shock (see GM guide
  \ref{GM-sec:shock-check}). This is not considered an attack; it occurs
  automatically during the creature's turn of the Action Round.

\item[Camouflage] (Sight). The creature may only be seen at half the
  normal range or less. Add \textbf{5} to the creature's Initiative Percentage.

\item[Capture] (Perception). When attacking a character, reduce any
  hits received by the character by \textbf{3}. If the adjusted number
  is \textbf{0} or higher, the character is considered restrained by the
  creature ( suffering the adjusted number of hits), and remains
  restrained until the creature is stunned, passes out or dies.

\item[Chameleon] (Perception). The creature may assume the form of
  any living entity within its sight. If the party is unaware of the
  creature, it may appear to the party as one of their own when they see
  it. The creature's Combat and Agility Ratings are \emph{halved} when
  assuming another form. One Action Round is required for the creature
  to change shape (it may not attack or move while changing).
  
\item[Chitinous Carapace] (Perception). Armor with a projectile
  Defense Strength ranging from \textbf{3} to \textbf{5} and a Beam Defense Strength of \textbf{1}.

\item[Dart Shooting] (Perception). May attack by projecting quills
  or other natural projectiles. Its range (in hexes) equals one half of
  its Combat Rating (rounded down).  The Hit Strength of a dart equals
  half the creature's Combat Rating minus the range.

\item[Edible] (Examination).  Fit for consumption by humans.

\item[Energy absorption] (Examination).  Stun weapons and beam
  weapons have no effect on the creature.

\item[Energy Blast] (Perception). Emits powerful blasts of heat,
  electricity or radioactivity. The creature may attack one character at
  a range (in hexes) equal to its Combat Rating. The attack is resolved
  by subtracting the range from the creature's Combat Rating (only).
  Roll one die, add the modified rating to the result, and locate the
  sum on the Hit Table to determine the hits received by the character.
  Armor may be used to aid defense.

\item[Energy Drain] (Perception). The creature lives off the
  surrounding electrical energy. At the beginning of each Action Round,
  one item belonging to the party that is within a number of hexes of
  the creature equal to or less than its Combat Rating immediately loses
  all its power. This will affect beam weapons, scientific equipment,
  and any other electrical/electronic devices. The creature will always
  move toward such items.

\item[Energy Ingestion] (Perception). The Combat Rating of the
  creature is \emph{increased} by the number of hits it receives. The
  Agility Rating is reduced normally. The first time the creature
  ingests energy, its Aggression Rating is increased to \textbf{12}. The
  creature is killed if it is attacked with a stun weapon.

\item[Entrapment] (Perception). The creature has constructed or is
  able to construct simple traps to capture unwary passers by. A
  creature with tunnelling creates camouflaged pits. A creature with web
  shooting creates a net or large web. If the party is \emph{unaware} of
  such a creature, the GM should have one character (determined at
  random) fall into one of the traps.

\item[Flight] (Sight). The creature is winged and capable of flight.
  When flying, its Movement Rate is three times its current Agility
  Rating. Its normal Agility Rating is always used when calculating its
  strength. If the creature's Agility Rating is reduced to less than
  half its original value, the creature can no longer fly.  Heightened
  Hearing/Smelling (Examination). The creature may see twice as far as
  humans can. It sees as well in darkness as humans do in the light. The
  creature is never unaware during an encounter and no character may
  successfully perform an ambush skill task against it.

\item[Illusions] (Perception). May create immaterial, harmless but
  entirely convincing visual illusions of anything within its sight. The
  creature may not move or attack while doing so.

\item[Immaterial] (Perception). The creature possesses no solid form
  and is not affected by projectile fire, beam fire, poisons, or close
  attack.  It is visible and may suffer the effects of a stun weapon.
  The GM may allow it to pass through solid objects.

\item[Inky Discharge] (Examination). Emits opaque substance that
  creates total darkness, Emission counts as the creature's attack for
  one Action Round. The discharge spreads one hex in every direction
  each Action Round (two hexes in Round of Emission). Usually restricted
  to underwater use.

\item[Invisibility] (Perception). The creature is not placed on the
  Action Display. The GM notes the position and movement of the creature
  in his mind, informing the characters when the creature performs any
  visible actions. The creature is not immaterial, it may be affected by
  attack and may not move through solid objects.

\item[Kinetic Absorption] (Perception). Projectile weapons and blows
  have no effect on the creature.

\item[Mental Attack] (Perception). If the creature is not moving, it
  may attack by directing a blast of psionic energy at one character.
  Roll one die. If the result is greater than the character's Mental
  Power Rating, the character is stunned for a number of Action Rounds
  equal to the difference.Exception: If a 10 is rolled, the character's
  Endurance Rating is immediately reduced to \item[0] (he passes out).
  If a 1 is rolled, roll percentile dice, locate the result on the
  psionic Backlash Table (GM guide \ref{GM-sec:psibacklashtable}) and
  apply the outcome to the creature.

\item[Mind Control] (Examination). The creature possesses this skill
  at a level ranging from \textbf{6} to \textbf{10} and a Mental Power
  Rating of \textbf{4} to \textbf{7} (see GM guide
  \ref{GM-sec:psionicskills}).

\item[Multiple Attack] (Perception). May attack each and every
  character within its range once during its turn of the Action Round.

\item[Poison] (Examination). Skin or other exterior area is coated
  with a toxin (to humans at least). If the character touches a poisoned
  area with his bare skin, the GM should implement one or more of the
  following:

  \begin{itemize}
  \item \textbf{Virulent.} The character loses one point from his
    Endurance Rating each Action Round. When that reaches \textbf{0} (and the
    character passes out), he loses one point off his Strength Rating
    each Action Round. When that reaches \textbf{0} , the character is dead.
  \item \textbf{Virus.} Same effects as virulent poison. However, the
    unit of time per point lost is increased to one minute, one hour or
    one day, at the GM's discretion.
  \item \textbf{Paralytic.} The character's Dexterity and Agility
    Ratings are immediately reduced to \textbf{0} (he may not move or handle
    devices at all). If he receives successful treatment, he heals at
    four times the normal rate.
  \item \textbf{Psychotic.} The character immediately goes into fits of
    hallucination, paranoia and mindless babble. The GM should control
    the character when it seems appropriate. These effects cease
    immediately upon successful diagnosis and treatment. No
    characteristics are affected.
  \item \textbf{Spore.} The character suffers no effects at all until
    some future moment determined by the GM. At that point, i mplement
    the effects of any of the above poisons.
  \item \textbf{Infectious.} Any of the above poisons may be made
    contagious at the GM's discretion, and will spread to any character
    that is exposed to it.  Successful diagnosis of a poison stops any
    additional damage from occurring to the victim. Treatment must
    succeed to reverse the effects of poison (to begin the healing
    process). The GM may wish to reduce the Skill Level of a character
    attempting to diagnose or treat a character suffering from alien
    poisons.
  \end{itemize}

\item[Poison Attack] (Perception). Toxin on the creature's primary
  attack appendage (claw, stinger, tail, fangs, quills, etc.). If a
  character's protective attire is punctured when attacked, he suffers
  the effects of poison (see above).

\item[Poison Discharge] (Examination). Creature may attack by
  emitting toxin in gas or liquid form from its mouth or other orifice.
  Substance may be projected a number, of hexes equal to half the
  creature's Combat Rating. A character in a hex entered by the
  substance is poisoned if not wearing protective attire, see preceding.
  
\item[Psionic Communication] (Examination). The creature possesses
  this skill at a level ranging from \textbf{6} to \textbf{9} and a
  Mental Power Rating ranging from \textbf{4} to \textbf{7} (see GM
  guide \ref{GM-sec:psionicskills}).

\item[Psychokinesis] (Examination). The creature possesses this
  skill at a level ranging from \textbf{6} to \textbf{10} and a Mental
  Power Rating ranging from \textbf{4} to \textbf{7} (see GM guide
  \ref{GM-sec:psionicskills}).

\item[Radar Sense] (Perception). The creature ``sees'' by emitting
  energy waves. The creature may be visually blind but sees as well as a
  human and is not affected by darkness.

\item[Reflective Carapace] (Perception). Armor with beam Defense
  Strength ranging from \textbf{3} to \textbf{5} and with a projectile Defense Strength of
  \textbf{1} or \textbf{2}.

\item[Regeneration] (Perception). The creature regains one lost
  Combat or Agility Point at the beginning of each Action Round (unless
  the Rating has been reduced to \textbf{0}).

\item[Swarm Mind] (Perception). All the creatures in the encounter
  share one mind. What one sees or feels, all see or feel (although hits
  suffered by one are not suffered by all). If more than one creature is
  eligible to attack a single character in one Action Round, they may
  combine their strengths into one total (instead of attacking
  separately).

\item[Time Dilation] (Perception). The creature may conduct two
  attacks (against the same character or different characters) during
  its turn of an Action Round. Its Agility Rating is subtracted from the
  chance to hit the creature with ranged fire.

\item[Tunnelling] (Perception). The creature has a network of
  tunnels with various openings determined by the GM that it may move
  through at its full Agility Rating. The creature may create new
  tunnels at the rate of one hex per every two Action Rounds (the GM may
  reduce or increase this rate). One Action Round must be spent for the
  creature to create an opening into or out of the ground.

\item[Unknown Propulsion] (Sight). Floats with no apparent means of
  support. The creature may move its full Agility Rating each Action
  Round, regardless of what is in its hex or along its intended path
  (however, it may not move through walls). Until its Agility Rating is
  reduced to \textbf{0}, the creature may always move at its full Agility Rating.

\item[Vegetation Control] (Examination). The creature is one with
  the vegetation around it. Any movable parts of plants and trees
  (branches, vines, roots, etc.) may be called upon to entwine or strike
  one character as the creature's attack for an Action Round.

\item[Web Shooting] (Perception). The creature spins webs and shoots
  web strands to entwine its prey. May conduct a close attack (see GM
  guide \ref{GM-sec:close-combat}) from a distance of three hexes or
  less.
\end{description}

\section{Non-Player Characters}
\label{sec:non-play-char}
% [48.0]

The GM should choose one of the following non-player characters when
an NPC encounter occurs. Being careful not to disrupt the adventure in
progress (unless the GM wishes to do so), these NPC's are extremely
useful in spicing up an otherwise slow-moving section of the adventure
or to set up plot lines which the players might want to follow up in
the future.

These NPC's are presented in a format, which gives the GM enough
information to play them on the spur of the moment, yet does not
restrict the GM by being too specific. The GM is free to use these
listings as he sees fit; he should change whatever does not suit his
tastes.

The Characteristic Ratings of each NPC are given as letters, which
represent ranges for each characteristic. Before actually placing the
NPC's into play, the GM should choose a value for each characteristic
from the range indicated in the table \vpageref{tab:npcchars}.


\begin{table}[htbp]
  \centering
  \fbox{%
    \begin{minipage}{3.5in}
      \centering
      \caption{NPC Characteristic Chart}
      \label{tab:npcchars}
      
      \medskip
      
      \begin{tabular}{lll}
        LETTER & ALL CHARACTERISTICS & MENTAL\\
        & EXCEPT MENTAL POWER & POWER \\
        \rowcolor{gray}
        \textbf{A} & Poor (1,2) & 0 \\
        \textbf{B} & Mediocre (3, 4) & 1, 2 \\
        \rowcolor{gray}
        \textbf{C} & Average (5,6) & 3 \\
        \textbf{D} & Good (7, 8, 9) & 4, 5 \\
        \rowcolor{gray}
        \textbf{E} & Exceptional (10, 11, 12) & 6
      \end{tabular}
    \end{minipage}}
\end{table}

An NPC's Environ Skill Level is for the environ in which he is
encountered. If in a spaceport, the GM chooses a home environ for the
NPC. All other Environ and Gravity Skill Levels for the NPC are
determined in accordance with GM guide \ref{GM-sec:habitat-table-use}.

Some NPC encounters involve a group of NPC's. In such a case, the GM
rolls one die to determine how many there are.

\newcounter{npc}
\newenvironment{npc}[3]{%
\refstepcounter{npc}\paragraph{NPC Nr. #1\thenpc:}

#2\\
\emph{#3}\\
}{%

\bigskip

}

\subsection{Common NPCs}
\label{sec:common-npcs}
% [48.1]

\begin{npc}{C}{Drug Dealer}{Spaceport}
  \textbf{ST} (B) \textbf{EN} (C) \textbf{DX} (C) \textbf{AY} (B)
  \textbf{IN} (C) \textbf{MP} (B) \textbf{EM} (A) \textbf{LD} (B)
  \textbf{AG} (C) \textbf{SS:} Lower middle class \\
  \textbf{Skills:} Urban \textbf{2}; Grav \textbf{HY-1}; Temp
  \textbf{HO}; Environ \textbf{2}; Trading \textbf{2}; Transport Econ
  \textbf{4}; Handguns \textbf{2}; Law \textbf{1}; Ambush
  \textbf{1}. \\ 
  \textbf{Possessions:} Plastic pistol. \\
\begin{itemize}
\item \textbf{First Description:} As the characters are preparing to
  leave on their voyage, they are approached by a man who asks if he
  could talk to them in private.
\item \textbf{GM's Description:} This individual is attempting to get
  past customs officials with a great deal of illegal drugs. He has
  such a vast amount he cannot get all of them on his ship. Without
  telling the party what the shipment is, he asks. them if they would
  help him deliver an over order of material. He will pay them
  handsomely and asks their discretion. If the party shows any
  inclination to turn down the offer, the man will withdraw, not
  wanting to arouse suspicion.
\end{itemize}
\end{npc}

\hrulefill

\begin{npc}{C}{Diplomat}{Spaceport}
  \textbf{ST} (B) \textbf{EN} (B) \textbf{DX} (C) \textbf{AY} (B)
  \textbf{IN} (D) \textbf{MP} (C) \textbf{EM} (E) \textbf{LD} (C)
  \textbf{AG} (A) \textbf{SS:} Skilled tech \\
  \textbf{Skills:} Urban \textbf{3}; Grav \textbf{EX-1}; Temp
  \textbf{HO}; Environ \textbf{4}; Diplomacy \textbf{6}; Linguistics
  \textbf{8}; Law \textbf{4}; Streetwise \textbf{2}; Laser
  pistol \textbf{1}. \\
  \textbf{Possessions:} Audio-sealed case, laser pistol, translator.
\begin{itemize}
\item \textbf{First Description:} As the characters prepare to leave the
  spaceport, they are called into the security office where they meet
  this individual.
\item \textbf{GM's Description:} This person is a diplomat from the
  planet the characters are on to the planet for which they are bound,
  He needs immediate transport and asks the characters if they are
  willing to take him on. He is all he seems to be; the characters
  will be rewarded financially for their efforts on his behalf. If
  they refuse, no harm will come to them in any way.
\end{itemize}
\end{npc}

\hrulefill

\begin{npc}{C}{Scholar}{Spaceport}
  \textbf{ST} (B) \textbf{EN} (A) \textbf{DX} (A) \textbf{AY} (A)
  \textbf{IN} (D) \textbf{MP} (B) \textbf{EM} (C) \textbf{LD} (A)
  \textbf{AG} (A) \textbf{SS:} Family dominates planet \\ 
  \textbf{Skills:} Urban \textbf{2}; Grav \textbf{LT-1}; Temp
  \textbf{NL}; Environ \textbf{1}; Economics \textbf{5}; Mining
  \textbf{4}; Geology \textbf{3}; Physics \textbf{1}. \\ 
  \textbf{Possessions:} Camera.
\begin{itemize}
\item \textbf{First Description:} As the characters load or unload
  their ship in port, they see someone taking pictures of them.
\item \textbf{GM's Description:} This is a scholar on vacation who has
  chosen this planet to visit. The character's ship looked
  interesting, and he wanted a picture.
\end{itemize}
\end{npc}

\hrulefill

\begin{npc}{C}{Smuggler}{Spaceport}
\textbf{ST} (B) \textbf{EN} (C) \textbf{DX} (C) \textbf{AY} (B)
\textbf{IN} (C) \textbf{MP} (B) \textbf{EM} (B) \textbf{LD} (C)
\textbf{AG} (D) \textbf{SS:}Independent trading family \\ 
\textbf{Skills:} Urban \textbf{2}; Grav \textbf{LT-1}; Temp
\textbf{CD}; Environ \textbf{2}; Handguns \textbf{2}; Automobile
\textbf{1}; Forgery/Counterfeiting \textbf{1}. \\ 
\textbf{Possessions:} Stun gun. 
\begin{itemize}
\item \textbf{First Description:} The characters see a person shoving
  large boxes down a waste disposal chute.
\item \textbf{GM's Description:} The person is a smuggler whose
  partner is downstairs at the other end of the chute waiting for the
  contraband.  This individual is very jumpy and will not hesitate to
  fire upon the party if approached too closely.
\end{itemize}
\end{npc}

\hrulefill

\begin{npc}{C}{Trader}{Spaceport}
  \textbf{ST} (B) \textbf{EN} (B) \textbf{DX} (C) \textbf{AY} (B)
  \textbf{IN} (D) \textbf{MP} (B) \textbf{EM} (C) \textbf{LD} (C)
  \textbf{AG} (C) \textbf{SS:}Independent trading family \\ 
  \textbf{Skills:} Urban \textbf{2}; Grav \textbf{NW-1}; Temp
  \textbf{CD}; Environ \textbf{2}; Trading \textbf{6}; Economics
  \textbf{3}; Handguns \textbf{2}; Pilot \textbf{2}; Streetwise
  \textbf{3}; Diplomacy \textbf{2}; Shuttle \textbf{1};  
  Planetology \textbf{4}. \\ 
  \textbf{Possessions:} Stun pistol, portable computer, Geoscanner,
  expedition suit, camera, trade stuffs.  
  \begin{itemize}
  \item \textbf{First Description:} As the characters are talking to
    the booking agent at the Starport, a distinguished looking person
    approaches them and asks for a minute of their time.
  \item \textbf{GM's Description:} This person is an important
    interstellar trader who has overheard the players state their next
    port of call. He may ask them to deliver something; find their
    ship valuable for a separate trip after the next; ask them to pick
    something up and deliver it on further; etc. He is what he appears
    to be.
  \end{itemize}
\end{npc}

\hrulefill

\begin{npc}{C}{Reporter}{Spaceport, Urban}
  \textbf{ST} (B) \textbf{EN} (B) \textbf{DX} (C) \textbf{AY} (C)
  \textbf{IN} (C) \textbf{MP} (B) \textbf{EM} (C) \textbf{LD} (A)
  \textbf{AG} (C) \textbf{SS:} Fallen on hard times \\
  \textbf{Skills:} Urban \textbf{2}; Grav \textbf{LT-1}; Temp
  \textbf{NL}; Environ \textbf{2}; Automobile \textbf{2}; Direct-lift
  \textbf{1}; Disguise \textbf{1}; Streetwise \textbf{4}; Diplomacy
  \textbf{3}; Linguistics 1 . \\ 
  \textbf{Possessions:} Automobile, camera, audio recording
  equipment. 
  \begin{itemize}
  \item \textbf{First Description:} While the party is in the
    spaceport, they are witness to an important incident of the GM's
    creation (a murder, the arrival of a celebrity, an escaped
    criminal). They are then accosted by an individual asking them a
    lot of questions.
  \item \textbf{GM's Description:} This reporter works for a planetary
    news service and will press the characters for all the information
    they have to give and then some. It will become very annoying very
    quickly as this individual is both tenacious and intimidating.
\end{itemize}
\end{npc}

\hrulefill

\begin{npc}{C}{Religious Fanatics (one die)}{Spaceport, Urban}
  \textbf{ST} (B) \textbf{EN} (B) \textbf{DX} (B) \textbf{AY} (A)
  \textbf{IN} (C) \textbf{MP} (B) \textbf{EM} (C) \textbf{LD} (A)
  \textbf{AG} (A) \textbf{SS:} Poor colonist family \\ 
  \textbf{Skills:} Urban \textbf{1}; Grav \textbf{LT-1}; Temp
  \textbf{NL}; Environ \textbf{2}; Recruiting \textbf{1}; Streetwise
  \textbf{2}; Diplomacy \textbf{1}. \\ 
  \textbf{Possessions:} None. 
  \begin{itemize}
  \item \textbf{First Description:} The characters are approached by a
    group of people who begin to talk about the party's obvious lack of
    belief.
  \item \textbf{GM's Description:} These apparent fanatics are actually
    a front for a political organization skirting taxes by using a
    religious front.  The group will try to sell the party books or
    buttons or some worthless junk for exorbitant prices. They will give
    the party a hassle if nothing is purchased. If the party refuses any
    offers at all, the GM must determine whether this group becomes
    unruly.
  \end{itemize}
\end{npc}

\hrulefill

\begin{npc}{C}{Robbery Victim}{Urban, Suburban}
  \textbf{ST} (B) \textbf{EN} (B) \textbf{DX} (B) \textbf{AY} (B)
  \textbf{IN} (C) \textbf{MP} (A) \textbf{EM} (B) \textbf{LD} (A)
  \textbf{AG} (A) \textbf{SS:} Local establishment \\ 
  \textbf{Skills:} Urban \textbf{1}; Grav \textbf{LT-1}; Temp
  \textbf{NL}; Environ \textbf{2}; Law \textbf{1}; Automobile
  \textbf{1}. \\ 
  \textbf{Possessions:} An automobile. 
  \begin{itemize}
  \item \textbf{First Description:} The characters find the
    unconscious body of a male lying in an alleyway.
  \item \textbf{GM's Description:} The man has been mugged and robbed.
    He is a lawyer who might have been carrying valuable papers (GM's
    discretion) which were also snatched. At first he will be groggy
    with a small chance of mistaking the characters for his
    assailants. After fully awakening, he will need the characters'
    help to reach the authorities.
  \end{itemize}
\end{npc}

\hrulefill

\begin{npc}{C}{Fence}{Urban}
  \textbf{ST} (B) \textbf{EN} (B) \textbf{DX} (C) \textbf{AY} (C) \textbf{IN} (D) \textbf{MP} (A) \textbf{EM} (B) \textbf{LD} (C) \textbf{AG} (C) \textbf{SS:} Lower middle class \\
  \textbf{Skills:} Urban \textbf{2}; Grav \textbf{LT-1}; Temp  \textbf{NL}; Environ \textbf{2}; Streetwise \textbf{4}; Gambling \textbf{2}; Disguise \textbf{3}; Forgery/Counterfeiting \textbf{3}; Handguns \textbf{3}. \\
  \textbf{Possessions:} False identity, plastic pistol, various stolen items. 
  \begin{itemize}
  \item \textbf{First Description:} As the characters are strolling down a street, they are approached by an individual who tries to entice them to buy 
    some goods. 
  \item \textbf{GM's Description:} At first, the items offered by this individual will be typical: cameras, watches, jewellery, etc. If the characters 
    express enough interest, the fence will begin to offer harder to find high-tech items for very low prices. The fence is not hostile and will 
    back off if threatened. 
  \end{itemize}
\end{npc}

\hrulefill

\begin{npc}{C}{Street Person}{Urban, Suburban}
\textbf{ST} (B) \textbf{EN} (B) \textbf{DX} (B) \textbf{AY} (C) \textbf{IN} (A) \textbf{MP} (A) \textbf{EM} (B) \textbf{LD} (B) \textbf{AG} (C) \textbf{SS:} Unfostered orphan \\
\textbf{Skills:} Urban \textbf{1}; Grav \textbf{HY-1}; Temp \textbf{HO};Environ \textbf{2}; Streetwise \textbf{4}; Gambling \textbf{2}; Automobile \textbf{1}; Ambush \textbf{1}. \\
\textbf{Possessions:} None. 
\begin{itemize}
\item \textbf{First Description:}  The party is spoken to by a figure in a shadowy doorway as they travel through the streets. 
\item \textbf{GM's Description:} This figure could be any number of things: a beggar asking for money; a drug seller hawking his wares; a hustler 
looking for an easy mark in a card game; a pimp. How far the street person will go to engage the party is up to the GM. 
\end{itemize}
\end{npc}

\hrulefill

\begin{npc}{C}{Sanitation Engineer}{Urban, Suburban}
\textbf{ST} (A) \textbf{EN} (A) \textbf{DX} (B) \textbf{AY} (B) \textbf{IN} (A) \textbf{MP} (A) \textbf{EM} (B) \textbf{LD} (A) \textbf{AG} (A) \textbf{SS:} Lower middle class \\
\textbf{Skills:} Urban \textbf{1}; Grav \textbf{LT-1}; Temp  \textbf{NL}; Environ \textbf{1}; Streetwise \textbf{1}. \\
\textbf{Possessions:} None. 
\begin{itemize}
\item \textbf{First Description:} The characters see a person walking down the street next to a moving truck. 
\item \textbf{GM's Description:} This is a sanitation engineer collecting garbage. If this has already been played, the person could be something 
else in disguise --- a lawman, perhaps, or a spy. 
\end{itemize}
\end{npc}

\hrulefill

\begin{npc}{C}{Thieves (one die)}{Urban, Suburban}
\textbf{ST} (B) \textbf{EN} (C) \textbf{DX} (B) \textbf{AY} (B) \textbf{IN} (C) \textbf{MP} (A) \textbf{EM} (B) \textbf{LD} (C) \textbf{AG} (C) \textbf{SS:} Oppressed minority \\
\textbf{Skills:} Urban \textbf{1}; Grav \textbf{LT-1}; Temp  \textbf{NL}; Environ \textbf{2}; Streetwise \textbf{1}; Automobile \textbf{3}; Laser Pistol \textbf{2}; Blades \textbf{2}; Ambush \textbf{2}; 
Unarmed Combat \textbf{1}. \\
\textbf{Possessions:} Automobiles, laser pistols, knives. 
\begin{itemize}
\item \textbf{First Description:} As the party travels in their vehicle down a semi-deserted street in a city, they are sideswiped by another car and 
stopped. As they get out, another vehicle pulls up behind them. 
\item \textbf{GM's Description:} These individuals are part of a highly organized auto theft ring. They will pull out their weapons, order the party to 
stand against the wall, and drive off with the party's vehicle, also taking any equipment inside. If the party puts up no resistance, there 
will be no cause for combat. This gang is willing to engage in a quick fire fight and then escape in their vehicles. 
\end{itemize}
\end{npc}

\hrulefill

\begin{npc}{C}{Thieves (one die)}{Urban}
\textbf{ST} (C) \textbf{EN} (B) \textbf{DX} (C) \textbf{EN} (B) \textbf{IN} (B) \textbf{MP} (A) \textbf{EM} (A) \textbf{LD} (C) \textbf{AG} (D) \textbf{SS:} Poor colonist family \\
\textbf{Skills:} Urban \textbf{3}; Grav \textbf{LT-1}; TempML;Environ \textbf{2}; Streetwise \textbf{3}; Battlefield \textbf{2}; Jetpack \textbf{1}; 
Handguns \textbf{2}; Blades \textbf{2}; Ambush \textbf{3}; Unarmed Combat \textbf{1}. \\
\textbf{Possessions:} Pistols, daggers, stun guns. 
\begin{itemize}
\item \textbf{First Description:} The first description in this case may be a hit on the head; the group of thieves will attempt to ambush and rob the 
party. 
\item \textbf{GM's Description:} Basically the same as preceding; the thieves will not kill the party except in open combat. If the party gives in, no 
one will be seriously harmed. 
\end{itemize}
\end{npc}

\hrulefill

\begin{npc}{C}{Travelling Doctor}{Road}
\textbf{ST} (A) \textbf{EN} (B) \textbf{DX} (C) \textbf{AY} (A) \textbf{IN} (C) \textbf{MP} (B) \textbf{EM} (C) \textbf{LD} (A) \textbf{AG} (A) \textbf{SS:} Skilled tech \\
\textbf{Skills:} Urban \textbf{1}; Grav \textbf{LT-1}; Temp  \textbf{NL}; Environ \textbf{2}; Diplomacy \textbf{2}; Chemistry \textbf{1}; Diagnostics \textbf{3}; Treatment \textbf{2};Biology \textbf{1}; Programming \textbf{1}; 
Compu/Robot Tech \textbf{1}. \\
\textbf{Possessions:} ATV (Civ Level 6), robot with medical systems (Civ Level 7). 
\begin{itemize}
\item \textbf{First Description:} As the party travels along, they are hailed by a man and his robot by the side of the road. Their vehicle is parked 
nearby. 
\item \textbf{GM's Description:} This doctor's vehicle has broken down and he cannot fix it. He roams the countryside helping the populace and 
keeping them fit. He needs the party for either help or transportation to the local town. He will offer money or services if the party 
seems hesitant. 
\end{itemize}
\end{npc}

\hrulefill

\begin{npc}{C}{Recalcitrant Farm Boys (two dice)}{Small settlement}
\textbf{ST} (B) \textbf{EN} (B) \textbf{DX} (C) \textbf{AY} (B) \textbf{IN} (A) \textbf{MP} (B) \textbf{EM} (B) \textbf{LD} (B) \textbf{AG} (C) \textbf{SS:} Communal colony \\
\textbf{Skills:} Urban \textbf{0}; Grav \textbf{LT-1}; Temp  \textbf{CD}; Environ \textbf{4}; Blades \textbf{2}; Longarms \textbf{1}; Unarmed Combat \textbf{2}; Agriculture \textbf{4}. \\
\textbf{Possessions:} Horse cart, farm equipment, truck, rifle. 
\begin{itemize}
\item \textbf{First Description:} The party comes upon a small farming community, which is in obvious need of some new equipment. They are 
greeted by some of the locals. 
\item \textbf{GM's Description:} These are a bunch of very proud people. They separated themselves from the rest of the population to prove 
their project could be done without assistance. If the party tries to leave and get them help, or suggests they do that, the farmers, led by 
the young men, will hinder the party from leaving. They do not want the knowledge they are actually struggling to leak out. Depending 
on the circumstances, violence is not impossible. 
\end{itemize}
\end{npc}

\hrulefill

\begin{npc}{C}{Land Owner}{Environ}
\textbf{ST} (B) \textbf{EN} (B) \textbf{DX} (A) \textbf{AY} (A) \textbf{IN} (B) \textbf{MP} (A) \textbf{EM} (B) \textbf{LD} (A) \textbf{AG} (D) \textbf{SS:} Local establishment \\
\textbf{Skills:} Urban \textbf{0}; Grav \textbf{LT-1}; Temp  \textbf{NL}; Environ \textbf{3}; Trading \textbf{2}; Agriculture \textbf{3};Geology \textbf{1}; 
Longarms \textbf{1}. \\
\textbf{Possessions:} Truck, rifle, short wave radio. 
\begin{itemize}
\item \textbf{First Description:}  The party sees a sign stating "No Trespassing" followed shortly by a man in a truck pulling up behind them. 
\item \textbf{GM's Description:} The man is a loner who has retreated to his tract of land hidden from society and deeply resents intrusions by 
anyone. If the party does not leave (as he will insist they do) he will not hesitate to fire on them. He is a bit nuts. 
\end{itemize}
\end{npc}

\hrulefill

\begin{npc}{C}{Colonist}{Environ}
\textbf{ST} (B) \textbf{EN} (C) \textbf{DX} (B) \textbf{AY} (A) \textbf{IN} (B) \textbf{MP} (B) \textbf{EM} (A) \textbf{LD} (B) \textbf{AG} (B) \textbf{SS:} Poor colonist \\
\textbf{Skills:} Urban \textbf{0}; Grav \textbf{HY-2}; Temp \textbf{HO};Environ \textbf{4}; Trading \textbf{2}; Life Support \textbf{2}; Shuttle \textbf{1}; Agriculture \textbf{2}; Mining \textbf{2}; Glider \textbf{1}. \\
\textbf{Possessions:} Glider, knife, expedition suit, respirator helmet. 
\begin{itemize}
\item \textbf{First Description:} As the party travels cross-country, they find a person sitting on a rock looking very dejected. 
\item \textbf{GM's Description:} This person is very tired of being a colonist and wants to leave. He will attempt to convince the characters to take 
him along anywhere they are going. If they refuse, he will not get violent, just more depressed. 
\end{itemize}
\end{npc}

\hrulefill

\begin{npc}{C}{Miner}{Mountains}
\textbf{ST} (B) \textbf{EN} (C) \textbf{DX} (C) \textbf{AY} (B) \textbf{IN} (D) \textbf{MP} (A) \textbf{EM} (C) \textbf{LD} (C) \textbf{AG} (C) \textbf{SS:} Lower middle class \\
\textbf{Skills:} Urban \textbf{0}; Grav \textbf{HY-1}; Temp  \textbf{NL}; Environ \textbf{5}; Planetology \textbf{4}; Trading \textbf{3}; Sled \textbf{4}; Mining \textbf{3}; Geology \textbf{5}; Handguns \textbf{1}; Laser Pistol \textbf{3}. \\
\textbf{Possessions:} Jetpack, Expedition Suit, Auto Sled, Laser Pistol, Rifle, Stun Gun, Shortwave Radio, Geoscanner, Chemanalysis Lab, 
Mining Robot. 
\begin{itemize}
\item \textbf{First Description:} The party stumbles upon a very small mining operation with a few people and robots. 
\item \textbf{GM's Description:} This trader has discovered a very rich gem deposit, which he slowly mines, using the gems for barter to receive 
supplies. He will claim he mines a low-yield ore deposit, which barely pays for itself. He is hiding the real information because he likes 
living relatively alone and independent. 
\end{itemize}
\end{npc}

\hrulefill

\begin{npc}{C}{Lawman}{Road}
\textbf{ST} (C) \textbf{EN} (C) \textbf{DX} (D) \textbf{AY} (C) \textbf{IN} (B) \textbf{MP} (A) \textbf{EM} (B) \textbf{LD} (D) \textbf{AG} (D) \textbf{SS:} Skilled tech \\
\textbf{Skills:} Urban \textbf{0}; Grav \textbf{LT-2}; Temp  \textbf{NL}; Environ \textbf{4}; Machine Gun \textbf{2}; Longarms \textbf{2}; Handguns \textbf{4}; Helicopter \textbf{3}; Battlefield \textbf{2}. \\
\textbf{Possessions:} Helicopter, machine gun, needle pistol, impact armor, short wave radio. 
\begin{itemize}
\item \textbf{First Description:} As the party drives along a road, they are hailed and told to stop by an individual in a helicopter. 
\item \textbf{GM's Description:} The person is a lawman l ooking for a group who stole a vehicle like the one the party is riding in. If the party's 
papers and ID'S are in order, they will not be hindered; otherwise the lawman will detain and possibly arrest the party, depending on 
how suspicious the party is. 
\end{itemize}
\end{npc}

\hrulefill

\begin{npc}{C}{Villagers (two dice)}{Jungle, Forest}
\textbf{ST} (C) \textbf{EN} (C) \textbf{DX} (B) \textbf{AY} (B) \textbf{IN} (B) \textbf{MP} (A) \textbf{EM} (A) \textbf{LD} (B) \textbf{AG} (C) \textbf{SS:} Communal colony \\
\textbf{Skills:} Urban \textbf{0}; Grav \textbf{LT-1}; Temp \textbf{HO};Environ \textbf{5}; Animal Drawn \textbf{4}; Bows \textbf{4}; Blades \textbf{2}; Survival \textbf{4}; Ambush \textbf{2}. \\
\textbf{Possessions:} A blade weapon, a bow weapon, arrows. 
\begin{itemize}
\item \textbf{First Description:} As the characters are foraging through the undergrowth, they run into a group of backward natives, who are very 
curious about these strange looking individuals. 
\item \textbf{GM's Description:} The villagers are harmless unless provoked, in which case they will fight to kill. They have valuable information 
concerning the area (flora, fauna, gaudy mineral deposits) and, if communication is possible, they will serve as guides and helpers 
given good treatment by the party. 
\end{itemize}
\end{npc}

\hrulefill

\subsection{Rare NPCs}
\label{sec:rare-npcs}
% [48.2]

\setcounter{npc}{0}
\begin{npc}{R}{Psion}{Spaceport}
\textbf{ST} (B) \textbf{EN} (B) \textbf{DX} (B) \textbf{AY} (C) \textbf{IN} (D) \textbf{MP} (D) \textbf{EM} (B) \textbf{LD} (A) \textbf{AG} (C) \textbf{SS:} Independent trading family \\
\textbf{Skills:} Urban \textbf{4}; Grav \textbf{NW-3};Temp \textbf{CD}; Environs (all) \textbf{-4}; Handguns \textbf{1}; Pilot \textbf{1}; Navigation \textbf{6}; Psionic Boost \textbf{4}; Mind Control \textbf{3}; 
Psionic Comm \textbf{4}; Psion Tech \textbf{2}; Life Sense \textbf{2}; Glider \textbf{1}. \\
\textbf{Possessions:} Expedition suit, Stun Gun, Internal Gravity Web, Propulsion Gun. 
\begin{itemize}
\item \textbf{First Description:} In the course of tiding over between trips, the characters meet a Psion who has just had a falling out with her old 
ship and crew. 
\item \textbf{GM's Description:} The psion is a very beautiful woman who was relieved of her duties due to an ill-fated liaison with the pilot. Her 
career is chequered with this kind of behaviour; her love affairs affect her performance too much. She will attempt to hide this as she 
needs work, and will try to sell herself just on her skills, which are impressive. 
\end{itemize}
\end{npc}

\hrulefill

\begin{npc}{R}{Stowaway}{Spaceport (just before or after leaving)} 
\textbf{ST} (C) \textbf{EN} (C) \textbf{DX} (B) \textbf{AY} (C) \textbf{IN} (D) \textbf{MP} (A) \textbf{EM} (B) \textbf{LD} (B) \textbf{AG} (B) \textbf{SS:} Lower middle class family \\
\textbf{Skills:} Urban \textbf{1}; Grav \textbf{LT-1}; Temp \textbf{NL}; Environ \textbf{2}; Programming \textbf{6}; Chemistry \textbf{4}; Geology \textbf{3}; Asteroid Mining \textbf{2};Planetology \textbf{1}; Laser 
Pistol \textbf{1}. \\
\textbf{Possessions:} None. 
\begin{itemize}
\item \textbf{First Description:} In a pressurized part of the ship, the characters discover a stowaway. He is unarmed, with little inclination to fight. 
\item \textbf{GM's Description:} This human is trying to escape from an insane asylum. He has determined the planet, which the characters are 
leaving, is in danger of a huge seismic shock, which will cause great damage. The authorities locked him up to prevent panic; they 
think he is crazy; the GM should decide for himself. 
\end{itemize}
\end{npc}

\hrulefill

\begin{npc}{R}{Stowaway}{Spaceport}
\textbf{ST} (A) \textbf{EN} (B) \textbf{DX} (B) \textbf{AY} (B) \textbf{EM} (C) \textbf{MP} (A) \textbf{EM} (B) \textbf{LD} (A) \textbf{AG} (A) \textbf{SS:} Local establishment \\
\textbf{Skills:} Urban \textbf{0}; Grav \textbf{LT-1}; Temp \textbf{HO}; Environ \textbf{3}; Streetwise \textbf{1}; Glider \textbf{1}; Blades \textbf{1}. \\
\textbf{Possessions:} None. 
\begin{itemize}
\item \textbf{First Description:} In a stateroom, the characters find a lovely young girl attempting to hide in a closet. She is very docile and very 
nervous. 
\item \textbf{GM's Description:} The girl is running away from her parents who treat her fairly well. She simply has never been in space and wants 
to see what it's all about. 
\end{itemize}
\end{npc}

\hrulefill

\begin{npc}{R}{Clergyman}{Spaceport, Urban}
\textbf{ST} (B) \textbf{EN} (B) \textbf{DX} (C) \textbf{AY} (B) \textbf{IN} (B) \textbf{MP} (B) \textbf{EM} (B) \textbf{LD} (B) \textbf{AG} (C) \textbf{SS:} Skilled tech family \\
\textbf{Skills:} Urban \textbf{1}; Grav \textbf{LT-1}; Temp \textbf{HO};Environ \textbf{2}; Streetwise \textbf{2}; Disguise \textbf{4}; Forgery/ Counterfeiting \textbf{2}; Blades \textbf{1}. \\
\textbf{Possessions:} Knife. 
\begin{itemize}
\item \textbf{First Description:} The party is approached by a man wearing odd clothes but with a strange smile on his face. 
\item \textbf{GM's Description:} This individual will claim to be part of a local religion whose purpose is to try to convince people to stay home 
rather than wander around the galaxy. He will engage the characters in conversation for as long as they will listen. During this time, he 
will pick their pockets. If caught, he will feign amnesia (or some other malady) and play upon the party's sympathy. If it looks as if he 
will be turned in, he will fight, but only to free himself and run away. 
\end{itemize}
\end{npc}

\hrulefill

\begin{npc}{R}{Lawmen (one die)}{Spaceport, Urban}
\textbf{ST} (B) \textbf{EN} (C) \textbf{DX} (C) \textbf{AY} (B) \textbf{IN} (B) \textbf{MP} (A) \textbf{EM} (B) \textbf{LD} (C) \textbf{AG} (C) \textbf{SS:} Communal colony \\
\textbf{Skills:} Urban \textbf{1}; Grav \textbf{LT-1}; Temp \textbf{NL}; Environ \textbf{2}; Streetwise \textbf{1}; Laser Pistol \textbf{3}; Automobile \textbf{2}; Body Armor \textbf{1}. \\
\textbf{Possessions:} Laser pistol, stun pistol, needle pistol, reflective armor. 
\begin{itemize}
\item\textbf{First Description:} As the characters are strolling through the Starport, a figure rushes by them followed shortly by other figures ob- 
viously chasing the first one. 
\item\textbf{GM's Description:} These lawmen are chasing an escaped criminal and will enlist the aid of the characters if they wish. This will be 
done by asking the crowd in general to help stop him. If the party aids and the criminal is apprehended, the lawmen will be extremely 
grateful and offer to assist the party in any logical way they can. 
\end{itemize}
\end{npc}

\hrulefill

\begin{npc}{R}{Entertainer}{Urban, Town}
\textbf{ST} (A) \textbf{EN} (B) \textbf{DX} (C) \textbf{AY} (C) \textbf{IN} (B) \textbf{MP} (A) \textbf{EM} (B) \textbf{LD} (A) \textbf{AG} (A) \textbf{SS:} Family runs planetary corporation \\
\textbf{Skills:} Urban \textbf{0}; Grav \textbf{LT-1}; Temp \textbf{NL}; Environ \textbf{1}; Disguise \textbf{8}; Linguistics \textbf{3}; Streetwise \textbf{1}; Diplomacy \textbf{1}. \\
\textbf{Possessions:} None. 
\begin{itemize}
\item \textbf{First Description:} As the characters are walking down a street, a luxurious ground vehicle stops and out steps a well-dressed in- 
dividual. 
\item \textbf{GM's Description:} This person is an entertainer in an opera being performed tonight at this location. Unless the party makes some 
move to communicate with this person, the encounter will pass on by. If approached by the characters, the entertainer will gladly talk to 
them, being off-worlders as they are. 
\end{itemize}
\end{npc}

\hrulefill

\begin{npc}{R}{Runaway 'Bot}{City}
Civ Level 8 Robot (\#3) 
Systems: Aerodynamic, learning, self-activation, weapon, weapon targeting. 
\begin{itemize}
\item \textbf{First Description:} The party sees a bot flying through the air toward them, firing a laser at random. 
\item \textbf{GM's Description:} The bot has malfunctioned en route to its owner, and has taken flight thinking it is mining for certain minerals. It 
must be disabled and brought under control. 
\end{itemize}
\end{npc}

\hrulefill

\begin{npc}{R}{Police Vehicle}{Road}\label{sec:rare-npcs-8}
\textbf{ST} (C) \textbf{EN} (C) \textbf{DX} (C) \textbf{AY} (C) \textbf{IN} (B) \textbf{MP} (A) \textbf{EM} (C) \textbf{LD} (D) \textbf{AG} (C) \textbf{SS:} Skilled tech family \\
\textbf{Skills:} Urban \textbf{2}; Grav \textbf{HY-1}; Temp \textbf{HO};Environ \textbf{3}; Streetwise \textbf{3}; Battlefield \textbf{1}; Laser Pistol \textbf{4}; Handguns \textbf{2}; Longarms \textbf{2}; Unarmed 
Combat \textbf{1}; Automobile \textbf{3}; Body Armor \textbf{1}. \\
\textbf{Possessions:} Automobile, stun gun, laser pistol, rifle, impact armor, short wave radio, photographic equipment, jetpack. 
\begin{itemize}
\item \textbf{First Description:} If the party is driving, the vehicle pulls them over. If they are on foot, the vehicle pulls up next to them. 
\item \textbf{GM's Description:} These are lawmen who saw the party, became curious, and decided to investigate. Unless there are any 
discrepancies, they will just ask routine questions and depart. 
\end{itemize}
\end{npc}

\hrulefill

\begin{npc}{R}{Police Impostors}{Road}
\textbf{ST} (C) \textbf{EN} (B) \textbf{DX} (B) \textbf{AY} (B) \textbf{IN} (B) \textbf{MP} (A) \textbf{EM} (B) \textbf{LD} (B) \textbf{AG} (C) 
\textbf{SS:} Poor colonist family \\
\textbf{Skills:} See R\ref{sec:rare-npcs-8}. \\
\textbf{Possessions:} See R\ref{sec:rare-npcs-8}. 
\begin{itemize}
\item \textbf{First Description:} A police vehicle pulls the party over to the side of the road. 
\item \textbf{GM's Description:} A group of thieves have stolen this vehicle and are out to make the most of the situation. They will, at the very 
least, claim the party's vehicle has improper registration and attempt to impound the vehicle. If the party is not riding, they will attempt 
to appropriate the party's weapons and equipment, claiming their possession is illegal. They will fight if pushed far enough. 
\end{itemize}
\end{npc}

\hrulefill

\begin{npc}{R}{Rangers (one die + 2)}{Environ}
\textbf{ST} (C) \textbf{EN} (B) \textbf{DX} (C) \textbf{AY} (C) \textbf{IN} (B) \textbf{MP} (A) \textbf{EM} (B) \textbf{LD} (C) \textbf{AG} (D) \textbf{SS:} Family fallen on hard times \\
\textbf{Skills:} Urban \textbf{1}; Grav \textbf{LT-1}; Temp \textbf{NL}; Environ \textbf{4}; Battlefield \textbf{2}; Machine Guns \textbf{3}; Paint Guns \textbf{2}; Military Ground Vehicles \textbf{2}; Vehicle 
Tech \textbf{2}; Treatment \textbf{1}. \\
\textbf{Possessions:} ATV (Civ Level 8), submachine guns (Civ Level 5), paint guns (Civ Level 6), reflect/impact armor (Civ Level 6), vehicle 
kit, first aid kit. 
\begin{itemize}
\item \textbf{First Description:} As the party is traversing a marsh, they come upon another vehicle paralleling their course. 
\item \textbf{GM's Description:} This vehicle contains two rangers with the rest of them scattered in the marsh away from the party. Upon en- 
countering the characters, the rangers will regroup and view the party with great suspicion. The rangers are searching for a downed 
satellite containing valuable information. It is suspected a military power wanted the satellite and the rangers are wary the party may be 
involved. 
\end{itemize}
\end{npc}

\hrulefill

\begin{npc}{R}{Hunter}{Woods, Forest, Jungle}
\textbf{ST} (D) \textbf{EN} (C) \textbf{DX} (C) \textbf{AY} (C) \textbf{IN} (B) \textbf{MP} (A) \textbf{EM} (A) \textbf{LD} (C) \textbf{AG} (D) \textbf{SS:} Independent trading family \\
\textbf{Skills:} Urban \textbf{0}; Grav \textbf{HY-1}; Temp \textbf{HO};Environ \textbf{4}; Bow \textbf{6}; Longarms \textbf{6}; Ambush \textbf{7}; Unarmed Combat \textbf{5}; ATV \textbf{4}; Streetwise \textbf{1}; 
Weapon Tech \textbf{1}; Battlefield \textbf{1}. \\
\textbf{Possessions:} One of the following: long bow and arrows; rifle, Neuroscanner; weapon kit. 
\begin{itemize}
\item \textbf{First Description:} As the characters are travelling overland, they come upon a figure moving slowly through the brush. 
\item \textbf{GM's Description:} This is a hunter who is after illegal game. He is extremely wary to the authorities and will be hesitant to deal 
openly with the characters. If he is questioned about what kind of game he is hunting, he will lie or divert the question, because 
poaching carries extreme punishment. He will resist capture violently. 
\end{itemize}
\end{npc}

\hrulefill

\begin{npc}{R}{Archaeologist}{Ruins}
\textbf{ST} (B) \textbf{EN} (B) \textbf{DX} (B) \textbf{AY} (B) \textbf{IN} (E) \textbf{MP} (C) \textbf{EM} (C) \textbf{LD} (A) \textbf{AG} (B) \textbf{SS:} Family runs planetary corporation \\
\textbf{Skills:} Urban \textbf{1}; Grav \textbf{HY-2}; Temp \textbf{HO};Environ \textbf{4}; Handguns \textbf{1}; ATV \textbf{3}; Sled \textbf{2}; Diplomacy \textbf{1}; Agriculture \textbf{2}; Planetology \textbf{2}; 
Programming \textbf{2};Biology \textbf{1};Geology \textbf{4}; Construction \textbf{4}; Linguistics \textbf{3}. \\
\textbf{Possessions:} ATV, life support equipment, one laser pistol, Superoid camera, Chemanalysis Lab, Bioscanner. 
\begin{itemize}
\item \textbf{First Description:} As the characters are exploring ruins (or possibly just passing the time) they notice a woman carefully picking her 
way through some rocks. 
\item \textbf{GM's Description:} She is an archaeologist looking for information concerning elder civilizations. She is convinced man explored this 
part of the galaxy in a previous generation; these are the ruins of that exploration. She will calmly present her evidence to the party. 
She is totally wrong. 
\end{itemize}
\end{npc}

\hrulefill

\subsection{Unique NPCs}
\label{sec:unique-npcs}
% [48.3]

\setcounter{npc}{0}
\begin{npc}{U}{Civil Inspector}{Spaceport}
\textbf{ST} (A) \textbf{EN} (A) \textbf{DX} (B) \textbf{AY} (B) \textbf{IN} (C) \textbf{MP} (A) \textbf{EM} (C) \textbf{LD} (B) \textbf{AG} (C) \textbf{SS:} Lower middle class family \\
\textbf{Skills:} Urban \textbf{0}; Grav \textbf{LT-1}; Temp \textbf{NL}; Environ \textbf{3}; Streetwise \textbf{1}; Handguns \textbf{1}; Unarmed Combat \textbf{1}; Agriculture \textbf{1}; Planetology \textbf{2}. \\
\textbf{Possessions:} Stun gun, briefcase with forms and papers. 
\begin{itemize}
\item \textbf{First Description:} As the characters lounge in the spaceport bar, a man approaches them with a very curious look on his face. 
\item \textbf{GM's Description:} The man is an inspector looking for planetary exploiters and abusers. Men whose descriptions match the 
characters were seen in this bar just last week. He is highly suspicious of the party and will probably call in the lawmen unless truly 
convinced of his error. The odds of this happening are small. 
\end{itemize}
\end{npc}

\hrulefill

\begin{npc}{U}{Technician}{Spaceport, Urban}
\textbf{ST} (C) \textbf{EN} (C) \textbf{DX} (C) \textbf{AY} (B) \textbf{IN} (C) \textbf{MP} (B) \textbf{EM} (C) \textbf{LD} (C) \textbf{AG} (D) 
\textbf{SS:} Oppressed minority \\
\textbf{Skills:} Urban \textbf{1}; Grav \textbf{LT-1}; TempIML; Environ \textbf{3}; Streetwise \textbf{4}; Demolitions \textbf{3}; Handguns \textbf{3}; Blades \textbf{5}; Disguise \textbf{3}; Electro Tech \textbf{3}. \\
\textbf{Possessions:} Plastic explosives, timing device, pistol (Civ Level 6), knife, Electrokit. 
\begin{itemize}
\item \textbf{First Description:} The party is i nformed while in port that electrical repairs have become necessary on their vehicles due to new fire 
control regulations. A repairman shows up to effect the repairs. 
\item \textbf{GM's Description:} The man is an agent of an organization dedicated to random violence. Unless the characters take some 
precautionary measures, a timed explosive device will be implanted in either their ship or vehicle, whichever is applicable. The agent, 
by the way, has mistaken the characters' ship or vehicle for someone else's. 
\end{itemize}
\end{npc}

\hrulefill

\begin{npc}{U}{Spy}{Urban}
\textbf{ST} (C) \textbf{EN} (D) \textbf{DX} (D) \textbf{AY} (C) \textbf{IN} (C) \textbf{MP} (C) \textbf{EM} (B) \textbf{LD} (D) \textbf{AG} (D) \textbf{SS:} Poor colonist family \\
\textbf{Skills:} Urban \textbf{2}; Grav \textbf{LT-2}; Temp \textbf{NL}; Environ \textbf{3}; Demolitions \textbf{3}; Longarms \textbf{3}; Handguns \textbf{4}; Paint Gun \textbf{7}; Linguistics \textbf{2}; Disguise \textbf{5}; 
Weapon Tech \textbf{2}. \\
\textbf{Possessions:} Needle pistol, plastic pistol, dagger, false ID. 
\begin{itemize}
\item \textbf{First Description:} A man in the same place as the characters is shot by an unknown assailant. 
\item \textbf{GM's Description:} If the characters help this man, they will eventually discover he is a federal spy operating to uncover a black 
market in computer parts which is undercutting the federally-sponsored corporations. The headquarters is on this world. The spy may 
die and leave the characters with clues and information; he may live and ask the party for some sort of aid in his task. 
\end{itemize}
\end{npc}

\hrulefill

\begin{npc}{U}{Local Doctor}{Town}
\textbf{ST} (B) \textbf{EN} (B) \textbf{DX} (C) \textbf{AY} (B) \textbf{IN} (C) \textbf{MP} (A) \textbf{EM} (B) \textbf{LD} (B) \textbf{AG} (C) \textbf{SS:} Skilled tech family \\
\textbf{Skills:} Urban \textbf{1}; Grav \textbf{LT-1}; Temp \textbf{HO};Environ \textbf{2}; Teaching \textbf{1}; Diagnosis \textbf{3}; Treatment \textbf{2}; Handguns \textbf{1}; Chemistry \textbf{1}. \\
\textbf{Possessions:} Medical scanner (Civ Level 8), Chemanalysis lab (Civ Level 6), Chemsynthesis lab (Civ Level 7), plus various laboratory 
items such as flasks, microscope, etc. 
\begin{itemize}
\item \textbf{First Description:} As the party enters a small settlement out in the wilderness, they are introduced to the local doctor who, they are 
told, is also the mayor. 
\item \textbf{GM's Description:} This individual has discovered rare trace elements in the drinking water which, when augmented with additives of 
his own creation, cause the local populace to be very docile. He is in semi-retirement and wishes nothing more than to remain here 
with the adulation of this town. He is not dangerous and will give up his dream if he is threatened. 
\end{itemize}
\end{npc}

\hrulefill

\begin{npc}{U}{Art Dealer}{Urban, Suburban}
\textbf{ST} (A) \textbf{EN} (A) \textbf{DX} (A) \textbf{AY} (B) \textbf{IN} (A) \textbf{MP} (B) \textbf{EM} (B) \textbf{LD} (A) \textbf{AG} (A) \textbf{SS:} Family dominates planet \\
\textbf{Skills:} Urban \textbf{1};Grav \textbf{LT-1}; Temp \textbf{NL}; Environ \textbf{1};Diplomacy \textbf{1}; Trading \textbf{1}; Forgery/ Counterfeiting \textbf{1}. \\
\textbf{Possessions:} Outside of his shop and clothing, none. 
\begin{itemize}
\item \textbf{First Description:} The party sees an interesting work of art in the window as they walk into an art shop. 
\item \textbf{GM's Description:} The work of art has been created to cover a masterpiece which was thought lost ages ago. The exact type of 
piece is left up to the GM. If the players are enticed to purchase the item, examination with a Chemlab will give information as to the 
existence of the hidden work. The storeowner will be very affable and will not push the item, which he believes to be a worthless piece 
of junk. 
\end{itemize}
\end{npc}

\hrulefill

\begin{npc}{U}{Slaver}{Farmland}
\textbf{ST} (B) \textbf{EN} (C) \textbf{DX} (B) \textbf{AY} (B) \textbf{IN} (C) \textbf{MP} (A) \textbf{EM} (A) \textbf{LD} (C) \textbf{AG} (D) \textbf{SS:} Lower middle class family \\
\textbf{Skills:} Urban \textbf{0}; Grav \textbf{NW-1};Temp \textbf{NL}; Environ \textbf{3}; Agriculture \textbf{6}; Animal Drawn \textbf{1}; Tractor \textbf{1}; Truck \textbf{1}; Trading \textbf{4}; Longarms \textbf{2}. \\
\textbf{Possessions:} A hydroponic farm, farm equipment, rifle, slaves. 
\begin{itemize}
\item \textbf{First Description:} The party happens upon a huge hydroponic farm with many workers, and their owner approaches them. 
\item \textbf{GM's Description:} This farm owner is also a slaver, the workers all being owned by him. On this planet there is no law preventing 
humans from owning others, and he has made such an arrangement economically feasible. He will hide the fact that he owns the 
workers for as long as possible and, if found out,, will evict the characters from his land. He will fight if provoked. 
\end{itemize}
\end{npc}

\hrulefill

\begin{npc}{U}{Artist}{Hills, Mountains}
\textbf{ST} (A) \textbf{EN} (B) \textbf{DX} (D) \textbf{AY} (C) \textbf{IN} (C) \textbf{MP} (C) \textbf{EM} (C) \textbf{LD} (B) \textbf{AG} (B) \textbf{SS:} Oppressed minority \\
\textbf{Skills:} Urban \textbf{0}; Grav \textbf{LT-1}; Temp \textbf{NL}; Environ \textbf{2}; Glider \textbf{1}; Teaching \textbf{2}; Survival \textbf{2}. \\
\textbf{Possessions:} Foodstuffs and sculpting tools. 
\begin{itemize}
\item \textbf{First Description:} As the characters round a bend on a backwoods road, they see an incredible edifice; it reflects light in all directions and seems unworldly. 
\item \textbf{GM's Description:} This is a piece of sculpture created by a hermit artist who lives in a hut nearby. The artist retreated away from 
society to create this masterpiece; even to the untrained eye it is beautiful. It is so large that moving it would be impossible without 
heavy equipment. Thus, there is little the party can do to hurt it. If the characters take hostile action, the artist will hide, confident his 
work will remain after the party has left. 
\end{itemize}
\end{npc}

\hrulefill

\begin{npc}{U}{Disease Carrier}{Jungle, Forest, Woods}
\textbf{ST} (B) \textbf{EN} (B) \textbf{DX} (B) \textbf{AY} (B) \textbf{IN} (B) \textbf{MP} (B) \textbf{EM} (C) \textbf{LD} (B) \textbf{AG} (B) \textbf{SS:} Poor colonist family \\
\textbf{Skills:} Urban \textbf{0}; Grav \textbf{NW-1};Temp \textbf{CD}; Environ \textbf{3}; Agriculture \textbf{4}. \\
\textbf{Possessions:} Farm tools. 
\begin{itemize}
\item \textbf{First Description:} As the party forages through the brush, they come upon a small farm growing wild exotic tubers. 
\item \textbf{GM's Description:} The farmer and his family all are carriers for a dread disease (the exact type is left up to the GM). They are immune because they eat the tubers daily. If the characters take some of the food (which the farmer will offer), they will avoid getting ill 
until the supply runs out. 
\end{itemize}
\end{npc}

\hrulefill

\section{Spaceship Encounters}
\label{sec:spaceship-encounters}
% [49.0]

A spaceship encounter may occur when using the \emph{Deep Space} or
\emph{Planet Space} column of the Encounter Table. The following
encounter descriptions indicate whether they occur in deep or planet
space and are divided into common, rare, and unique categories.

Remember, when the party is travelling in a spaceship, an encounter
check is conducted \emph{once per day} (24-hour period).


\subsection{Common Spaceship Encounters}
\label{sec:comm-spac-enco}
% [49.1]

\begin{itemize}
\item \textbf{Deep Space.} While heading out of the system, the
  characters are hailed by a trading vessel heading into the system
  with some refined ore for the settlements.
  
  The vessel is a Corco \emph{Mu}, and the voice on the radio is
  friendly.  The GM should decide where this vessel is from and what
  world it is heading to.  It poses no threat to the party.
\item \textbf{Planetary Space.} The party's vessel is approached by a
  \emph{Dagger} bearing federal ID and hailing the party to stop.
  
  This is a federal ship calling for routine inspection of the goods
  and material the ship is carrying.  They will inspect closely and
  will hand out fines and arrest guilty parties if such items are
  found.
\item \textbf{Planetary Space.} While in orbit, the party finds that a
  \emph{Piccolo} has gained rapidly on them and hails them to slow up
  and prepare to be boarded.
  
  This is the Astroguard needing to ascertain where the party's vessel
  came from, where it is going, and with what purpose.
\item \textbf{Deep Space.} The party's ship passes a \emph{Corco
    Gamma} heading out of the system.
  
  This is a Biological Survey Mission heading for a new system.  The
  presence of habitable worlds has just been confirmed and this is the
  first survey mission to go there to collect specimens.  There is a
  suspicion of danger, and the ship has brought along some weaponry to
  defend itself.  The crew is hopeful no danger will be present.
\item \textbf{Planetary Space.} At first radar indicates and then
  visual sightings confirm a \emph{Terwillicker 5000} is rapidly
  tumbling end over end toward the party's vessel.
  
  A collision is imminent unless the pilot can avoid the accident.
  Make a check for collision, using the accident procedure (see GM
  guide \ref{GM-sec:accidents}).  The rationale for this
  out-of-control spacecraft is left up to the GM.
\end{itemize}


\subsection{Rare Spaceship Encounters}
\label{sec:rare-spac-enco}
% [49.2]

\begin{itemize}
\item \textbf{Deep Space.} As the party heads into a system from the
  jump point, they notice a Corco \emph{Zeta} heading toward them at a
  brisk clip.  No radio contact can be made, and there is no
  possibility of collision.
  
  The ship has been disabled by a meteor strike and cannot hail the
  party's ship nor maneuver in any fashion.  If the party does not
  attempt to change course and catch them (a maneuver the GM should
  play out), they will drift helplessly into interstellar space.  If
  it appears the party will not stop to help, the \emph{Zeta} will
  attempt to gain their attention by firing their burster.
\item \textbf{Deep Space.} Near the jump point into the system, the
  party encounters a Corco \emph{Gamma} drifting slowly.
  
  If the party attempts to raise the ship by radio, they will receive
  a brusque reply indicating that all is well.  Actually, on board
  that vessel an argument is taking place between the navigator and
  the captain concerning the fact they have jumped to a system they
  did not wish to.  The psion on the \emph{Gamma} will attempt to
  contact the Psion on the party's vessel to ask for help.
\item \textbf{Planetary Space.} The party's vessel passes a
  \emph{Clarinet}.
  
  This is a scientific mission carrying supplies to a colony on a
  far-off star.  The colony is new and needs all the help it can get.
  The \emph{Clarinet's} communications officer asks if any on the
  party's ship are interested.
\end{itemize}


\subsection{Unique Spaceship Encounters}
\label{sec:uniq-spac-enco}
% [49.3]

\begin{itemize}
\item \textbf{Deep Space.} A \emph{Piccolo}, obviously damaged, is
  drifting out near the jump point.
  
  The crew of 10 are criminals and wish to hijack the party's vessel.
  Their maneuver power is gone, but they can still fight.  They will
  feign innocence and will try to get on board the other vessel armed
  and try for a take over.  They jumped into this system escaping from
  authorities.
\item \textbf{Planetary Space.} A ship of totally unknown design is in
  a slow elliptical orbit around a moon.
  
  This ship is either (a) and old space probe or (b) a wrecked alien
  craft.  The GM should decide for himself, as he should also decide
  why it is derelict.  This encounter may need a great deal of time to
  play out.
\end{itemize}

\section{Accidents}
\label{sec:accidents}
% [50.0]

The accidents listed fall into two categories: those the characters
can prevent and those they cannot. The GM may choose from either
category. The accidents they can prevent will involve a check, which
determines whether their skill has avoided the unfortunate occurrence.
When any incident occurs in which equipment becomes damaged, the
characters' tech skills may be called upon to effect repairs.

The procedure for affecting the repair of an item is detailed in GM
guide \ref{GM-sec:repair}.  See GM guide \ref{GM-sec:accidents} for
additional information on accidents.

The base chance to avoid a common \textbf{(C)} accident is 75\%; a
rare \textbf{(R)} accident, 50\%; and a unique \textbf{(U)} accident,
25\%. Each of the following accident types can occur in any \emph{one}
of the three possible degrees indicated (\textbf{C}, \textbf{R}, or
\textbf{U}).


\subsection{Avoidable Accidents}
\label{sec:avoidable-accidents}
% [50.1]


\subsubsection*{Vehicle Accidents}
\label{sec:vehicle-accidents}

\begin{description}
\item[(C)] A rut in the road causes possible loss of control.
\item[(R)] An underground stream has caused a sinkhole to appear just
  in front of the vehicle; try to avoid getting stuck.
\item[(U)] A landslide threatens to bury the vehicle.
\item[(C)] While travelling near a marsh, the ground begins to sink
  away; try to avoid being mired.
\item[(R)] A sudden seismic tremor causes possible loss of control.
\item[(U)] A sudden increase in volcanic activity causes magma to
  spurt toward the vehicle; try to avoid it.
\item[(C)] While travelling along a cliff, a wheel slips toward the
  edge; try to avoid becoming stuck.
\item[(R)] Brakes begin to act sporadically while rolling down a hill;
  try to avoid possible accident.
\item[(U)] Earthquake causes large fissure to form in road; try to
  avoid falling in.
\item[(C)] In a city, another vehicle ignores a traffic signal and
  heads directly toward your vehicle; try to avoid it.
\item[(R)] While driving by a construction site, a beam has broken
  loose; try to avoid it falling on you.
\item[(U)] A crazed pedestrian steps right out in front of your
  vehicle; try to avoid running him over (an NPC encounter may result
  if you succeed).
\end{description}

\subsubsection*{Suit/Armor Accidents}
\label{sec:suitarmor-accidents}


\begin{description}
\item[(C)] If character's Gravity Skill Level is less than \textbf{1}
  a fall is possible.
\item[(R)] The character has wedged his boot into a crevice; try to
  extricate it without causing damage.
\item[(U)] A violent windstorm has sprung up carrying with it debris
  and small rocks; try to reach cover without being pummelled.
\item[(C)] A boulder comes rolling down a hill; try to avoid being
  struck.
\item[(R)] Ice cracks open in front of you; try to avoid falling in.
\item[(U)] Jet-pack is buffeted by winds; try to land safely.
\end{description}

\subsubsection*{Air Vehicle Accidents}
\label{sec:air-vehicle-accid}


\begin{description}
\item[(C)] Downdraft threatens to pull craft off course; try to regain
  control.
\item[(R)] Electrical storm causes instrument loss; try to land
  safely.
\item[(U)] Violent hailstorm forces an emergency landing.
\end{description}


\subsubsection*{Marine Vehicle Accidents}
\label{sec:marine-vehicle-accid}


\begin{description}
\item[(C)] Large swells cause difficulty in reaching land; try to
  return safely.
\item[(R)] Violent storm threatens to swamp vessel.
\item[(U)] Huge creature collides with vessel and possibly capsizes
  it.
\end{description}


\subsubsection*{Outer Space Accidents}
\label{sec:outer-space-accid}

\begin{description}
\item[(C)] As you try to land your shuttlecraft, unexpected terrain
  threatens to cause landing problems.
\item[(R)] As your ship comes out of hyperjump, you encounter a meteor
  storm; try to avoid collision.
\item[(U)] Flying through the tail of a comet causes instrument
  malfunction; the ship must be flown manually.
\item[(C)] Near the spacecraft, a shuttle loses control and heads for
  your craft; try to avoid collision.
\item[(R)] Some unexplained disturbance causes your orbit to decay;
  try to avoid falling into atmosphere before gaining control.
\item[(U)] Shuttle landing gear fails; try to avoid crash.
\end{description}


\subsection{Unavoidable Accidents}
\label{sec:unav-accid}
% [50.2] 


\subsubsection*{Common}
\label{sec:common}

\begin{itemize}
\item If in a poison or corrosive atmosphere, a scanner will stop
  giving correct readings.
\item If in a thin atmosphere, a Chemsynthesizer will stop cold.
\item If in any star system other than an "M" spectral class, star
  flares cause photographic equipment to malfunction.
\item If in a contaminated atmosphere, any radio will not work.
\item If in a vacuum, any holographic equipment will not work.
\item A violent storm brews up and any exposed equipment is damaged.
\item A mild seismic quake topples any exposed equipment.
\item If in a temperature zone above 100\textdegree, any cooling
  system will malfunction.
\item If in a corrosive atmosphere, any air system will shut down.
\item The ignition for any ship drive system fails.
\item Computer freezes up.
\end{itemize}

\subsubsection*{Rare}
\label{sec:rare}

\begin{itemize}
\item Any character in a temperature type (CD or HO) 2 removed from
  his home type contracts a disease.
\item A captured creature destroys its cage and attempts to escape.
\item A robot stop following commands.
\item A meteor strikes ship causing damage.
\item In a stressful situation, any force field overloads.
\end{itemize}

\subsubsection*{Unique}
\label{sec:unique}

\begin{itemize}
\item If in a corrosive atmosphere, any equipment exposed will
  deteriorate.
\item Any hand-held energy weapon explodes.
\item Any grenade explodes.
\item Any spaceship pod begins to leak.
\item Any computer will not allow access to information.
\item Any Civ Level 8 robot attacks its owner.
\end{itemize}

As with the other lists, this one is intended to guide and inspire the
GM, not limit him. Feel free to invent your own accidents.


%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "adventure_guide"
%%% End: 
