%%
%% $Id: ag_intro.tex,v 1.1 2004/10/07 18:26:39 goltz20707 Exp $
%%
\section*{Introduction}
\label{sec:introduction}

The \emph{Universe} \emph{Adventure Guide} is divided into two
distinct parts.

\emph{Encounters} is used by the GM as a continuing source of
creatures, non-player characters, spaceship encounters, and accidents
that may be used from adventure to adventure. The players should not
be allowed to see Chapter \ref{cha:encounters}, Encounters, at all.
Although they will eventually run out, the encounters presented will
continue to be a source of inspiration and ideas for the imaginative
GM.

\emph{Lost on Laidley} is a single self-contained adventure, intended
to be played once by the GM and the players during their first session
of \emph{Universe}. The Chapter presents a completely generated star
system that serves three purposes. It is the setting for \emph{Lost on
  Laidley}; it is a comprehensive example of world generation; and it
may be the backdrop for adventures created by the GM. The players
should be allowed to see only Sections \ref{sec:orionis-system} and
\ref{sec:players-characters} of the adventure. The contents of the
remaining Sections will be revealed to them during play.  After the
adventure has been played, the format of Lost on Laidley and the ideas
presented within it will continue to guide the GM in the conduct of
his own adventures.

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "adventure_guide"
%%% End: 
