%%
%% $Id: changes.tex,v 1.3 2004/10/27 20:03:50 goltz20707 Exp $
%%
\chapter{Changes from Original}
\label{cha:changes-from-original}

This book is based on \emph{Universe} Second Edition, published by
SPI.  Some changes have been made for formatting and grammatical
reasons.  Other changes have been made for the following reasons:


\begin{description}
\item[Section \ref{sec:universe-future-change-tran}:] The definition
  of the Tran had read ``This was the equal of 500 1980 dollars.''
  This was changed to account for inflation and for grammatical
  reasons.
\item[Section \ref{sec:universe-future-change-swahili}:] Afrikaans was
  dropped as a language and Swahili substituted. Afrikaans is only
  spoken in South Africa today, whereas Swahili is a \emph{lingua
    franca} spoken in numerous countries in Africa and the Arabian
  peninsula.
%  (It might be argued that South Africa is more likely to
%   found a colony than other African nations, but I don't accept that
%   argument.) 
\item[Section \ref{sec:contents}:] The section listing game
  components had read:
  
  \begin{quote}
    {\normalfont\bfseries [\ref{sec:contents}] A complete copy
      of \emph{Universe} should include:}
    
    \begin{itemize}
    \item 1 22'' x 33'' Interstellar Display
    \item 1 Gamemaster's Guide Book
    \item 1 Adventure Guide Book
    \end{itemize}
    
    The following parts are included in the boxed version only:
    
    \begin{itemize}
    \item One 17'' x 22'' Tactical Space Combat game map
    \item One \emph{Delta Vee} rules booklet (tactical space combat)
    \item One 200 die-cut \emph{Delta Vee} counter sheet
    \item Two 20-Sided dice
    \item One Counter tray
    \item One Game Box
    \end{itemize}
  \end{quote}

  Since this version of \emph{Universe} exists in electronic form
  only, the list of contents was changed to reflect this fact.

\item[Section \ref{sec:player-uses-habitat-change-temp}:] The
  temperature range descriptions were edited to add equivalents in
  degrees Celsius and Kelvins.

\item[Section \ref{sec:play-records-inform-change-display} and Figure
  \ref{fig:example-environ-disp}:] The figure was edited to include
  examples of how to fill out the Gravity Skill Level boxes and the
  Temperature Range box.

\item[Section \ref{sec:prof-doctor-bot}:] The original text said
  benefit F for doctor included a ``Manner 36sdf robot with medical
  system.''  However, there is no such robot chassis.  Since a medical
  system takes 1 Hardware and 2 Software points, and a Manner
  \emph{38sdf} has exactly this many, the bot chassis type was changed
  to ``Manner 38sdf''.

\item[Section \ref{sec:scientific-skills-change-ten}:] The word
  ``ten'' was changed to the boldface numeral \textbf{10} to agree in
  appearance with other skill descriptions.

\item[Table \ref{tab:robot-chassis}:] Two columns were added by Ian
  Taylor:
  \begin{itemize}
  \item \textbf{Availability} is the estimated restrictions of
    purchase and the requirement of permits based on the robot's
    potential danger to the public in strength and armor.
  \item \textbf{Base Repair Time (Hours)} is the base time to repair a
  damaged robot chassis of this type.
  \end{itemize}

  A description of the weight and availability columns was also added.

\item[Section \ref{sec:robot-systems-change-1}:] The descriptions of
  robot systems had availability and Civ Level added for ease of use.
  (Ian Taylor)
\item[Table \ref{sec:robot-systems-change-2}:] This table was created
  for ease of use.  It summarizes the information in section
  \ref{sec:robot-systems-change-1}. 

\item[Section \ref{sec:marine-vehicles-hyphens-1}:] Hyphens added for
  grammatical correctness.

\item[Section \ref{sec:personal-equipment-split-tables}:] Equipment
chart split into scientific, repair, communications and miscellaneous
equipment.  (Ian Taylor)
\item[Section \ref{sec:law-level} and Table
  \ref{tab:world-development}:] Two columns were added by Ian Taylor:
  \begin{description}
  \item[Patrol Rating:] The frequency (out of 10) of Astroguard
    patrols, the sum of the Spaceport Class and Law Level.
  \item[Security Rating:] The probability (out of 10) of enforcement
    agencies discovering illegal commodities once a spacecraft has
    landed, the sum of the Spaceport Class and Law Level found above,
    plus \textbf{2}.
  \end{description}

  These were adapted from SPI's \emph{Star Trader} game.

\item[Section \ref{sec:coast-1g-accel}:] The text on ``Coasting 1G
  Acceleration'' was added, from \emph{Ares} Magazine issue \#17.


\item[Section \ref{sec:skill-perception}:] The ``Perception'' skill
  was added from \emph{Ares} Magazine (issue unknown).
  
% \item[Section \ref{sec:char-may-incr-change-zero}:] The rule regarding
%   experience points for using perception skill was changed to award
%   one EP on rolling a \textbf{0} on either die instead of a
%   \textbf{1}, to be more in agreeement with other skills' EP rules.

\item[Table \ref{tab:psibacklash-change-ref}:] Added a reference to
  the Hyperjump Table, since that table was mentioned.

\item[Table \ref{tab:weapon-change-table}:] Table was rearranged
  slightly, and a column added (Availability), by Ian Taylor.

\item[Section \ref{sec:air-vehicles-change-largest}:] Added a note to
  the effect that the shuttle listed in 
  the Air Vehicles Table is the largest of its kind. 

\item[Section \ref{sec:add-two-energy}:] Added language from 2nd
  Edition \emph{Universe} about the fact that non-streamlined ships
  can carry two Energy pods in one pod berth.
\end{description}

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "gm_guide"
%%% End: 
