%%
%% $Id: charactergen.tex,v 1.3 2004/10/27 20:04:37 goltz20707 Exp $
%%
\chapter{\index{character!generation|(}Character Generation}
\label{sec:character-generation}
\newcounter{myprofs}

Each player of Universe assumes the role of a character. The player
develops the distinct personality and abilities of his character
during adventures, as the character interacts with the other players
characters and the various entities controlled by the Gamemaster.

The player generates a character by using the rules in Sections
\ref{sec:character-heritage}, \ref{sec:char-devel}, and
\ref{sec:char-prof}. By rolling dice and. making choices in the order
laid out in these Sections, the character is developed into an adult
who is a result of the luck of his parentage, the pressures of his
environment, and the decisions he has made throughout adolescence and
young adulthood. A completely generated character is defined by a
series of numerical \emph{Characteristic Ratings} and \emph{Skill
  Levels} that represent the character's strengths and weaknesses in
all situations that might occur while playing adventures.

Each player keeps track of his character's attributes and acquisitions
during character generation and throughout his adventuring life on a
\emph{Character Record}. The GM should provide each player with a
photocopy of the Character Record included with this booklet. The
players should use pencil to fill in their Character Record, as much
of the information they record will be altered during play. At some
point during character generation, the player must select a name and a
sex for his character. These decisions are entirely up to the player,
but it is suggested that a character's sex be the same as that of the
player. The GM should supervise character generation and all record
keeping on each Character Record, to make sure the current status of a
character is noted correctly. If the GM wishes, he may make a second
copy of each Character Record to verify all character information
and/or to note things about a character that he does not want the
player to know (such as a price on his head or an undetectable alien
disease).


\subsection*{Terms Used in Character Generation}
\label{sec:terms-used-character}



\begin{description}
\item[Characteristic.] A numerical \emph{rating}, ranging from
  \textbf{1} to \textbf{12}, which describes a facet of a character's
  physical, mental, or emotional makeup. Characteristic Ratings are
  determined during character generation and may not be increased once
  play begins. The nine characteristics each character receives are:

  \begin{description}
  \item[\index{strength|bb}\index{characteristics!strength|see{strength}}Strength.] The brute force a character can exert using the
    muscles of his arms, legs, and torso. Strength determines how much
    a character can carry, how far he can throw an object, and the
    force with which he may strike.
  \item[Endurance.] The amount of punishment the character's body can
    absorb before he loses consciousness or dies. Endurance determines
    the length of time a character may keep up an exertive action
    (such as sprinting) and the amount of time required for the
    character to heal from wounds or diseases.
  \item[\index{characteristics!dexterity|see{dexterity}}Dexterity.] The character's digital control and eye-hand
    coordination. Dexterity affects the character's use of any
    hand-held weapon and the efficiency with which he may perform
    complex tasks with his hands.
  \item[\index{characteristics!agility|see{agility}}Agility.] The quickness with which the character uses his
    body. Agility determines the speed with which a character can move
    and his  ability to dodge and maneuver in close quarters.
    
    \textbf{Note:} The preceding four characteristics are collectively
    referred to as the \emph{Physical Characteristics}.
  \item[\index{characteristics!intelligence|see{intelligence}}Intelligence.] The character's intuitive and acquired
    knowledge. Intelligence affects the character's ability to note
    peculiarities and to deduce information in a situation.
  \item[\index{characteristics!mental power|see{mental power}}Mental Power.] The character's control of his mind and
    emotions, especially in stress situations.  A character with high
    mental power (\textbf{4}, \textbf{5} or \textbf{6}) may be
    psionic.  \textbf{Note:} The mental power Characteristic Rating
    ranges from \textbf{0} to \textbf{6}.
  \item[\index{characteristics!leadership|see{leadership}}Leadership.] The character's ability to lead others and to
    control a situation. Leadership determines whether or not a party
    of characters have the initiative during combat (see
    \ref{sec:initiative}).
  \item[\index{characteristics!empathy|see{empathy}}Empathy.] The character's ability to interact with humans and
    other intelligent and semi-intelligent beings. Empathy affects the
    chances of communication with alien life forms and may allow a
    character to perceive another human's unspoken intentions.
  \item[\index{characteristics!aggression|see{aggression}}Aggression.] The base instincts a character exhibits,
    especially in a stress situation. A character with a low
    Aggression Rating will often behave in a cowardly or
    self-preserving manner, while a character with a high Aggression
    Rating will often take reckless actions in combat. A character's
    aggression may be controlled by his mental power.
  \end{description}
  
\item[\index{field of study}Field of Study.] One of seven areas of schooling that a
  character undertakes in his youth, including theoretical science,
  applied science, business, the humanities, the mind, the body, and
  the military. In addition, each character's basic education is
  represented by the \emph{general} field of study. The fields of
  study a character undertakes affect his characteristics, the
  professions he may choose, and the skills he may acquire.
\item[\index{habitat|bb}Habitat.] The environment in which the character was raised.
  Each character's habitat consists of a home environ (expressed in
  terms of the contour of the land and its major features), a gravity
  type, a temperature range and, in some cases, an urban background.
\item[\index{potential|bb}Potential.] A numerical rating, expressed as a multiplier from
  \textbf{$\frac12$} to \textbf{4}, representing the strengths and
  weaknesses exhibited by the character in his childhood. There are
  four potentials: physique, coordination, intellect, and social
  background. Once character generation is completed, potentials have
  no effect on the character.
\item[\index{skill}Skill.] A specific area of expertise that may be acquired by a
  character. During character generation, each character receives
  \emph{Initial} Skill Points, with which to acquire skills before
  entering a profession, and normal Skill Points, with which to
  acquire skills while practicing his profession. The proficiency that
  a character has with a given skill is represented by a numerical
  \emph{level} ranging from \textbf{1} up to a maximum level that
  varies from one skill to another and is sometimes limited by one of
  his Characteristic Ratings.  \textbf{Exception:} An Environ Skill
  Level or a Gravity Skill Level may be a negative number or
  \textbf{0}.  Unlike characteristics, skills may be acquired and
  improved during play. The nature and use of all skills are detailed
  in Chapter \ref{cha:skills}.  Three skills are introduced in the
  early part of character generation:

  \begin{description}
  \item[Environ Skill.] The character's knowledge and acclimation in a
    specific environ. Each character receives Skill Levels in all 33
    environs listed on the Environ Skill Display of the Character
    Record during character generation. The single highest Environ
    Skill Level received is his Skill Level in his home environ.
  \item[Gravity Skill.] The character's ability to function in a
    specific gravity type. Each character receives a Skill Level in
    all four gravity types on the Character Record during character
    generation. The single highest Gravity Skill Level received is his
    Skill Level in the gravity type of his home planet.
  \item[Urban Skill.] The character's familiarity with the ways of
    city life and the upper classes. Certain characters receive an
    Urban Skill Level during character generation; those that do not
    are considered to come from essentially rural areas.
  \end{description}
  
\item[Study Point.] A representation of a commitment by the character
  to a certain field of study. Each character receives \textbf{2} to
  \textbf{6} Study Points, which he expends to enter any of the seven
  fields of study.
\end{description}




\subsection*{\index{character!generation sequence}Character Generation Sequence}
\label{sec:char-gener-sequ}



To generate a character, the player should conduct the following steps
in order. The specific procedures mentioned in this outline are
explained in detail in Sections \ref{sec:character-heritage},
\ref{sec:char-devel}, and \ref{sec:char-prof}.

\renewcommand{\theenumi}{\Alph{enumi}}
\renewcommand{\theenumii}{\arabic{enumii}}
\begin{enumerate}
\item \textsc{character heritage}
  \begin{enumerate}
  \item \textbf{Determine 4 Potential Multipliers.}  Each character
    begins as a series of four numbers, representing his potential in
    the areas of physique, coordination, intellect, and social
    background.
  \item \textbf{Calculate the number of Study Points the character
      receives.} A character receives from \textbf{2} to \textbf{6}
    Study Points, depending on his potentials.
  \item \textbf{Determine the character's natural habitat.} This
    includes the character's home environ, his Skill Levels in all
    environs, his Skill Levels in various gravities, his comfortable
    temperature range, and his Skill Level in urban areas. Certain
    characters may come from a deep space environ.
  \item \textbf{Determine the character's social standing.} Each
    character receives a particular family history and, initial
    wealth. Each character receives from \textbf{1} to \textbf{5}
    Initial Skill Points, depending on his social standing.
  \end{enumerate}
\item \textsc{character development}
  \begin{enumerate}
    \setcounter{enumii}{4}
  \item \textbf{Choose fields of study for the character.} Study
    Points are expended to choose any of seven fields of study.
  \item \textbf{Choose initial skills for the character.} Initial
    Skill Points are expended to choose a variety of basic skills.
    Skills that may be chosen depend on the character's fields of
    study.
  \item \textbf{Determine the character's 9 Characteristic Ratings.}
    The character's Potential Multipliers and fields of study affect
    the determination of his characteristics.
  \end{enumerate}
\item \textsc{character professions}
  \begin{enumerate}
    \setcounter{enumii}{7}
  \item \textbf{Choose a profession for the character.} Any one of 
%    23
    \ref{item:myprofs}
    professions is available, depending on the character's fields of
    study and Characteristic Ratings.
  \item \textbf{Declare how many years the character will practice his
      profession.} A character may not be able to spend as many years
    at his profession as the player declares; he may be discharged or
    without work for part of the declared time.
  \item \textbf{Determine the effects of age on the character.} His
    physical characteristics (strength, endurance, dexterity, and
    agility) may be reduced if the character is beyond the age of 24.
  \item \textbf{Calculate the number of Skill Points the character
      receives.}  The character's profession, his Intelligence Rating,
    and the number of years he spent at it affect the number of Skill
    Points received.
  \item \textbf{Choose skills for the character.} Skill Points are
    expended to choose and improve a variety of skills related to the
    character's profession and his fields of study. Mark an \textbf{\textsf{X}}
    next to the skills for which the character is eligible that were
    not chosen for him.
  \item \textbf{Determine benefits the character receives from his
      profession.}  Weapons, equipment, money and/or prestige are avail
    able to a character, depending on how long he spent at his
    profession.
  \end{enumerate}
\end{enumerate}
\renewcommand{\theenumi}{\arabic{enumi}}
\renewcommand{\theenumii}{(\alph{enumii})}


\section{\index{character!heritage}Character Heritage}
\label{sec:character-heritage}

The early years of a character's life are very important in forming
the basis of his physical and mental makeup. The well being of his
predecessors determines the opportunities the character will be able
to exploit. Each character receives a series of potential values
representing an early assessment of high attributes, skills in the
habitat in which he was raised, and a financial endowment base on his
family's social standing.



\subsection[Potentials]{The player begins generating his character by
  determining the character's \index{potential}potential in the areas of physique,
  coordination, intellect, and social background.}
\label{sec:potentials}

For each of these four, the player roll one die and locates the result
on the \textbf{Character Heritage Table} (table \vref{tab:heritage})
to find the character's \emph{Potential Multiplier} for that area.

The player must roll for each Potential Multiplier in the order listed
above. As each potential is determined, its Multiplier is recorded on
the Character Record.

After determining all four potentials, the player \emph{adds} the
multipliers together to find how many \emph{Study Points} his
character receives, in accordance with the \emph{Character Heritage
  Table} (\vref{tab:heritage}).

Note the number of Study Points received on the Character Record.

\begin{table}[htbp]
  \centering
  \fbox{%
    \begin{minipage}{3.3in}
      \centering
      \caption{Character Heritage}
      \label{tab:heritage}

      \medskip

      \begin{tabular}{cllc}
        Die & \parbox{2.5cm}{\centering Potential\\Multiplier} &
          \parbox{1.5cm}{\centering Sum of\\
          Potential\\Multiplers} & \parbox{1.5cm}{\centering Study Points\\Received}\\
        \rowcolor{grey}
        1 & $\frac12$ (Poor) & 4 or fewer & 6\\
        2--3 & 1 (Below Average) & $4\frac12$ to $6\frac12$ & 5\\
        \rowcolor{grey}
        4--6 & 2 (Average) & 7 to $9\frac12$ & 4\\
        7--9 & 3 (Above Average) & 10 to $11\frac12$ & 3\\
        \rowcolor{grey}
        10 & 4 (Exceptional) & 12 or more & 2\\
      \end{tabular}

      \medskip

      \parbox{0.9\textwidth}{To determine potential
        multiplier for physique, 
        coordination, intellect, and social background, roll one die and
        locate the result under ``Potential Multiplier''. After all four
        potentials are determined; add all four multipliers together to
        determine how many Study Points the character receives under the
        ``Study Points Received'' column.}
    \end{minipage}}
  \end{table}

\subsection[Habitat Table]{The player uses the  \index{habitat}Habitat Table to
  determine the 
  natural environment in which his character grew up and in which he
  will be most effective during play.}
\label{sec:habitat-table}

The player rolls one die to determine the column of the Table used and
rolls a second die to determine the entry referred to in that column.
The second die result is modified by \emph{adding} the character's
\emph{Physique Multiplier} and subtracting his \emph{Coordination
  Multiplier}. A multiplier of \textbf{$\frac12$} is considered
\textbf{0} for this purpose. Each entry on the Home Environ Table
contains the following information:

\begin{itemize}
\item \textbf{The specific type of environ from which the character
    comes.}  Each environ type is stated as a two-part abbreviation,
  explained on the Table. The first part represents the contour of the
  land (flat, hills, mountains, or peaks). The second part represents
  the major features of the land (volcanic, craters, barren, light
  vegetation, woods, forest, jungle, marsh, or ice). Exceptions to
  this two-part system include the three \emph{water} environs (inland
  waterways, sea surface, or sea submerged). Certain characters may be
  able to choose a \emph{deep-space} environ (see
  \ref{sec:deep-space}).
\item \textbf{The character's Home Environ Skill Level.} Each
  character receives a Skill Level ranging from \textbf{1} to
  \textbf{6} in his home environ. This level may be reduced if the
  player chooses to increase his character's \emph{Gravity} Skill
  Level (see \ref{sec:habitat-table-use}). A character's Skill Level
  in his home environ affects his Skill Level in other environs and
  may be improved during play of the game.
\item \textbf{The gravity on the character's home planet.} Each
  gravity type is stated as an abbreviation, explained on the Table.
  \emph{Near weightless gravity} represents from \textbf{0.0} to
  \textbf{0.4 G}, \emph{light gravity} represents from \textbf{0.7} to
  \textbf{1.0 G}, \emph{heavy gravity} represents from \textbf{1.3} to
  \textbf{1.7 G}, and \emph{extreme gravity} represents from
  \textbf{2.0} to \textbf{2.5 G}. Each character is considered to be at
  Skill Level 1 in his own gravity type and at lesser Skill Levels in
  other gravity types (unless his Gravity Skill Level is increased per
  \ref{sec:habitat-table-use}). Gravity Skills may also be improved
  during play.
\item \textbf{The temperature range in the character's home
    environ.}\label{sec:player-uses-habitat-change-temp}
  Each temperature range is stated as an abbreviation, explained on
  the Table. \emph{Cold} ranges from \textbf{-50} to \textbf{0}
  degrees Fahrenheit (\textbf{-45} to \textbf{-18} degrees Celsius,
  \textbf{228} to \textbf{255} Kelvins); \emph{normal} ranges from
  \textbf{25} to \textbf{100} degrees (\textbf{-4} to \textbf{38}
  degrees C, \textbf{269} to \textbf{311} K); and \emph{hot} ranges
  from \textbf{125} to \textbf{175} degrees (\textbf{52} to
  \textbf{79} degrees C, \textbf{325} to \textbf{353} K).  Characters
  receive no Skill Level in their temperature range.
\item \textbf{The character's Urban Skill Level.} Certain results give
  a character an Urban Skill Level ranging from \textbf{1} to
  \textbf{4}. If a result from the Table lists no Urban Skill Level,
  the character receives none. A character's Urban Skill may be
  improved during play.
\end{itemize}


\subsection{Habitat Table}
\label{sec:habitat-table-1}

See table \vref{tab:habitat}.

\newlength{\cellwidth}
\setlength{\cellwidth}{1.2cm}
\newlength{\extralinespace}
\setlength{\extralinespace}{3mm}
\begin{table}[htbp]
  \setlength{\tabcolsep}{0.25\tabcolsep}
%  \tiny
  \scriptsize
  \centering
  \fbox{%
    \begin{minipage}{0.9\textwidth}
      \centering
      \caption{Habitat Table}
      \label{tab:habitat}

      \medskip

      \begin{tabular}{c|cccccccccc}
    & \multicolumn{10}{c}{First Die Roll}\\[\extralinespace]
    \\[\extralinespace]
    \parbox{\cellwidth}{Second Die + Ph - Co} & 1 & 
    2 & 3 & 4 & 5 & 6 & 7 & 8 &
    9 & 10\\[\extralinespace] 
    \hline
    -3 & 
    \parbox{\cellwidth}{4:HL-VO\\NW-HO} & 
    \parbox{\cellwidth}{6:SF\\HY-NL} & 
    \parbox{\cellwidth}{4:FL-CR\\LT-NL} & 
    \parbox{\cellwidth}{4:HL-IC\\NW-NL/1} & 
    \parbox{\cellwidth}{2:HL-CR\\NW-NL/2} & 
    \parbox{\cellwidth}{2:HL-CR\\HY-NL/2} & 
    \parbox{\cellwidth}{3:FL-CR\\LT-NL/1} & 
    \parbox{\cellwidth}{2:HL-BN\\HY-CD/2} & 
    \parbox{\cellwidth}{4:FL-MA\\HY-NL/1} & 
    \parbox{\cellwidth}{3:HL-IC\\NW-NL/3}\\[\extralinespace]
    \rowcolor{grey}
    -2 & 
    \parbox{\cellwidth}{3:HL-WD\\LT-NL/1} & 
    \parbox{\cellwidth}{4:HL-LV\\NW-NL} & 
    \parbox{\cellwidth}{4:FL-CR\\NW-CD/1} & 
    \parbox{\cellwidth}{3:PL-BN\\LT-HO/1} & 
    \parbox{\cellwidth}{4:SF\\LT-NL/4} & 
    \parbox{\cellwidth}{2:HL-CR\\HY-NL/1} & 
    \parbox{\cellwidth}{3:HL-FT\\LT-NL/1} & 
    \parbox{\cellwidth}{4:MN-VO\\NW-HO/1} & 
    \parbox{\cellwidth}{3:HL-VO\\LT-HO/1} & 
    \parbox{\cellwidth}{2:HL-WD\\LT-HO}\\[\extralinespace]
    -1 & 
    \parbox{\cellwidth}{4:HL-JU\\EX-NL} & 
    \parbox{\cellwidth}{6:FL-LV\\NW-HO} & 
    \parbox{\cellwidth}{3:HL-WD\\LT-NL} & 
    \parbox{\cellwidth}{3:HL-LV\\HY-NL/1} & 
    \parbox{\cellwidth}{4:HL-IC\\HY-CD/1} & 
    \parbox{\cellwidth}{2:HL-WD\\LT-NL/1} & 
    \parbox{\cellwidth}{4:FL-BN\\LT-CD} & 
    \parbox{\cellwidth}{3:PK-WD\\LT-NL/1} & 
    \parbox{\cellwidth}{3:MN-LV\\LT-NL/1} & 
    \parbox{\cellwidth}{5:SF\\LT-NL/2}\\[\extralinespace]
    \rowcolor{grey}
    0 & 
    \parbox{\cellwidth}{5:FL-IC\\HY-NL} & 
    \parbox{\cellwidth}{3:HL-CR\\HY-NL/1} & 
    \parbox{\cellwidth}{3:FL-JU\\LT-HO/1} & 
    \parbox{\cellwidth}{3:HL-VO\\NW-CD/1} & 
    \parbox{\cellwidth}{2:MN-LV\\NW-NL/2} & 
    \parbox{\cellwidth}{3:MN-CR\\HY-NL/1} & 
    \parbox{\cellwidth}{3:MN-WD\\HY-NL/1} & 
    \parbox{\cellwidth}{3:FL-WD\\LT-NL/1} & 
    \parbox{\cellwidth}{3:MN-VO\\HY-NL/2} & 
    \parbox{\cellwidth}{3:HL-VO\\LT-NL/2}\\[\extralinespace]
    1 & 
    \parbox{\cellwidth}{4:MN-WD\\HY-NL} & 
    \parbox{\cellwidth}{3:MN-LV\\HY-NL/1} & 
    \parbox{\cellwidth}{4:PK-FT\\HY-HO} & 
    \parbox{\cellwidth}{5:FT-IC\\HY-CD} & 
    \parbox{\cellwidth}{2:HL-BN\\EX-NL/2} & 
    \parbox{\cellwidth}{4:FL-WD\\EX-NL} & 
    \parbox{\cellwidth}{3:HL-BN\\LT-HO} & 
    \parbox{\cellwidth}{3:HL-JU\\LT-NL/2} & 
    \parbox{\cellwidth}{3:MN-FT\\LT-NL/1} & 
    \parbox{\cellwidth}{2:FL-JU\\HY-NL/3}\\[\extralinespace]
    \rowcolor{grey}
    2 & 
    \parbox{\cellwidth}{4:MN-WD\\HY-NL} & 
    \parbox{\cellwidth}{6:SB\\HY-NL} & 
    \parbox{\cellwidth}{3:HL-LV\\HY-HO/1} & 
    \parbox{\cellwidth}{2:MN-BN\\LT-HO/1} & 
    \parbox{\cellwidth}{4:FL-BN\\LT-NL} & 
    \parbox{\cellwidth}{2:FL-LV\\LT-NL/1} & 
    \parbox{\cellwidth}{2:FT-LV\\LT-NL/1} & 
    \parbox{\cellwidth}{3:MN-FT\\HY-NL/1} & 
    \parbox{\cellwidth}{4:FL-WD\\HY-CD/1} & 
    \parbox{\cellwidth}{3:MN-CR\\NW-HO/3}\\[\extralinespace]
    3 & 
    \parbox{\cellwidth}{5:MN-JU\\EX-NL} & 
    \parbox{\cellwidth}{4:HL-LV\\LT-NL} & 
    \parbox{\cellwidth}{4:HL-JU\\HY-HO} & 
    \parbox{\cellwidth}{3:HL-BN\\NW-NL/1} & 
    \parbox{\cellwidth}{6:FL-IC\\EX-CD} & 
    \parbox{\cellwidth}{5:IN\\HY-NL} & 
    \parbox{\cellwidth}{4:FL-MA\\HY-NL/2} & 
    \parbox{\cellwidth}{2:MN-LV\\LT-NL/1} & 
    \parbox{\cellwidth}{4:FL-IC\\HY-NL/1} & 
    \parbox{\cellwidth}{2:HL-CR\\LT-NL/2}\\[\extralinespace]
    \rowcolor{grey}
    4 & 
    \parbox{\cellwidth}{5:NL-IC\\LT-NL} & 
    \parbox{\cellwidth}{4:FL-WD\\HY-HO} & 
    \parbox{\cellwidth}{2:MN-BN\\LT-NL/1} & 
    \parbox{\cellwidth}{4:HL-WD\\LT-NL} & 
    \parbox{\cellwidth}{3:FL-CR\\LT-NL/1} & 
    \parbox{\cellwidth}{3:FL-WD\\HY-NL/1} & 
    \parbox{\cellwidth}{2:HL-BN\\LT-HO/2} & 
    \parbox{\cellwidth}{3:HL-LV\\LT-NL/1} & 
    \parbox{\cellwidth}{4:MN-VO\\EX-CD/2} & 
    \parbox{\cellwidth}{2:HL-LV\\HY-NL/1}\\[\extralinespace]
    5 & 
    \parbox{\cellwidth}{4:HL-LV\\HY-NL} & 
    \parbox{\cellwidth}{4:FL-BN\\HY-NL} & 
    \parbox{\cellwidth}{3:HL-WD\\LT-NL/1} & 
    \parbox{\cellwidth}{5:FL-IC\\HY-NL} & 
    \parbox{\cellwidth}{3:HL-CR\\HY-CD/1} & 
    \parbox{\cellwidth}{5:IN\\LT-NL/1} & 
    \parbox{\cellwidth}{3:HL-VO\\HY-NL/1} & 
    \parbox{\cellwidth}{2:FL-LV\\EX-HO/3} & 
    \parbox{\cellwidth}{3:MN-WD\\EX-NL/2} & 
    \parbox{\cellwidth}{1:HL-WD\\HY-NL/4}\\[\extralinespace]
    \rowcolor{grey}
    6 & 
    \parbox{\cellwidth}{4:FL-LV\\HY-CD} & 
    \parbox{\cellwidth}{4:FL-LV\\LT-HO} & 
    \parbox{\cellwidth}{4:PK-LV\\HY-NL} & 
    \parbox{\cellwidth}{4:IN\\LT-NL/1} & 
    \parbox{\cellwidth}{3:PK-BN\\HY-HO/1} & 
    \parbox{\cellwidth}{3:FL-FT\\LT-CD/2} & 
    \parbox{\cellwidth}{3:MN-VO\\EX-NL/1} & 
    \parbox{\cellwidth}{3:FL-JU\\HY-HO/1} & 
    \parbox{\cellwidth}{5:PL-MA\\EX-HO/2} & 
    \parbox{\cellwidth}{2:FL-LV\\NW-HO/4}\\[\extralinespace]
    7 & 
    \parbox{\cellwidth}{4:MN-WD\\LT-HO} & 
    \parbox{\cellwidth}{3:HL-VO\\LT-NL/1} & 
    \parbox{\cellwidth}{4:MN-FT\\HY-NL} & 
    \parbox{\cellwidth}{2:FL-WD\\HY-HO/1} & 
    \parbox{\cellwidth}{3:SB\\EX-NL/3} & 
    \parbox{\cellwidth}{3:HL-FT\\HY-NL/1} & 
    \parbox{\cellwidth}{3:MN-FT\\LT-NL/2} & 
    \parbox{\cellwidth}{3:PK-LV\\HY-NL/1} & 
    \parbox{\cellwidth}{3:HL-JU\\LT-HO/2} & 
    \parbox{\cellwidth}{2:HL-CR\\LT-HO/4}\\[\extralinespace]
    \rowcolor{grey}
    8 & 
    \parbox{\cellwidth}{4:FL-BN\\HY-NL} & 
    \parbox{\cellwidth}{4:MN-FT\\HY-NL} & 
    \parbox{\cellwidth}{3:FL-FT\\HY-NL/1} & 
    \parbox{\cellwidth}{3:MN-FT\\HY-NL/1} & 
    \parbox{\cellwidth}{4:HL-WD\\EX-HO} & 
    \parbox{\cellwidth}{3:FL-FT\\LT-CD/1} & 
    \parbox{\cellwidth}{4:FL-IC\\HY-NL/1} & 
    \parbox{\cellwidth}{4:HL-FT\\EX-CD/1} & 
    \parbox{\cellwidth}{2:HL-CR\\HY-NL/1} & 
    \parbox{\cellwidth}{1:HL-LV\\HY-NL/2}\\[\extralinespace]
    9 & 
    \parbox{\cellwidth}{5:SB\\EX-NL} & 
    \parbox{\cellwidth}{3:HL-WD\\HY-NL} & 
    \parbox{\cellwidth}{4:HL-LV\\LT-HO} & 
    \parbox{\cellwidth}{3:HL-LV\\HY-HO/1} & 
    \parbox{\cellwidth}{4:FL-IC\\HY-NL/1} & 
    \parbox{\cellwidth}{4:FL-IC\\LT-NL/1} & 
    \parbox{\cellwidth}{4:SF\\HY-NL/1} & 
    \parbox{\cellwidth}{4:MN-JU\\EX-NL/1} & 
    \parbox{\cellwidth}{3:HL-VO\\HY-HO/1} & 
    \parbox{\cellwidth}{3:HL-FT\\HY-NL/1}\\[\extralinespace]
    \rowcolor{grey}
    10 & 
    \parbox{\cellwidth}{5:FL-MA\\EX-NL} & 
    \parbox{\cellwidth}{4:FL-JU\\LT-NL} & 
    \parbox{\cellwidth}{4:HL-JU\\LT-HO} & 
    \parbox{\cellwidth}{4:HL-LV\\LT-HO} & 
    \parbox{\cellwidth}{3:FL-WD\\LT-NL/1} & 
    \parbox{\cellwidth}{2:MN-LV\\LT-NL/1} & 
    \parbox{\cellwidth}{3:HL-WD\\LT-HO} & 
    \parbox{\cellwidth}{3:MN-FT\\NW-NL/1} & 
    \parbox{\cellwidth}{4:HL-FT\\EX-HO/2} & 
    \parbox{\cellwidth}{2:MN-BN\\HY-NL/2}\\[\extralinespace]
    11 & 
    \parbox{\cellwidth}{4:FL-JU\\LT-NL} & 
    \parbox{\cellwidth}{4:FL-WD\\HY-CD} & 
    \parbox{\cellwidth}{3:HL-BN\\LT-CD/1} & 
    \parbox{\cellwidth}{4:FL-BN\\HY-CD} & 
    \parbox{\cellwidth}{3:HL-FT\\EX-NL/1} & 
    \parbox{\cellwidth}{4:MN-LV\\HY-NL} & 
    \parbox{\cellwidth}{3:PK-BN\\LT-NL/1} & 
    \parbox{\cellwidth}{3:FL-WD\\LT-CD/1} & 
    \parbox{\cellwidth}{2:HL-CR\\EX-NL/2} & 
    \parbox{\cellwidth}{1:FL-LV\\LT-CD/3}\\[\extralinespace]
    \rowcolor{grey}
    12 & 
    \parbox{\cellwidth}{3:HL-BN\\HY-NL} & 
    \parbox{\cellwidth}{5:MN-FT\\EX-CD} & 
    \parbox{\cellwidth}{3:HL-CR\\HY-NL/1} & 
    \parbox{\cellwidth}{3:MN-CR\\HY-HO/1} & 
    \parbox{\cellwidth}{5:FL-CR\\EX-CD} & 
    \parbox{\cellwidth}{4:PL-WD\\LT-NL} & 
    \parbox{\cellwidth}{3:FL-FT\\HY-HO/1} & 
    \parbox{\cellwidth}{3:HL-FT\\LT-CD/1} & 
    \parbox{\cellwidth}{4:HL-FT\\EX-HO/1} & 
    \parbox{\cellwidth}{2:FL-CR\\LY-CD/3}\\[\extralinespace]
    13 & 
    \parbox{\cellwidth}{4:MN-CR\\HY-NL} & 
    \parbox{\cellwidth}{3:MN-BN\\LT-NL} & 
    \parbox{\cellwidth}{3:PK-CR\\LT-CD/1} & 
    \parbox{\cellwidth}{3:FL-BN\\EX-NL/1} & 
    \parbox{\cellwidth}{4:FL-MA\\EX-NL/1} & 
    \parbox{\cellwidth}{3:PK-CR\\HY-HO/1} & 
    \parbox{\cellwidth}{2:MN-BN\\EX-NL/2} & 
    \parbox{\cellwidth}{4:MN-JU\\HY-NL/1} & 
    \parbox{\cellwidth}{2:HL-BN\\HY-NL/1} & 
    \parbox{\cellwidth}{3:FL-MA\\HY-HO/4}\\[\extralinespace]
    \rowcolor{grey}
    14 & 
    \parbox{\cellwidth}{3:MN-LV\\EX-NL/1} & 
    \parbox{\cellwidth}{4:MN-BN\\EX-NL} & 
    \parbox{\cellwidth}{4:MN-FT\\EX-NL} & 
    \parbox{\cellwidth}{3:MN-VO\\EX-NL/1} & 
    \parbox{\cellwidth}{2:HL-WD\\HY-HO/1} & 
    \parbox{\cellwidth}{3:HL-LV\\EX-NL/1} & 
    \parbox{\cellwidth}{4:HL-VO\\EX-NL/1} & 
    \parbox{\cellwidth}{2:FL-LV\\LT-NL/1} & 
    \parbox{\cellwidth}{3:FL-WD\\HY-NL/2} & 
    \parbox{\cellwidth}{3:FL-LV\\EX-HO/1}\\[\extralinespace]
  \end{tabular}

  \medskip

  \parbox{0.9\textwidth}{Environ Skill Level:Coutour-Feature\\
    Gravity-Temperature/Urbal Skill Level (if any)

    All abbreviations are explained in the Environ, Gravity, and
    Temperature sections of Character Record. See
    \ref{sec:habitat-table} for explanation of use.}
\end{minipage}}
\end{table}
              

\subsection[Recording Habitat Table Info]{The player records the
    information obtained for his 
    character from the Habitat Table on the \index{character!record}Character Record as
    follows:}
\label{sec:habitat-table-use}

\begin{itemize}
\item If an Urban Skill Level is obtained, note the number in the
  Urban Skill Level box.
\item Place an \textbf{\textsf{X}} rather than a number in the
  Temperature space corresponding to the character's comfortable
  temperature range.
\item Record a \textbf{1} in the proper Gravity Type box.
  
  \index{gravity skill!reducing}\index{environ skill!increasing}If desired, a character's Gravity Skill Level may be increased at
  this time by \emph{reducing} the \emph{Environ Skill Level} the
  character received. Thus, a character that received an Environ Skill
  Level of \textbf{3} may receive a Gravity Skill Level of \textbf{2}
  (or \textbf{3}) in his gravity type by reducing his Environ Skill
  Level to \textbf{2} (or \textbf{1}). A character's Environ Skill
  Level may never be reduced below \textbf{1}. A character's
  \emph{Urban Skill Level} may not be reduced to increase his Gravity
  Skill Level. After recording the character's Gravity Skill Level,
  subtract \textbf{2} from the Level and record this result in the
  Gravity Type boxes \emph{adjacent} to the character's home Gravity
  Type box (even if the number is negative). Then subtract \textbf{2}
  \emph{more} and note this result in the box adjacent to the boxes
  just filled in. Repeat this once more if necessary. These numbers
  represent the character's Gravity Skill Levels in gravity types
  other than the one in which he grew up.
  
  \textbf{Example:} A character has a Skill Level of \textbf{1} in
  \emph{Light Gravity}. A \textbf{-1} is entered in the \emph{Near
    Weightless} box and in the \emph{Heavy} box. A \textbf{-3} is
  entered in the \emph{Extreme} box.
  
\item Record the character's Environ Skill Level in the proper box of
  the Environ Skill Display. The Level may be reduced if the
  character's Gravity Skills were increased (see preceding). Then
  \emph{reduce} the Environ Skill Level by \textbf{1} and enter this
  number in every box on the Display \emph{directly adjacent to a
    side} (not diagonal) of the character's Home Environ box. Reduce
  the Skill Level by \textbf{1} \emph{again} and enter this number in
  every box on the display directly adjacent to the boxes just filled
  in. Repeat this process until the entire Environ Skill Display is
  filled. Negative numbers will be used in some boxes. However, an
  Environ Skill may never be below \textbf{-6}. These numbers are the
  character's Environ Skill Levels in environ types other than the one
  in which he grew up.
\end{itemize}

\label{sec:play-records-inform-change-display}\textbf{Example:} The
player has filled out the Display (Figure 
\ref{fig:example-environ-disp}) to show the Environ Skill Levels that
his character (who has a Skill Level of \textbf{4} in his home
flat/jungle environ) has in all possible environs, plus a home gravity
type of \emph{light} and a home temperature range of \emph{normal}
(see Figure \ref{fig:example-environ-disp}).

\begin{figure}[htbp]
  \centering
  \fbox{%
    \begin{minipage}{0.95\textwidth}
      \centering
      \caption{Example Environ Skill Display}
      \label{fig:example-environ-disp}

      \medskip
      \begin{tabular}{l|c|c|c|c|c|c|c|c|c|c|c|c|c|}
        \multicolumn{1}{c}{}& \multicolumn{1}{c}{VO} &
        \multicolumn{1}{c}{CR} & \multicolumn{1}{c}{BN} &
        \multicolumn{1}{c}{LV} & \multicolumn{1}{c}{WD} &
        \multicolumn{1}{c}{FT} & \multicolumn{1}{c}{JU} &
        \multicolumn{1}{c}{MA} & \multicolumn{1}{c}{IC} &  
        \multicolumn{4}{l}{Urban Skill Level: 1}\\
        \cline{2-7}\cline{12-14}
        PK & --5 & --4 & --3 & --2 & --1 &  0 & \multicolumn{3}{c}{} &
        & Gravity &
        --1 & 0.0--0.4 G (NW)\\
        \cline{2-8}\cline{13-13}
        MN & --4 & --3 & --2 & --1 &  0 &  1 &  2 & \multicolumn{2}{c}{} &
        & Skill & 1 & 0.7--1.0 G (LT)\\
        \cline{2-10}\cline{13-13}
        HL & --3 & --2 & --1 &  0 &  1 &  2 &  3 &  2 &  1 &
        & Levels & --1 &
        1.3--1.7 G (HY)\\
        \cline{2-10}\cline{13-13}
        \multicolumn{1}{l}{FL} &    & --1 &  0 &  1 &  2 &  3 &  4 &  3 &
        2 &  &  & --3 & 2.0--2.5 G (EX)\\
        \cline{3-10}\cline{12-14}
        \multicolumn{8}{r}{IN} \vline &  2 & \multicolumn{1}{c}{} & & Temp &  &
        --50\textdegree--0\textdegree (CD)\\
        \cline{9-9}\cline{13-13}
        \multicolumn{8}{r}{SF} \vline & 1 & \multicolumn{1}{c}{} & & Range
        & \textsf{X} & 
        25\textdegree--100\textdegree (NL)\\
        \cline{9-9}\cline{13-13}
        \multicolumn{8}{r}{SB} \vline & 0 & \multicolumn{1}{c}{} & & & &
        125\textdegree--175\textdegree (HO)\\
        \cline{9-9}\cline{12-14}
      \end{tabular}
    \end{minipage}}
\end{figure}

\subsection[Deep-Space Environs]{The player may choose a \index{home environ!deep-space}deep-space
  environ for his character if the character meets the following
  prerequisites:}
\label{sec:deep-space}

\begin{itemize}
\item He has a \emph{Physique Multiplier} of \textbf{$\frac12$} or
  \textbf{1}. 
\item He has a \emph{Coordination Multiplier} of \textbf{3} or
  \textbf{4}.
\item His home planet gravity type is \emph{near weightless}.
\end{itemize}

The player may voluntarily declare that such a character is from a
deep space environ, such as an orbiting space station, a spaceport,
or an asteroid mining complex. The character's Habitat Table result is
ignored and his Character Record is \emph{not} filled out in accordance with
\ref{sec:habitat-table-use}. Instead, the following steps are
conducted:

\begin{enumerate}
\item The character receives an Urban Skill Level of \textbf{4}; note this in
  the Urban Skill Level box on the Character Record.
\item The character is accustomed to normal temperatures; place an
  \textbf{\textsf{X}} in the Normal Temperature space.
\item The character has a Near Weightless Gravity Skill Level of
  \textbf{3}; enter a \textbf{3} in the Near Weightless box, a
  \textbf{0} in the Light Gravity box, a \textbf{--2} in the Heavy
  Gravity box, and a \textbf{--5} in the Extreme Gravity box.
\item The character is accustomed to no natural environs; enter a
  \textbf{--4} in all 33 boxes of the Environ Skill Display.
\end{enumerate}

When using the Social Standing Table (see
\ref{sec:social-standing-table}), add twice the character's Social
Background Multiplier to the die result \emph{only} (the character's
Urban and Environ Skill Levels are ignored).

When expending Initial Skill Points (see \ref{sec:init-sp}), the
following skills may be chosen in addition to those listed for the
character in \ref{sec:fos-skills}: \index{agriculture!and deep space environ}agriculture, EVA, and asteroid
mining. These skills are \emph{not} available to the character when
expending Skill Points acquired from his profession (unless they are
listed in his profession or one of his fields of study).


\subsection[Social Standing Table]{The player uses the \index{social standing}Social Standing
  Table to determine his 
  character's family history and the amount of money with which he is
  endowed.}
\label{sec:social-standing-table}

 

The player rolls one die and modifies the die result as follows:

\begin{itemize}
\item \emph{Double} the character's \emph{Social Background
    Multiplier} and \emph{add} the product to the die result.
\item \emph{Add} the character's \emph{Urban Skill Level} to the die
  result.
\item \emph{Subtract} the character's \emph{Home Environ Skill Level}
  from the die result (use the number derived from the Habitat Table,
  before it was  reduced to increase Gravity Skill Levels).
\end{itemize}

Note the social standing result corresponding to the modified die
result on the Character Record.

Roll one die again and \emph{multiply} the die result by the amount of
\index{money!initial}money shown for the acquired social standing. Note this product on the
Character Record. Remember, one Tran equals 1,000 Mils.

The Initial Skill Point Modifier shown for the acquired social
standing is used in conjunction with \ref{sec:initial-skill-points}.

Aside from 
determining how much money and how many Initial Skill Points he
receives, a character's social standing will affect him during play.
When encountering NPC's and when dealing with bureaucracy and
government, the GM should take into account a character's social
standing. Class distinctions in \emph{Universe} are not as pronounced
as in, say, a feudal society; but people still feel most comfortable
with those of similar background.


\begin{table}[htbp]
  \centering
  \fbox{%
    \begin{minipage}{5in}
      \centering
      \caption{Social Standing Table}
      \label{tab:social-standing-table}

      \medskip

      \begin{tabular}{cp{2in}lc}
        \parbox{1.5cm}{\centering Modified Die Result} &   Social Standing &
        \parbox{1.5cm}{\centering Wealth: One Die $\times$} &
        \parbox{2.5cm}{\centering Initial Skill
          Point Modifier}\\
        \rowcolor{grey}
        --1 or Less & Unfostered Orphan & 1 Mil & +3\\
        0 & Oppressed minority & 10 Mils & 0\\
        \rowcolor{grey}
        1,2 & Poor colonist family & 10 Mils & +1\\
        3 & Family fallen on hard times & 100 Mils & +1\\
        \rowcolor{grey}
        4 & Communal colony & 10 Mils & +2\\
        5--7 & Lower middle class family & 100 Mils & 0\\
        \rowcolor{grey}
        8--10 & Skilled tech family & 1 Tran & +1\\
        11--13 & Local establishment (business or farm run by family) & 1
        Tran & --1\\
        \rowcolor{grey}
        14--15 & Independent trading family & 10 Trans & +2\\
        16--17 & Family runs planetary corporation & 10 Trans & --1\\
        \rowcolor{grey}
        18 & Family dominates planet & 100 Trans & --2\\
        19 & Family dominates star system & 100 Trans & --2\\
        \rowcolor{grey}
        20 & Family influences entire Federation & 200 Trans & --3\\
        21 & Wealthy dilettante & 200 Trans & --4
      \end{tabular}

      \medskip

      \parbox{0.9\textwidth}{See \ref{sec:social-standing-table} for
      explanation of use.}
    \end{minipage}}
  \end{table}


\subsection[Initial Skill Points]{The player determines how many
  \index{skill points!initial}Initial Skill Points his character receives for experience gained in
  the early part of his life.}
\label{sec:initial-skill-points}

The player rolls one die, adds the Initial Skill Point Modifier
(listed with the character's social standing on the Social Standing
Table) to the die result, and refers to the \textbf{Initial Skill
  Points Table}
(table \vref{tab:initial-skill-points}).

A player expends
Initial Skill Points after choosing fields of study for his character
(see \ref{sec:init-sp}).

           

\begin{table}[htbp]
  \centering
  \fbox{%
    \begin{minipage}{.6\textwidth}
      \centering
      \caption{Initial Skill Points}
      \label{tab:initial-skill-points}

      \medskip

      \begin{tabular}{ll}
        If the modified  result is{\ldots} &  The character
        receives{\ldots}\\
        \rowcolor{grey}
        0 or less &  1 Initial Skill Point\\
        1--3 &  2 Initial Skill Points\\
        \rowcolor{grey}
        4--6 &  3 Initial Skill Points\\
        7--9 &  4 Initial Skill Points\\
        \rowcolor{grey}
        10 or more &  5 Initial Skill Point\\
      \end{tabular}
      
      \medskip
      
      \parbox{0.9\textwidth}{Roll one die and add the Initial Skill Point
        Modifier (listed with character's social standing on the
        Social Standing Table) to the die result.}
    \end{minipage}}
\end{table}


\section{\index{character!development}Character Development}
\label{sec:char-devel}

Each character undertakes one or more fields of study during his teen
years. His schooling provides him with basic skills and, along with
his potentials, determines his final characteristics.


\subsection[Study Points and Fields of Study]{The player expends
  \index{study points|bb}\index{fields of study|bb}Study
  Points to have his character undertake any of 7 fields of study.}
\label{sec:fos}

These include theoretical science, applied science, business, the
humanities, the mind, the body, and the military. An eighth field of
study, general, is automatically undertaken by every character, with
no charge in Study Points.

Each field of study costs \textbf{1} Study Point to enter
(\textbf{Exception:} See the \emph{mind}, following). The player
declares which fields of study his character is entering and notes the
choices on his Character Record. A character may enter any given field
of study \emph{twice} at a cost of \textbf{2} Study Points
(representing heavy concentration by the character in that field). If
this is done, write the field of study twice on the Character Record.
A given field of study may not be studied more than twice.

\textbf{Restrictions on the Study of the \index{fields of study!mind,
    restrictions on study}Mind}: A character must have an
\emph{Intellect} Potential of at least \textbf{2} in order to study
the mind. The expenditure of \textbf{2} Study Points is required to
study the mind \emph{once}; \textbf{4} Study Points must be expended
to study the mind \emph{twice}. A character may \emph{not} study the
mind if he is studying the \emph{military} and/or \emph{business}.

A player should expend all his available Study Points when generating
his character. Unused Study Points are useless during play.


\subsection[Initial Skill Points]{The player expends \index{initial skill
  points|bb}Initial Skill
  Points to give his character basic skills acquired in youth.}
\label{sec:init-sp}



The expenditure of \textbf{1} Initial Skill Point allows the character
to receive any skill for which he is eligible at Skill Level
1 \emph{only}. No more than \textbf{1} Initial Skill Point
may be expended for a particular skill at this time.

The types of skills that may be chosen for a character depend on his
fields of study. The skills listed in \ref{sec:fos-skills} may be chosen for
a character that has undertaken the appropriate field of study. At
this time, a character may receive no skill that is not mentioned in
one of his chosen fields of study. All characters may receive skills
in the \emph{general} category on the skill list.

Initial Skill Points that are not expended before determining
characteristics are lost.


\subsection[Initial Skills]{\index{skills!initial}The following skills may be chosen for a
  character that has undertaken the appropriate field of study.}
\label{sec:fos-skills}

See Chapter \ref{cha:skills} for a detailed explanation of each of these
skills.

\begin{description}
\item[Theoretical Science:]\index{fields of study!theoretical science} Chemistry, physics, programming, biology,
  geology, astronomy.
\item[Applied Science:]\index{fields of study!applied science} Suit tech, electro tech, construction, vehicle
  tech, programming.
\item[Business:]\index{fields of study!business} Programming, recruiting, law, economics, trading.
\item[Humanities:]\index{fields of study!humanities} Linguistics, diplomacy, law, teaching.
\item[The Mind:]\index{fields of study!mind} Psionic boost, psionic communication, life sense.
\item[The Body:]\index{fields of study!body} Unarmed combat, ambush, EVA, gravity (home gravity
  type only), jetpack, survival.
\item[Military:]\index{fields of study!military} Battlefield, longarms, handguns, demolitions,
  grenades.
\item[General:]\index{fields of study!general} Streetwise, laser/stun pistol, gambling, blades,
  ground vehicles, urban (only if character already has an Urban Skill
  Level of \textbf{1} or greater), environ (home environ only).
\end{description}


\subsection[Generating Characteristics]{\index{characteristics!generating}The player determines his
  character's 9 characteristics one at a time, by using the
  Characteristic Modifier Chart and the Characteristic Generation
  Table.}
\label{sec:characteristics}


The Characteristic Modifier Chart is used to calculate the effect of
the character's potentials and fields of study on each characteristic.
The Characteristic Generation Table is used to find the actual rating
of each characteristic.

For each characteristic, the player completes the following steps:  

\begin{enumerate}
\item \emph{Add together} all numbers listed for the characteristic
  on the Modifier Chart that correspond to the fields of study
  undertaken by the character. If a given field was studied
  \emph{twice}, double the corresponding number. Note that every
  character is considered to have undertaken the \emph{general} field
  of study and may add its number to the total.\label{item:chargen-1}
\item Check the \emph{potential} section of the chart to find which
  potential affects the characteristic and multiply the total
  determined in Step \ref{item:chargen-1} by the Potential Multiplier.
  If \emph{two} potentials affect the characteristic, multiply the
  total by the first Potential Multiplier and then multiply this
  product by the second Potential Multiplier.\label{item:chargen-2}
\item Roll percentile dice and add the final product derived in Step
  \ref{item:chargen-2} to the dice result. Locate this modified result
  on the Characteristic Generation Table to find the rating for the
  characteristic. Mark the rating on the Character Record.
\end{enumerate}

\textbf{Important:} If an \emph{\index{agility!minimum}Agility Rating} of \emph{less than}
\textbf{5} is received at this time, it is considered a \textbf{5}.
\index{mental power!rounded down}The \emph{Mental Power Rating} is \emph{divided by \textbf{2}}
(rounding fractions down) before being noted on the Character Record.
The Mental Power Rating may range from \textbf{0} to \textbf{6}.

This procedure is conducted nine times when generating a
character --- once for each characteristic.

Note that no potentials affect the \index{aggression!and potentials}aggression characteristic. The
modifiers listed under aggression on the Modifier Chart are simply
added to (or subtracted from) the characteristic generation dice
result, as appropriate.


\subsection{Characteristic Modifier Chart}
\label{sec:char-modif-chart}

See page \pageref{tab:char-mod}.

\begin{table}[htbp]
  \setlength{\tabcolsep}{0.5\tabcolsep}
  \centering
  \fbox{%
    \begin{minipage}{4.75in}
      \centering
      \caption{Characteristic Modifier Chart}
      \label{tab:char-mod}

      \medskip

      \begin{tabular}{lccccccccr}
        Characteristic\\
        Field of Study      & ST & EN & DX & AY & IN & MP & LD & EM &  \multicolumn{1}{c}{AG}\\[1ex]
        Theoretical Science &  0 &  0 &  0 &  0 &  8 &  2 &  0 &  0 &   0\\
        \rowcolor{grey}
        Applied Science     &  0 &  0 &  4 &  0 &  8 &  0 &  0 &  0 &   0\\
        Business            &  0 &  0 &  0 &  0 &  0 &  0 &  2 &  4 &   0\\
        \rowcolor{grey}
        Humanities          &  0 &  0 &  0 &  0 &  6 &  2 &  0 &  4 & --15\\
        The Mind            &  0 &  2 &  0 &  0 &  4 &  6 &  0 &  0 & --25\\
        \rowcolor{grey}
        The Body            &  8 &  4 &  4 &  4 &  0 &  0 &  0 &  0 & +15\\
        Military            &  4 &  5 &  0 &  4 &  0 &  0 &  5 &  0 & +25\\
        \rowcolor{grey}
        General             &  1 &  1 &  1 &  1 &  1 &  1 &  1 &  1 &   0\\
        Multiply by: & \multicolumn{1}{c}{Ph,} & \multicolumn{1}{c}{Ph}
        & \multicolumn{1}{c}{Co,} & \multicolumn{1}{c}{Ph,} &
        \multicolumn{1}{c}{In} & \multicolumn{1}{c}{In} &
        \multicolumn{1}{c}{In,} &
        \multicolumn{1}{c}{In}\\
         & \multicolumn{1}{c}{Co} & 
        & \multicolumn{1}{c}{In} & \multicolumn{1}{c}{Co} &
        & &
        \multicolumn{1}{c}{SB} &
        \\
      \end{tabular}

      \medskip

      \parbox{0.9\textwidth}{Modifiers are applied to the
        percentile dice roll for the listed characteristic if the
        character studied the listed field. All applicable modifiers for a
        single characteristic are  added together. Then the total is
        multiplied by the Potential Multipliers that apply to that
        characteristic.
        
        See \ref{sec:characteristics} for further explanation.}
    \end{minipage}}
\end{table}


\subsection{Characteristic Generation Table}
\label{sec:char-gener-table}

See page \pageref{tab:characteristic-gen}.

\begin{table}[htbp]
  \centering
  \fbox{%
    \begin{minipage}{2.65in}
      \centering
      \caption{Characteristic Generation Table}
      \label{tab:characteristic-gen}

      \medskip

      \begin{tabular}{lc}
        \multicolumn{1}{c}{Percentile Dice} & Characteristic\\
        \multicolumn{1}{c}{Result + Modifier} & Rating\\[1ex]
        7 or less & 1\\
        \rowcolor{grey}
        8--15 & 2\\
        16--24 & 3\\
        \rowcolor{grey}
        25--34 & 4\\
        35--57 & 5\\
        \rowcolor{grey}
        58--83 & 6\\
        84--96 & 7\\
        \rowcolor{grey}
        97--108 & 8\\
        109--119 & 9\\
        \rowcolor{grey}
        120--129 & 10\\
        130--139 & 11\\
        \rowcolor{grey}
        140 or more & 12\\
      \end{tabular}
      
      \medskip

      \parbox{0.9\textwidth}{When generating the Mental Power
        Rating, divide the rating  by 2 (rounding fractions down). An
        Agility Rating of less than 5 is considered a 5.
        
        See \ref{sec:characteristics} for explanation of use.}
    \end{minipage}}
\end{table}

\section{\index{professions}Character Professions}
\label{sec:char-prof}

Every character in \emph{Universe} undertakes a profession --- a
particular area of employment where he may hone his abilities, gain
skills, accumulate wealth, and/or acquire weapons and equipment
otherwise difficult to obtain. There are 
% 23
\ref{item:myprofs}
professions, including
seven branches of military service. However, the character must
fulfill certain requirements to enter a profession, and sometimes
these are quite strict. A character with poor attributes can always be
a \emph{colonist} for a short while and start his adventuring early,
while still young and naive. Each character practices his profession
for a number of years and, if he has been in the military, is then
discharged; if he has been in a commercial profession, he is
considered ``for hire'' at the end of his term of employment, and his
life of individual exploits then commence.

Each character is considered to be 20 years old upon entering a
profession. The player should look over the entire list of professions
before choosing one, and then note his choice on the Character Record.
The procedures of the following Cases should be carried out in order
by each player that is generating a character.


\subsection[Prerequisites]{Each profession has a list of \index{professions!prerequisites}prerequisites
  that a 
  character must  fulfill in order to enter that profession.}
\label{sec:each-profession-has}

Certain fields of study must have been undertaken by the character.
Some of his Characteristic Ratings must meet certain minimums. In some
professions, a certain amount of initial wealth is required and/or
other special requirements must be met. A character that does not meet
all the prerequisites of a given profession may not enter that
profession.


\subsection[Mental Power Waiver]{A character with a \index{mental power!waiver}Mental Power
  Rating of 5 or 6 may enter 
  certain professions whether or not he  meets any prerequisites of
  that profession.}
\label{sec:mental-power-waiver}

Such professions list a \emph{mental power waiver} in their
prerequisites, and include star sailors, scouts, space pirates,
armed traders, and merchant explorers.

A character that enters one of these professions in this manner is
considered to be a \emph{Psionic navigator}, an interstellar pilot
whose Psionic powers enhance the efficiency and accuracy of
hyperjumping. Such a character must spend \textbf{2} Skill Points
acquired from his profession (see \ref{sec:each-prof-descr}) to
achieve \emph{Navigator} Skill Level 2 and \textbf{2} more Skill
Points to achieve \emph{Pilot} Skill Level 2. If the character
receives fewer than \textbf{4} Skill Points for practicing his
profession, he must apply all the points to these two skills as evenly
as possible. Other than these requirements, a Psionic navigator
undertakes his profession like any other character. If he wishes,
however, he may choose additional Skill Levels in navigation and
piloting within the restrictions of \ref{sec:skill-points-are}.


\subsection[Years In Profession]{The player must declare how many
  \index{professions!years in}years his character 
  wishes to spend at his profession and then must use the Employment
  Table to determine how many years are actually spent in the
  profession.}
\label{sec:player-must-declare}

The player declares that his character wishes to spend \textbf{4},
\textbf{8}, \textbf{12}, \textbf{16} or \textbf{20} years at his
profession. Then he carries out one of the following procedures:

\begin{itemize}
\item \emph{If the character is a colonist or a thinker}, the declared
  number of years is the \emph{actual} number of years employed. The
  Employment Table is not used. Add \textbf{20} to the number of years
  spent in the profession and note this sum as the character's age on
  the Character Record.
\item \emph{If the character is in a military profession}, roll one
  die and cross-reference the die result with the column of the
  Employment Table corresponding to the declared years of employment.
  The result represents the \emph{actual} number of years the
  character served before being discharged. Add \textbf{20} to the actual
  number of years employed and note this sum as the character's age on
  the Character Record. The character is considered to be at the point
  in his life immediately after his discharge; his declared years of
  employment has no further effect on play and does not count against
  him if greater than his actual years of employment.
\item \emph{If the character is in any other profession}, add
  \textbf{20} to the \emph{declared} years of employment and note this
  sum as the character's age on the Character Record. Then roll one
  die and cross-reference the die result with the column of the
  Employment Table corresponding to the declared years of employment.
  The result represents the \emph{actual} number of years in which the
  character was gainfully employed. If the result is less than the
  \emph{declared} number of years, the difference in years is
  considered to be time of \emph{unemployment}; the character does not
  receive the benefits of experience from those years, but has aged
  through them.
\end{itemize}

\begin{table}[htbp]
  \centering
  \fbox{%
    \begin{minipage}{2in}
      \centering
      \caption{Employment Table}
      \label{tab:employment}

      \medskip

      \begin{tabular}{lrrrrr}
        & \multicolumn{5}{c}{\textbf{Declared Years of}}\\
        & \multicolumn{5}{c}{\textbf{Employment}}\\
        \textbf{Die}   & \textbf{4} & \textbf{8} & \textbf{12} &
        \textbf{16} & \textbf{20}\\
        \rowcolor{grey}
        \textbf{1--2} & 4 & 4 &  8 & 12 & 12\\
        \textbf{3--5} & 4 & 8 &  8 & 12 & 16\\
        \rowcolor{grey}
        \textbf{6--8} & 4 & 8 & 12 & 12 & 16\\
        \textbf{9--10} & 4 & 8 & 12 & 16 & 20\\
      \end{tabular}
      
      \medskip

      \parbox{0.9\textwidth}{A character in the Thinker or Colonist
        profession does not use this table. See
        \ref{sec:player-must-declare}.}
    \end{minipage}}
\end{table}


\subsection[Effects of Age]{If a character is 28 years of age or
  older, the player 
  must use the \index{age!effects of}Effects of Age Table to determine if any of the
  character's physical characteristics are permanently reduced.}
\label{sec:if-character-28}

The player rolls one die and \emph{adds} the character's \emph{age} to
the die result. He then locates the modified result on the Table to
find how many points must be removed from his character's
\index{endurance!decrease due to aging|see{age, effects of}}Endurance,
\index{agility!decrease due to aging|see{age, effects of}}Agility,
\index{strength!decrease due to aging|see{age, effects of}}Strength,
and/or \index{dexterity!decrease due to aging|see{age, effects of}}Dexterity Ratings. Points must be removed from among the
characteristics as evenly as possible \emph{in the order indicated}.
Erase the listed rating and replace it with the new reduced rating.
\textbf{Example:} A 36 year old character receives a reduction result
of \textbf{5} from the Table. The player reduces the character's
Endurance Rating by \textbf{2} and his Agility, Strength, and
Dexterity Ratings each by \textbf{1}.

A single rating may not be reduced below \textbf{1} in this manner. If
this were about to happen, skip to the next characteristic.

\begin{table}[htbp]
  \centering
  \fbox{%
    \begin{minipage}{2.25in}
      \centering
      \caption{Effects of Age Table}
      \label{tab:age}

      \medskip

      \begin{tabular}{lc}
        \textbf{1 Die + Age} & \parbox{1in}{\centering\textbf{Physical
        Characteristic Points Lost}}\\
        \rowcolor{grey}
        \textbf{34 or less} & 0\\
        \textbf{35, 36} & 1\\
        \rowcolor{grey}
        \textbf{37, 38} & 2\\
        \textbf{39, 40} & 3\\
        \rowcolor{grey}
        \textbf{41, 42} & 4\\
        \textbf{43, 44} & 5\\
        \rowcolor{grey}
        \textbf{45, 46} & 6\\
        \textbf{47 or more} & 7\\
      \end{tabular}

      \medskip

      \parbox{0.9\textwidth}{See \ref{sec:if-character-28} for explanation.}
    \end{minipage}}
\end{table}


\subsection[Skill Point Modifier]{Each profession description contains
  a \index{skill point modifier}Skill  Point 
  Modifier used with the Skill Point Table to determine the number of
  Skill Points the character  receives.}
\label{sec:each-prof-descr}

The player refers to the column of the table matching his character's
\emph{actual} years of employment. He then rolls one die,
\emph{adding} the character's \emph{Intelligence Rating} and the
\emph{Skill Point Modifier} to the die result. Locating the modified
die result in the proper column yields the number of Skill Points that
the character receives for practicing his profession.

\subsection[Skill Points]{\index{skill points!initial}Skill Points are expended to purchase a
  variety of skills 
  for the character, representing expertise acquired during his years
  of employment.}
\label{sec:skill-points-are}


A player may choose any skills for his character that are listed in
his profession description. He may also choose any skills in the
listings for the character's fields of study in \ref{sec:fos-skills}.
\textbf{Exception:} A character that studied the mind may choose or
improve only a skill from that field of study if his Mental Power
Rating is \textbf{4} or greater. Any skills listed in the
\emph{general} category of \ref{sec:fos-skills} are also available.
The expenditure of \textbf{1} Skill Point allows the character to
receive \textbf{1} Skill Level in any skill for which he is eligible.
Skill Points may be expended to receive more than one level in a given
skill (at the rate of \textbf{1} point per level), within the
following \index{skills!maximum level (generation)}restrictions:

\begin{itemize}
\item If the character was employed for \textbf{4} years, no single
  skill may exceed Level 3.
\item If the character was employed for \textbf{8} to \textbf{12}
  years, no single skill may exceed Level 4.
\item If the character was employed for \textbf{16} to \textbf{20}
  years, no single skill may exceed Level 5.
\end{itemize}

The skill list in each profession may place additional restrictions on
specific skills or may require the acquisition of a certain skill. A
character that received a skill by the expenditure of an
\emph{Initial} Skill Point (see \ref{sec:init-sp}) may now expend
Skill Points to increase the level of that skill, but he is still
subject to these restrictions.

A player may \index{environ skill!improving}improve an environ skill (if allowed by his fields of
study or profession) by expending a Skill Point. The expenditure of
\textbf{1} Skill Point allows the player to increase the Skill Level
in \textbf{1} environ \emph{and the} \textbf{4} \emph{environs
  directly} (\emph{not} diagonally) \emph{adjacent to it} on the
Environ Skill Display by \textbf{1} level each. Thus, a Skill Point
expended to increase woods/hills would also increase light
vegetation/hills, woods/mountains, woods/ flat, and forest/hills.
Environ, urban, and gravity Skill Levels are not affected by the years
of employment restrictions in this Case; however, the limits listed in
\ref{sec:environmental-skills} must be adhered to.


\index{skills!familiar (X)}After all chosen skills have been assigned Skill Levels, the player
places an \textbf{\textsf{X}} in each Skill Level space for every
skill his character was eligible to receive but did not take. This
includes all skills listed for his profession, his fields of study
and, if from a deep space environ, those listed in
\ref{sec:deep-space}. Skills marked in this manner are familiar to the
character and he will be able to acquire them easily during play (see
\ref{sec:experience-points}).

\begin{table}[htbp]
  \centering
  \fbox{%
    \begin{minipage}{2.30in}
      \centering
      \caption{Skill Point Table}
      \label{tab:skillpoint}
      
      \medskip
      
      \begin{tabular}{clllll}
        & \multicolumn{5}{c}{\textbf{Actual Years}}\\
        & \multicolumn{5}{c}{\textbf{Of Employment}}\\
        \textbf{Modified}\\
        \textbf{Die Result} & \textbf{4} & \textbf{8} & \textbf{12} &
        \textbf{16} & \textbf{20}\\
        \rowcolor{grey}
        2--5 & 1 & 4 &  6 &  7 &  7\\
        6--9  & 2 & 5 &  7 &  8 &  8\\
        \rowcolor{grey}
        10--13 & 2 & 6 &  8 &  9 & 10\\
        14--17 & 3 & 6 &  8 & 10 & 11\\
        \rowcolor{grey}
        18--21 & 3 & 7 &  9 & 11 & 12\\
        22--26 & 4 & 7 & 10 & 12 & 14\\
        \rowcolor{grey}
        27--32 & 4 & 8 & 11 & 13 & 15\\
      \end{tabular}
    \end{minipage}}
\end{table}

\subsection[Benefits]{Each profession description contains a variety of
  material \index{benefits}benefits the character may receive from  practicing that
  profession.}
\label{sec:benefits}

The player rolls one die and adds the character's actual years of
employment to the die result. He then locates the modified die result
on the Benefit Table to find his character's Benefit Level (a letter
from \textbf{A} to \textbf{F}). The character receives the benefits
listed for the acquired Benefit Level in the description of his
profession.

Benefits include weapons, equipment and/or cash. The capabilities of
weapons and equipment are detailed in Chapter \ref{cha:equipment}.
Cash is received as \emph{lump sum} (add this to any money the
character may possess) or as a \emph{multiplier} (multiply the listed
amount of money by this number). \index{uniform}Characters in a military profession
also receive a \emph{rank of office}, a duty uniform, a dress uniform
(uniforms are not listed in the benefits) and, in some instances, an
annual pension.

\textbf{GM Note:} The rank a character achieves in a military
profession will help him during play when dealing with other military
characters and non-player characters, especially if he is wearing his
uniform. For example, a captain of the space sailors may receive
special favours when travelling in a government spacecraft, such as
preferential accommodations or extra information concerning the
destination. The GM is responsible for handing out pensions (although
the player should record the amount in the Money section of his
Character Record). A character's first pension payment is received one
game-year after the character is generated. See \ref{sec:keeping-time}
for timekeeping guidelines.


\begin{table}[htbp]
  \centering
  \fbox{%
    \begin{minipage}{2.35in}
      \centering
      \caption{Benefits Table}
      \label{tab:benefits}
      
      \medskip
      
      \begin{tabular}{cc}
        \parbox{1in}{\centering 1 DIE + ACTIVE YEARS IN PROFESSION} &
        \parbox{1in}{\centering BENEFIT LEVEL}\\
        \rowcolor{grey}
        5--10  & A \\
        11--14 & B \\
        \rowcolor{grey}
        15--18 & C \\
        19--23 & D \\
        \rowcolor{grey}
        24--28 & E \\
        29--30 & F \\
      \end{tabular}

      \medskip

      \parbox{0.9\textwidth}{Explanation of each 
        result is listed in 
        appropriate 
        Profession. 
        See \ref{sec:benefits} for 
        explanation of use.}
    \end{minipage}}
\end{table}

\subsection[Profession Tables]{The Profession Tables include the
  Employment Table, 
  Effects of Age Table, Skill Point  Table, and Benefits Table.}
\label{sec:prof-tabl-incl}

See pages \pageref{tab:employment} through \pageref{tab:benefits}.


\subsection[Professions]{Each player may choose one of the following
%  23 
  \ref{item:myprofs}
  \index{professions}professions 
  for his character. }
\label{sec:each-player-may}

\pagebreak[2]

\noindent{\large\textbf{ASTROGUARD (Military)}}\label{sec:prof-astroguard}\index{astroguard}
\stepcounter{myprofs}

A member of a planetary or star system spaceship force. The
astroguards protect the space lanes from privateers and monitor all
ship traffic coming into and out of local space.  Usually operating in
groups of small craft, astroguards also respond to distress calls and
are sometimes called upon for
combat duty.

\renewcommand{\theenumi}{\textbf{\Alph{enumi}}}
\textbf{Prerequisites:} Study of the military; Aggression Rating of at
least \textbf{5}.

\textbf{Skill Point Modifier:} 1.

\textbf{Skills Available:} Gunnery, missile guidance, space tactics,
pilot, air vehicles, EVA, spaceship tech (\textbf{1} Level only).

\pagebreak[2]
\textbf{Benefits:}
\begin{enumerate}
\item Corporal: 300 Mils cash.
\item Flight Sergeant: 1 Tran cash.
\item Flight Lieutenant: 3 Trans cash; \hyperlink{tag:gravity-web}{internal gravity web}.
\item Squadron Leader: 10 Trans cash; 250 Mils per year pension;
  \hyperlink{tag:gravity-web}{internal gravity web}.
\item Wing Commander: 25 Trans cash; 750 Mils per year  pension;
  \hyperlink{tag:gravity-web}{internal gravity web}.
\item Group Captain: 60 Trans cash; 2.5 Trans per year pension;
  \hyperlink{tag:gravity-web}{internal gravity web}.
\end{enumerate}

\bigskip

\pagebreak[2]


\noindent{\large\textbf{CIVIL
    INSPECTOR}}\label{sec:prof-civil-inspector}\index{civil inspector}
\stepcounter{myprofs}

A civilian employed by the federation's Agency for Interstellar
Development. A civil inspector travels from world to world
administering the progress of colonization programs. He watches for
and prevents private exploitation of planets still in the early stages
of development. Although he receives some self-defense training (in
case of hostile merchants in isolated areas), a civil inspector
usually calls for military aid when things get rough.

\textbf{Prerequisites:} Study of the humanities and business;
Characteristic Ratings of at least Intelligence \textbf{6}, Leadership
\textbf{4}, Empathy \textbf{6}, and Mental Power \textbf{2}.

\textbf{Skill Point Modifier:} 4.  

\textbf{Skills Available:}  Handguns, unarmed combat, agriculture,
  biology, planetology, urban.

\pagebreak[2]
\textbf{Benefits:}

\begin{enumerate}
\item 250 Mils cash.
\item  700 Mils cash; \hyperlink{tag:translator}{translator}.
\item  2 Trans cash; \hyperlink{tag:translator}{translator} or \hyperlink{tag:commlink}{interstellar commlink}.
\item  7 Trans cash; \hyperlink{tag:translator}{translator} and \hyperlink{tag:commlink}{interstellar commlink}.
\item 20 Trans cash; \hyperlink{tag:bot-frazette-blue}{Frazette Blue
  robot} with \hyperlink{tag:bot-information}{information system} and 
  \hyperlink{tag:bot-recorder}{recorder system}.
\item 40 Trans cash; \hyperlink{tag:bot-manner-36sd}{Manner 36sd robot} with \hyperlink{tag:bot-information}{information system},
  \hyperlink{tag:bot-weapon-target}{weapon targeting system}, \hyperlink{tag:bot-weapon}{laser pistol system}, \hyperlink{tag:bot-recorder}{recorder system}.
\end{enumerate}

\bigskip

\pagebreak[2]

\noindent{\large\textbf{COLONIST}}\label{sec:prof-colonist}\index{colonist}
\stepcounter{myprofs}

A citizen of the federation, living and working on a newly settled
planet.

\textbf{Prerequisites:} None.

\textbf{Skill Point Modifier:} 0.

\textbf{Skills Available:}  Trading, air vehicles, marine vehicles,
agriculture, mining.

\pagebreak[2]
\textbf{Benefits:}
\begin{enumerate}
\item Nothing.
\item 1 Tran.
\item  2 Trans; \hyperlink{tag:cl6-laser}{Civ Level 6 laser pistol}.
\item  4 Trans; \hyperlink{tag:cl6-laser}{Civ Level 6 laser pistol}.
\item  8 Trans; \hyperlink{tag:auto}{car}.
\item  15 Trans; \hyperlink{tag:auto}{Civ Level 5 automobile}; \hyperlink{tag:cl6-laser}{Civ Level 6 laser pistol}.  
\end{enumerate}

\bigskip

\pagebreak[2]

\noindent{\large\textbf{DIPLOMAT}}\label{sec:prof-diplomat}\index{diplomat}
\stepcounter{myprofs}

An employee of the federation or any planet government serving as a
representative to other planets. Diplomats are ``couriers'' of
official decisions and agreements. As a symbol of goodwill, diplomats
enjoy great hospitality whenever visiting an area in an official
capacity.

\textbf{Prerequisites:} Study of the humanities and the mind;
Characteristic Ratings of at least Intelligence \textbf{7}, Empathy \textbf{7}, Mental
Power \textbf{2}. The Aggression Rating may not exceed \textbf{6}. The character's
initial wealth must be at least 1 Tran.

\textbf{Skill Point Modifier:} 8.

\textbf{Skills Available:}  Recruitment, programming, computer/robot
   tech (\textbf{1} Level only), urban.

\pagebreak[2]
\textbf{Benefits:}
\begin{enumerate}
\item 300 Mils cash.
\item 1 Tran cash; \hyperlink{tag:audio-sealed-case}{audio-sealed case}.
\item 3 Trans cash; \hyperlink{tag:audio-sealed-case}{audio-sealed case}; \hyperlink{tag:translator}{translator}.
\item 10 Trans cash; \hyperlink{tag:audio-sealed-case}{audio-sealed case}; \hyperlink{tag:translator}{translator}; interstellar
  commlink.
\item 25 Trans cash; \hyperlink{tag:bot-manner-38sdf}{Manner 38sdf robot} with \hyperlink{tag:bot-holographer}{holographer system},
  \hyperlink{tag:bot-language}{language system}.
\item 60 Trans cash; \hyperlink{tag:bot-soidistant-v-201}{Soidistant V-201} with \hyperlink{tag:bot-creative-thought}{creative thought system},
  \hyperlink{tag:bot-holographer}{holographer system}.
\end{enumerate}

\bigskip

\pagebreak[2]

\noindent{\large\textbf{DOCTOR}}\label{sec:prof-doctor}\index{doctor}
\stepcounter{myprofs}

An esteemed member of the medical profession. The state of medicine in
the 24th Century allows MDs to perform feats of healing approaching
the miraculous. Doctors work out of medical offices in urban areas or,
in sparsely populated colonies, travel from settlement to settlement
as called for.

\textbf{Prerequisites:} Study of theoretical and applied science;
   Characteristic Ratings of at least Dexterity  \textbf{7}, Intelligence  \textbf{8},
   Mental Power \textbf{2}, and Empathy  \textbf{6}.

\textbf{Skill Point Modifier:} 10.

\textbf{Skills Available:} Teaching, diagnosis, treatment.

\pagebreak[2]
\textbf{Benefits:}
\begin{enumerate}
\item \hyperlink{tag:first-aid-kit}{first aid kit}.
\item 3 Trans cash; \hyperlink{tag:first-aid-kit}{first aid kit}.
\item 1 Tran x Intelligence Rating in cash; \hypertarget{tag:cl6-mediscanner}{Civ Level 6 mediscanner}.
\item 4 Trans x Intelligence Rating in cash; \hypertarget{tag:cl6-mediscanner}{Civ Level 6 mediscanner}.
\item 10 Trans x Intelligence Rating in cash; \hypertarget{tag:cl8-mediscanner}{Civ Level 8 mediscanner}.
\item 20 Trans x Intelligence Rating in cash; 
%% Original:
%%Manner 36sdf robot with medical system.
%% New:
\hyperlink{tag:bot-manner-38sdf}{Manner 38sdf robot} with \hyperlink{tag:bot-medical}{medical system}.
%% End change

\end{enumerate}
%% Label placed here so it generates the right reference.
\label{sec:prof-doctor-bot}

\bigskip

\pagebreak[2]

\noindent{\large\textbf{ENFORCER}}\label{sec:prof-enforcer}\index{enforcer}
\stepcounter{myprofs}

A privately employed bodyguard, soldier, or security specialist.
Enforcers usually work for corporations that control a large area of a
world, executing the rules of the company or patrolling restricted
lands. Enforcers are sometimes hired by backwater colonies to quell
criminal activity or civil unrest. Enforcers are generally disliked by
those they are watching over.

\textbf{Prerequisites:} Study of the mind and the body; all Physical
Characteristics must be at least \textbf{4}; Intelligence must be at least \textbf{5};
Leadership must be at least \textbf{5}.

\textbf{Skill Point Modifier:} 6.

\textbf{Skills Available:} Battlefield, demolitions, machine guns,
longarms, handguns, paint gun, recruiting, law, military ground
vehicles.

\pagebreak[2]
\textbf{Benefits:}
\begin{enumerate}
\item 500 Mils cash.
\item 1.5 Trans cash.
\item 3 Trans cash; Civ Level 5 pistol.
\item 10 Trans cash; \hyperlink{tag:cl8-laser}{Civ Level 8 laser pistol}.
\item 25 Trans cash; \hyperlink{tag:cl8-paintgun}{Civ Level 8 paint gun}.
\item 60 Trans cash; \hyperlink{tag:cl8-paintgun}{Civ Level 8 paint gun}; Civ Level 5 submachine
  gun.
\end{enumerate}

\bigskip

\pagebreak[2]

\noindent{\large\textbf{EXPLORER}}\label{sec:prof-explorer}\index{explorer}
\stepcounter{myprofs}

A crewman or part-owner of an interstellar exploration vessel.
Explorers organize and carry out private expeditions to unsettled
planets, gathering information to facilitate future commercial
development. Often their efforts are financed by large corporations
and private investors. Explorers are most concerned with the
profitable gain of knowl edge; they rarely transport large cargoes and
their ships are not usually armed.

\textbf{Prerequisites:} Study of business and either theoretical or
applied science; Characteristic Ratings of at least Intelligence
\textbf{7}, Mental Power \textbf{3}, and Empathy \textbf{4}. The
character's initial wealth must be at least 500 Mils. The mental power
waiver applies (see \ref{sec:mental-power-waiver}).

\textbf{Skill Point Modifier:} 7.

\textbf{Skills Available:} Handguns, paint gun, gunnery, pilot,
linguistics, survival, air vehicles, marine vehicles, planetology,
biology, geology, astronomy, navigation, any environs, any gravity.

\pagebreak[2]
\textbf{Benefits:}
\begin{enumerate}
\item Nothing.
\item Initial wealth x 3.
\item Initial wealth x 5; Civ Level 8 \hyperlink{tag:bioscanner}{bioscanner}.
\item Initial wealth x 10; Civ Level 8 \hyperlink{tag:geoscanner}{geoscanner}.
\item Initial wealth x 25; \hyperlink{tag:neuroscanner}{neuroscanner}.
\item Initial wealth x 50; \hyperlink{tag:bot-manner-50sd}{Manner 50sd robot} with \hyperlink{tag:bot-bio}{bio system},
  \hyperlink{tag:self-activation}{self-activation} system, and \hyperlink{tag:holographer}{holographer} system.
\end{enumerate}

\bigskip

\pagebreak[2]

\noindent{\large\textbf{FREEFALLER (Military)}}\label{sec:prof-freefaller}\index{freefaller}
\stepcounter{myprofs}

A soldier in the zero gravity branch of the federal armed forces.
Freefallers are usually assigned to a fleet patrolling a single star
system. They attack and defend asteroid belts, small moons, and space
colonies, and specialize in spaceship boarding operations.

\textbf{Prerequisites:} Study of \textbf{2} of the following fields:
applied science, the body, and the military; Characteristic Ratings of
at least Dexterity \textbf{6}, Agility \textbf{7}, and Aggression
\textbf{4}. The character's home gravity type may not be \emph{heavy}
or \emph{extreme}.

\textbf{Skill Point Modifier:} 3.  

\textbf{Skills Available:} Near Weightless Gravity Skill Level is
automatically increased to \textbf{2} (if not already equal to or greater than
\textbf{2}) at no cost in Skill Points. Battlefield, body armor, machine guns,
arc guns, longarms, hand guns, jetpack, demolitions, EVA, suit tech,
weapon tech, near weightless gravity (in addition to any levels
received as described in the preceding).

\pagebreak[2]
\textbf{Benefits:}
\begin{enumerate}
\item Lancer: 300 Mils cash.
\item Corporal: 1 Tran cash; Civ Level 6 \hyperlink{tag:expedsuit}{expedition suit}.
\item Sergeant: 3 Trans cash; Civ Level 7 \hyperlink{tag:expedsuit}{expedition suit}.
\item Lieutenant: 10 Trans cash; 250 Mils per year pension; Civ Level
  7 \hyperlink{tag:expedsuit}{expedition suit}.
\item Captain: 25 Trans cash; 750 Mils per year pension; Civ Level 7
  \hyperlink{tag:expedsuit}{expedition suit}; \hyperlink{tag:arcgun}{arc gun}.
\item Colonel: 60 Trans cash; 2.5 Trans per year pension; Civ Level 7
  \hyperlink{tag:expedsuit}{expedition suit}; \hyperlink{tag:arcgun}{arc gun}; \hyperlink{tag:jetpack}{jetpack}.
\end{enumerate}

\bigskip

\pagebreak[2]

\noindent{\large\textbf{HANDYMAN}} \label{sec:prof-handyman}\index{handyman}
\stepcounter{myprofs}

An engineer or technician working with any kind of planet-based
technology. Handymen build and repair electrical devices, protective
suits, weapons, computers, robots, and vehicles. They are employed by
all branches of government, military and commercial concerns.

\textbf{Prerequisites:} Study of applied science \emph{twice}; Characteristic
Ratings of at least Dexterity \textbf{6} and Intelligence \textbf{6}.

\textbf{Skill Point Modifier:} 5.

\textbf{Skills Available:} Agriculture, programming, physics, energy
tech, spaceship tech (\textbf{1} Level only), weapon tech, compu/robot tech.

\pagebreak[2]
\textbf{Benefits:}
\begin{enumerate}
\item \hyperlink{tag:basicrepair}{Basic repair kit}. 
\item 1 Tran cash; \hyperlink{tag:basicrepair}{basic repair kit}.  
\item 3 Trans cash; Civ Level 7 \hyperlink{tag:electrokit}{electrokit}.   
\item 10 Trans cash; Civ Level 8 \hyperlink{tag:electrokit}{electrokit}.  
\item 20 Trans cash; Civ Level 8 \hyperlink{tag:vehiclekit}{vehicle kit}.  
\item 40 Trans cash; Civ Level 8 \hyperlink{tag:vehiclekit}{vehicle kit} and \hyperlink{tag:robotkit}{robot kit}.
\end{enumerate}

\bigskip

\pagebreak[2]

\noindent{\large\textbf{INTERSTELLAR
    TRADER}}\label{sec:prof-interstellar-trader}\index{interstellar trader}
\stepcounter{myprofs}

A crewman or part owner of an interstellar freight vessel. Traders
travel from world to world, purchasing and selling goods for profit.
Trader ships are usually armed, as a defense against space pirates and
unfriendly planet governments. Although traders stay on the right side
of the law, they are always willing to venture away from established
trade routes if good money can be made.

\textbf{Prerequisites:} Study of business; Characteristic Ratings of
at least Intelligence \textbf{6} and Mental Power \textbf{2}. The
character's initial wealth must be at least 1 Tran. The mental power
waiver applies (see \ref{sec:mental-power-waiver}).

\textbf{Skill Point Modifier:} 4.

\textbf{Skills Available:} Longarms, handguns, arc gun, gunnery,
missile guidance, space tactics, pilot, linguistics, diplomacy,
asteroid mining, astronomy, EVA.

\pagebreak[2]
\textbf{Benefits:}
\begin{enumerate}
\item Nothing; all initial wealth lost.   
\item Civ Level 7 \hyperlink{tag:buscomp}{business computer}.    
\item Initial wealth x 2; \hyperlink{tag:buscomp}{business computer}.  
\item Initial wealth x 5; \hyperlink{tag:buscomp}{business computer}.   
\item Initial wealth x 12; \hyperlink{tag:buscomp}{business computer}.   
\item Initial wealth x 30; \hyperlink{tag:bot-brummagen-ii}{Brummagen II robot} with \hyperlink{tag:bot-spacecraft-tech}{spaceship tech
  system}; \hyperlink{tag:buscomp}{business computer}.
\end{enumerate}

\bigskip

\pagebreak[2]

\noindent{\large\textbf{LAWMAN (Military)}}\label{sec:prof-lawman}\index{lawman}
\stepcounter{myprofs}

A member of a planetary law enforcement agency. Lawmen perform the
functions of today's policemen, protecting the citizenry and
investigating crimes. However, the sparse and spread out populations
of most worlds often require lawmen to travel from settlement to
settlement executing justice wherever necessary, much like the
marshals of the old west.

\textbf{Prerequisites:} Study of the body and the humanities; all Physical Characteristic Ratings must be at least  \textbf{4}.

\textbf{Skill Point Modifier:} 3.

\textbf{Skills Available:} Machine guns, longarms, handguns, paint
gun, air vehicles, marine vehicles, urban, any environs, any gravity.

\pagebreak[2]
\textbf{Benefits:}   
\begin{enumerate}
\item  Patrolman: 250 Mils cash.  
\item Constable: 700 Mils cash; \hyperlink{tag:cl6-laser}{Civ Level 6 laser pistol}.  
\item Sergeant: 2 Trans cash; \hyperlink{tag:cl8-laser}{Civ Level 8 laser pistol}.   
\item Marshal: 7 Trans cash; \hyperlink{tag:cl6-paint}{Civ Level 6 paint gun} and used rover (ATV).  
\item Inspector: 20 Trans cash; 500 Mils per year pension; \hyperlink{tag:cl8-paint}{Civ Level 8
  paint gun}.
\item Chief: 40 Trans cash; 2 Trans per year pension; \hyperlink{tag:cl8-paint}{Civ Level 8
  paint gun} and any hand gun.
\end{enumerate}

\bigskip

\pagebreak[2]

\noindent{\large\textbf{MERCHANT}}\label{sec:prof-merchant}\index{merchant}
\stepcounter{myprofs}

A trader of goods on the local (planetary) level. Merchants travel
from settlement to settlement on a sparsely populated world, or
arrange for transport of goods from point to point on a world.
Merchants handle every thing from basic foodstuffs to fine electronic
equipment to contraband.

\textbf{Prerequisites:} Study of business \emph{twice} or study of
business \emph{and} applied science; Characteristic Ratings of at
least Intelligence \textbf{5} and Empathy \textbf{7}. The character's
initial wealth must be at least 1 Tran.

\textbf{Skill Point Modifier:} 5.

\textbf{Skills Available:} Diplomacy, air vehicles, marine vehicles,
agriculture, electro tech, computer/robot tech, vehicle tech, urban,
any environs.

\pagebreak[2]
\textbf{Benefits:}
\begin{enumerate}
\item Nothing; all initial wealth lost.  
\item Civ Level 7 \hyperlink{tag:buscomp}{business computer}.   
\item Initial wealth x 3; Civ Level 7 \hyperlink{tag:buscomp}{business computer}.  
\item Initial wealth x 10; Civ Level 7 \hyperlink{tag:buscomp}{business computer}.  
\item Initial wealth x 20; Civ Level 5 truck.   
\item Initial wealth x 40; Civ Level 7 jet.
\end{enumerate}

\bigskip

\pagebreak[2]

\noindent{\large\textbf{RANGER (Military)}}\label{sec:prof-ranger}\index{ranger}
\stepcounter{myprofs}

A soldier in the standard federal ground force. A ranger's
responsibilities cover all aspects of line and garrison duty on a
planet surface, including air and marine patrol. During combat,
rangers come in as the ``second wave'' following the Space Troopers
and carrying out any ``mopping up'' operations.

\textbf{Prerequisites:} Study of the military and the body; all
 Physical Characteristic Ratings must be at least  \textbf{4}.

\textbf{Skill Point Modifier:} 2.  

\textbf{Skills Available:} Artillery, machine guns, paint gun, bows,
military ground vehicles, air vehicles, marine vehicles, weapon tech,
vehicle tech, treatment, any environs.

\pagebreak[2]
\pagebreak[2]
\textbf{Benefits:}
\begin{enumerate}
\item Ranger: 300 Mils cash; \hyperlink{tag:respirator}{respirator}.
\item Corporal: 1 Tran cash; Civ Level 6 \hyperlink{tag:respirator-helmet}{respirator helmet}.
\item Sergeant: 3 Trans cash; Civ Level 8 \hyperlink{tag:respirator-helmet}{respirator helmet}.
\item Lieutenant: 10 Trans cash; 250 Mils per year pension; Civ Level
  8 \hyperlink{tag:respirator-helmet}{respirator helmet}; \hyperlink{tag:cl6-laser}{Civ Level 6 laser pistol}.
\item Captain: 25 Trans cash; 750 Mils per year pension; Civ Level 8
  \hyperlink{tag:respirator-helmet}{respirator helmet}; \hyperlink{tag:cl8-laser}{Civ Level 8 laser pistol}.
\item Colonel: 60 Trans cash; 2.5 Trans per year pension; Civ Level 8
  \hyperlink{tag:respirator-helmet}{respirator helmet} and \hyperlink{tag:cl8-paintgun}{paint gun}.
\end{enumerate}

\bigskip

\pagebreak[2]

\noindent{\large\textbf{REPORTER/JOURNALIST}}\label{sec:prof-reporter-journalist}\index{reporter/journalist}
\stepcounter{myprofs}

A newsperson or media personality, this individual can become widely
known, can establish contacts on many levels of federation society,
from the underworld to the high echelons of government, and can be
sent to dangerous and exotic places on assignment at no cost to self
(other than to his safety).

\textbf{Prerequisites:} Study of the humanities and either business or
theoretical science; Characteristic Ratings of at least Intelligence
\textbf{7}, Mental Power \textbf{2}, Aggression \textbf{5}.

\textbf{Skill Point Modifier:} 7.

\textbf{Skills Available:}  Unarmed combat, survival, disguise,
   forgery/counterfeiting, urban, air vehicles, programming, electro
   tech, recruiting.

\pagebreak[2]
\textbf{Benefits:} All reporters receive an audio recorder.
\begin{enumerate}
\item Nothing additional.
\item 2 Trans cash; photographic equipment.
\item Credentials with planetary news service; 500 Mils x Intelligence
  Rating in cash; \hyperlink{tag:superoid}{superoid camera}.
\item Credentials with star system news service; 3 Trans x
  Intelligence Rating in cash; basic \hyperlink{tag:holographer}{holographer}.
\item Credentials with federation-wide news service; 10 Trans x
  Intelligence Rating in cash; shoulder \hyperlink{tag:holographer}{holographer}.
\item Media celebrity throughout federation; 25 Trans x Intelligence
  Rating in cash; \hyperlink{tag:bot-manner-38sdf}{Manner 38sdf robot} with \hyperlink{tag:bot-information}{information system},
  \hyperlink{tag:bot-holographer}{holographer system}.
\end{enumerate}

The GM should take a reporter's status into account when the character
requests information, assist ance, or financial backing from
commercial, government, or media of ficials.

\bigskip

\pagebreak[2]

\noindent{\large\textbf{SCIENTIST}}\label{sec:prof-scientist}\index{scientist}
\stepcounter{myprofs}

A researcher or inventor employed by the federation or by a private
concern. Scientists specialize in any of the physical sciences and are
the powers behind the advanced technology and discoveries of the 24th
Century. As such, any expedition or venture of an exploratory nature
will receive much more support if a qualified scientist participates.

\textbf{Prerequisites:} Study of \textbf{2} of the following: theoretical
science, applied science, or the humanities; Intelligence Rating of at
least \textbf{8}.

\textbf{Skill Point Modifier:} 9.  

\textbf{Skills Available:} Teaching, agriculture, chemistry,
planetology, programming (\textbf{1} Level only), biology, geology, astronomy,
energy, computer/robot tech (\textbf{1} Level only).

\pagebreak[2]
\textbf{Benefits:}
\begin{enumerate}
\item 1 Tran cash.  
\item 2 Trans cash; Civ Level 5 \hyperlink{tag:chemlab}{chemlab}.
\item 500 Mils x Intelligence Rating in cash; Civ Level 6
  \hyperlink{tag:chemsynth}{chemsynthesizer}.
\item 2 Trans x Intelligence Rating in cash; Civ Level 7 \hyperlink{tag:chemlab}{chemlab}.
\item 6 Trans X Intelligence Rating in cash; Civ Level 8 \hyperlink{tag:bioscanner}{bioscanner}
  and \hyperlink{tag:geoscanner}{geoscanner}.
\item 15 Trans x Intelligence Rating in cash; \hyperlink{tag:bot-frazette-amber}{Frazette Amber robot}
  with \hyperlink{tag:bot-geo}{geo system}, \hyperlink{tag:bot-chemical}{chemical system}, \hyperlink{tag:bot-bio}{bio system}, and \hyperlink{tag:bot-information}{information
  system}.
\end{enumerate}

\bigskip

\pagebreak[2]

\noindent{\large\textbf{SCOUT (Military)}}\label{sec:prof-scout}\index{scout}
\stepcounter{myprofs}

A member of the exploration branch of the federal armed forces. Scouts
lead the way to unexplored worlds, gathering specimens and
information, forging the path for future colonization. Scouts usually
work in small teams when conducting an expedition and sometimes work
alone. Although administered by the military, the scouts rarely
participate in actual combat.

\textbf{Prerequisites:} Study of theoretical or applied science; all
  Physical Characteristics  must be at least  \textbf{3}; Characteristic
  Ratings of at least Mental Power  \textbf{2} and Empathy  \textbf{6}; the mental power
  waiver applies  (see \ref{sec:mental-power-waiver}).

\textbf{Skill Point Modifier:} 4.

\textbf{Skills Available:}  Longarms, handguns, pilot, EVA, survival,
   air vehicles, marine vehicles, jetpack, planetology, treatment,
   biology, geology, astronomy, suit tech, any environs, any gravity.

\pagebreak[2]
\textbf{Benefits:}
\begin{enumerate}
\item Scout Second Class: 300 Mils cash.
\item Scout First Class: 1 Tran cash; Civ Level 6 \hyperlink{tag:expedsuit}{expedition suit}.
\item Master Scout: 3 Trans cash; Civ Level 7 \hyperlink{tag:expedsuit}{expedition suit}.
\item Single Scout: 5 Trans cash; Civ Level 7 \hyperlink{tag:expedsuit}{expedition suit},
  \hypertarget{tag:neuroscanner}{neuroscanner}, and \hypertarget{tag:rover}{rover}.
\item Expedition Leader: 25 Trans cash; 750 Mils per year pension; Civ
  Level 7 \hyperlink{tag:expedsuit}{expedition suit} and \hyperlink{crawler}{crawler}.
\item Scout Commodore: 60 Trans cash; 2.5 Trans per year pension; Civ
  Level 7 \hyperlink{tag:expedsuit}{expedition suit} and used \hyperlink{tag:pod-explorer}{explorer pod}.
\end{enumerate}

\bigskip

\pagebreak[2]

\noindent{\large\textbf{SPACE
    PIRATE}}\label{sec:prof-space-pirate}\index{space pirate}
\stepcounter{myprofs}

A crewmember or owner of an outlaw spaceship. Space pirates waylay
commercial vessels in interplanetary space, smuggle legal and illegal
goods, and terrorize asteroid mining operations and small colonies.
Pirate vessels may sometimes be chartered for secret journeys to
distant parts at exorbitant prices.

\textbf{Prerequisites:} All Physical Characteristics must be at least
\textbf{5}; Characteristic Ratings must be at least Aggression
\textbf{7} and Mental Power \textbf{2}. The character \emph{must} have
an initial wealth \emph{less than} \textbf{4} Trans and must
\emph{not} have studied the humanities. The mental power waiver
applies (see \ref{sec:mental-power-waiver}).

\textbf{Skill Point Modifier:} 6.

\textbf{Skills Available:} Body armor, handguns, paint gun, arc gun,
bows, unarmed combat, gunnery, missile guidance, space tactics, pilot,
linguistics, disguise, forgery/counterfeiting, EVA, air vehicles,
asteroid mining, planetology, astronomy, spaceship tech, weapon tech,
near weightless gravity.

\pagebreak[2]
\textbf{Benefits:}\\
All space pirates receive an \hyperlink{tag:gravity-web}{internal gravity web}.   
\begin{enumerate}
\item Nothing; all initial wealth lost.  
\item Initial wealth x 2.   
\item Initial wealth x 4; wanted by federation.
\item Initial wealth x 10.   
\item Initial wealth x 10; \hyperlink{tag:terwillicker-5000}{Terwillicker 5000 Battlecraft}; wanted by
  federation.
\item Initial wealth x 20; \hypertarget{tag:harmonics-piccolo}{Piccolo spaceship} with \hyperlink{tag:pod-light-weapon}{light weapon pod}.
\end{enumerate}

Before beginning play, a space pirate may turn any amount of his money
into booty. He may purchase any weapons, robots, or equipment as if he
were on a Civ Level 6 world with no law restrictions. No bargaining is
allowed. If a space pirate is wanted by the federation, the GM should
take this into account when the character encounters any planetary law
enforcement NPC's, and especially when he encounters any federal
authorities (see \ref{sec:creating-encounters}). The character should
be kept constantly on the run.

\bigskip

\pagebreak[2]

\noindent{\large\textbf{SPACE
    TECHNICIAN}}\label{sec:prof-space-technician}\index{space technician}
\stepcounter{myprofs}

A specialist in the design, construction, and repair of spaceships and
related equipment. Space techs are
indispensable members of any lengthy space voyage. Usually  employed
by commercial shipping companies and  ship manufacturers, space techs
are sometimes called upon by governments and military branches for
their specialized  services.

\textbf{Prerequisites:} Study of applied science  \emph{twice};
Characteristic  Ratings of at least  Dexterity  \textbf{7}, Intelligence  \textbf{7},
and Mental Power \textbf{2}.   

\textbf{Skill Point Modifier:} 8.

\textbf{Skills Available:}  Psion tech ( \textbf{1} Level only), EVA,
   programming, astronomy, physics, energy tech, spaceship tech,
   compu/robot tech.

\pagebreak[2]
\textbf{Benefits:}
\begin{enumerate}
\item 500 Mils cash.   
\item 2 Trans cash.  
\item 5 Trans cash; Civ Level 7 \hyperlink{tag:electrokit}{electrokit}.    
\item 15 Trans cash; Civ Level 8 \hyperlink{tag:electrokit}{electrokit}.   
\item 30 Trans cash; Civ Level 7 \hyperlink{tag:spaceshipkit}{spaceship kit}.    
\item 60 Trans cash; \hyperlink{tag:bot-frazette-amber}{Frazette Amber
  robot} with  \hyperlink{tag:bot-spaceshiptech}{spaceship tech system},
   \hyperlink{tag:bot-electrotech}{electro tech system}, \hyperlink{tag:bot-self-activation}{self-activation system}.
\end{enumerate}

\bigskip



\pagebreak[2]

\noindent{\large\textbf{SPACETROOPER (Military)}} \label{sec:prof-spacetrooper}\index{spacetrooper}
\stepcounter{myprofs}

A soldier in the federal assault force, specializing in body armor
combat on planet surfaces. Troopers are famous for their mass
assaults, rapid victories, and \emph{esprit de corps}. Spacetroopers
are always at the fore of any major attack and move on once the way is
cleared for the standard military forces. Although their operations
are sometimes launched from the air, spacetroopers are trained
exclusively for ground combat.

\textbf{Prerequisites:} Study of the military and the body; all
Physical Characteristics must be at least \textbf{6}; Aggression
Characteristic Rating must be at least \textbf{5}. The character's
home gravity type may not be \emph{Near Weightless}.

\textbf{Skill Point Modifier:} 5.

\textbf{Skills Available:} All Gravity Skill Levels except Near
Weightless are \emph{automatically} increased to \textbf{0} (if not
already equal to or greater than 0) at no cost in Skill Points. Body
armor (\textbf{1} Level required), artillery, machine guns, paint
gun, arc gun, military ground vehicles, suit tech, any environs, any
gravity except Near Weightless (in addition to any levels received as
described above).

\pagebreak[2]
\textbf{Benefits:} 
\begin{enumerate}
\item Trooper: 300 Mils cash; Civ Level 6 \hyperlink{tag:armor-vest}{armor vest} and \hyperlink{tag:respirator-helmet}{respirator
  helmet}.
\item Corporal: 1 Tran cash; Civ Level 6 reflective \hyperlink{tag:body-armor}{body armor}.
\item Sergeant: 3 Trans cash; Civ Level 7 impact \hyperlink{tag:body-armor}{body armor}.
\item Lieutenant: 10 Trans cash; 250 Mils per year pension; Civ Level
  7 reflect/impact \hyperlink{tag:body-armor}{body armor}.
\item Captain: 25 Trans cash; 1 Tran per year pension; Civ Level 8
  reflective \hyperlink{tag:body-armor}{body armor}.
\item Colonel: 60 Trans cash; 3 Trans per year pension; Civ Level 8
  reflect/impact \hyperlink{tag:body-armor}{body armor} and \hyperlink{tag:arcgun}{arc gun}.
\end{enumerate}

\bigskip

\pagebreak[2]

\noindent{\large\textbf{SPY}} \label{sec:prof-spy}\index{spy}
\stepcounter{myprofs}

A member of a federal or planetary intelligence agency.  Spies are
used throughout the federation to gather information and to conduct
secret missions on planets that the federation or rival planetary
governments are suspicious of. Spies usually work alone, under a
variety of identities.
  
\textbf{Prerequisites:} Study of the mind and the body; Characteristic
Ratings of at least Dexterity \textbf{7}, Agility \textbf{5}, and
Intelligence \textbf{6}; Urban Skill Level of at least \textbf{1}.

\textbf{Skill Point Modifier:} 7.

\textbf{Skills Available:} Demolitions, longarms, handguns, paint gun,
arc gun, bows, linguistics, disguise,
forgery/\linebreak[0]counterfeiting, weapon 
tech, urban.  

\pagebreak[2]
\textbf{Benefits:}\\
All spies receive a set of \hyperlink{tag:false-identity}{false identity papers}.
\begin{enumerate}
\item Nothing additional. 
\item Civ Level 6 \hyperlink{tag:plastic-pistol}{plastic pistol}. 
\item Any one handgun or laser pistol.   
\item 700 Mils cash; \hyperlink{tag:cl8-laser}{Civ Level 8 laser pistol} and any other handgun. 
\item 4 Trans cash; \hyperlink{tag:cl8-paintgun}{Civ Level 8 paint gun} and any one  handgun. 
\item 20 Trans cash; \hyperlink{tag:arcgun}{arc gun} and any one handgun.  
\end{enumerate}

\bigskip

\pagebreak[2]

\noindent{\large\textbf{STAR SAILOR
    (Military)}}\label{sec:prof-star-sailor}\index{star sailor}
\stepcounter{myprofs}

A member of the federal navy, serving aboard a spaceship.
Responsibilities may range from plotting courses to piloting a craft
to manning the guns during battle. A star sailor is familiar with all
manner of known space craft and is more comfortable in a zero-G
environment than he is on the ground.

\textbf{Prerequisites:} Study of the military and either applied
sciences or the body; Characteristic Ratings of at least Agility
\textbf{5}, Intelligence \textbf{7}, and Mental Power \textbf{2}. The
character's home gravity type may not be \emph{extreme}. The mental
power waiver applies (see \ref{sec:mental-power-waiver}).

\textbf{Skill Point Modifier:} 6.

\textbf{Skills Available:} Near Weightless Gravity Skill Level is
\emph{automatically} raised to \textbf{0} (if not already equal to or
greater than \textbf{0}) at no cost in Skill Points. Pilot, astronomy,
energy tech, suit tech, space ship tech (\textbf{1} Level only),
weapon tech, space tactics, gunnery, missile guidance, EVA, near
weightless gravity (in addition to any levels received as described
above).

\pagebreak[2]
\textbf{Benefits:}\\
All star sailors receive an \hyperlink{tag:gravity-web}{internal gravity web}
and free passage on any interstellar federal craft (for the distance
listed in Light Years).
\begin{enumerate}
\item Midshipman: 500 Mils cash; no passage.  
\item Ensign: 1.5 Mils cash; 10 LY passage.  
\item Lieutenant: 4 Trans cash; 300 Mils per year pension; 20  LY passage.  
\item Captain: 15 Trans cash; 1 Tran per  year pension; 50 LY passage.  
\item Commander: 36 Trans cash; 3 Trans per year pension; 100 LY passage.   
\item Admiral: 80 Trans cash; 10 Trans per year pension;  perpetual passage.  
\end{enumerate}

\bigskip

\pagebreak[2]

\noindent{\large\textbf{THINKER}} \label{sec:prof-thinker}\index{thinker}
\stepcounter{myprofs}

A member of the exclusive psionic community.
The advanced study of the mind has turned telepathy, telekinesis, and
other extraor dinary mental abilities into the premier science of the
24th Century. Thinkers effectively control interstellar travel and
communications and are thus treated with great respect by government
and corporate officials. Because of the freak nature of their powers,
however, the general public greets them with a mixture of awe, fear,
and mistrust. Thinkers often set up small colonies in luxurious
isolation so that they may probe the mysteries of thought undisturbed.
A thinker will keep his powers concealed when with non-Psions.

\textbf{Prerequisites:} Study of the mind and either applied science
or the humanities; Characteristic Ratings of at least Intelligence
\textbf{7} and Mental Power \textbf{5}.

\textbf{Skill Point Modifier:} 10.

\textbf{Skills Available:}  Navigation, mind control, psychokinesis,
  psion tech, teaching.

\pagebreak[2]
\textbf{Benefits:}
\begin{enumerate}
\item Nothing. 
\item \hyperlink{tag:commlink}{interstellar commlink}.  
\item 1 Tran cash; \hyperlink{tag:commlink}{interstellar commlink}.  
\item 3 Trans cash; \hyperlink{tag:psionic-rig}{psionic rig}.
\item 10 Trans cash; \hyperlink{tag:psionic-rig}{psionic rig}.  
\item 25 Trans cash; \hyperlink{tag:augmented-jump}{augmented  jump pod}. 
\end{enumerate}

\bigskip

\pagebreak[2]

\noindent{\large\textbf{ZERO-G
    MINER}}\label{sec:prof-zero-g-miner}\index{zero-g miner}
\stepcounter{myprofs}
  
A prospector of ores and precious metals in asteroid belts and on
small planetoids. Miners work together in small groups, surveying the
entire ``ring'' of an asteroid belt, staking claims on those rocks
that appear promising. Actual mining operations range from a single
miner with a rock blaster to orbital mining cities.
\
textbf{Prerequisites:} Study of applied science or business;
Characteristic Ratings of at least Dexterity \textbf{5} and Agility
\textbf{3}. The character's initial wealth must be at least 500 Mils.
The character's home gravity type may not be \emph{extreme}.

\textbf{Skill Point Modifier:} 4.

\textbf{Skills Available:} Near Weightless Gravity Skill Level is
automatically increased to \textbf{0} (if not already equal to or
greater than \textbf{0}) at no cost in Skill Points. Pilot, economics,
trading, EVA, mining, asteroid mining, geology, electro tech, near
weightless gravity (in addition to any levels received as described
above).

\pagebreak[2]
\textbf{Benefits:}
\begin{enumerate}
\item Nothing.  
\item Initial wealth x 2.   
\item Initial wealth x 3; \hyperlink{tag:rock-blaster}{rock blaster}.  
\item Initial wealth x 5;  Civ Level 8 \hyperlink{tag:geoscanner}{geoscanner}.  
\item Initial wealth x 10; \hyperlink{tag:rock-blaster}{rock blaster} and Civ Level 8  \hyperlink{tag:geoscanner}{geoscanner}. 
\item Initial wealth x 20; \hyperlink{tag:bot-manner-44}{Manner 44 robot} with \hyperlink{tag:bot-force-field}{force field system},
  \hyperlink{tag:bot-geo}{geo system}, \hyperlink{tag:bot-pilot}{pilot system}, and \hyperlink{tag:bot-chemical}{chemical system}.
\end{enumerate}

\addtocounter{myprofs}{-1}
\refstepcounter{myprofs}\label{item:myprofs}
\index{character!generation|)}
%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "gm_guide"
%%% End: 
