%%
%% $Id: equipment.tex,v 1.2 2004/10/08 19:42:07 goltz20707 Exp $
%%
\chapter{Equipment}
\label{cha:equipment}

In order to adventure on the vast and varied worlds of known space,
the characters must take advantage of the technology of the
federation. Equipment is a loose term encompassing all weapons,
protective attire, vehicles, and scientific, technical, and personal
equipment the characters may purchase or receive. Most items of
equipment and their pertinent specifications are presented on the
equipment charts. Additional information about particular items and
other items not mentioned on the charts are included in the
appropriate Sections.

The hostile environments in which characters will often operate demand
the use of costly equipment. If characters do not receive what they
need from their professional benefits and do not have the funds to
purchase such things as an expedition suit or simple vehicle, the GM
should see that they are provided with those items necessary for
survival at the least (even if on loan from an employer).

\section{Economic Guidelines}
\label{sec:economic-guidelines}
 
The basic costs of every piece of equipment are listed on the
equipment charts. These prices may vary, depending on many factors.

If the price of an item is \textbf{1 Tran} or less, the basic price is used
when purchasing the item on a world that has the same (or a higher)
Civ Level as the item. If the item's Civ Level is one higher than the
world of purchase, the price is \emph{doubled} and the item will be
available only in urban areas (concentrated population of 200,000 or
more). If the item's Civ Level is two higher than that of the world of
purchase, the price is \emph{multiplied by} \textbf{10} and the item will be
available only in urban areas, if it is available at all. In such a
case, the GM rolls one die; if the result is \textbf{4} or less, the item is
available in that area. A separate check may be made for each distinct
urban area on the world.


If an item costs more than \textbf{1 Tran}, the \textbf{Actual Price
  Table} (Table \vref{tab:actual-price}) is used to determine the
price of the item on a given world. The table lists modifiers that are
applied to the percentile dice roll. These include a modifier based on
whether the item is \emph{exported}, \emph{manufactured}, or not
manufactured on the world (see the World Resources Table and Resource
Development, \ref{tab:world-resource}.  Each product falls into one of
these three categories on each world.  \textbf{Example:} The price of
a Translator (Resource Development CC) has a --30 modifier is
\emph{silicon} is \emph{abundant} and \emph{iron} is \emph{limited} on
the world, no modifier if silicon is \emph{limited}, and a +30
modifier if there is no silicon on the world.  Other modifiers take
the Civ Level of the world into account. A character may use his
\emph{trading skill} to modify the Actual Price Table dice results (see
\ref{sec:interpersonal-skills}). The general availability of an item
costing more than \textbf{1 Tran} that is of a higher Civ Level than the world
of purchase is subject to the same restrictions as items costing less
than a Tran.

An item with a Civ Level of \emph{more than two} above the Civ Level
of a given world is \emph{illegal} on that world. If the item can be
found at all (at the GM's discretion), its cost is multiplied by 10
\emph{times the Law Level} of that world. A weapon that is classified
as \emph{military} is illegal (see \ref{sec:weapons}). A weapon that is
classified as \emph{restricted} is illegal unless the purchaser has
received a permit from the authorities (if he was going to an area
populated with dangerous creatures, for example). The GM may make an
illegal weapon available to the characters, but if he does so, he
should multiply its price by the Law Level of the world (and should
also apply modifiers based on the Civ Level and the other variables
discussed above).

Even though robots and spaceships are not included in this Chapter,
their price is affected by all these variables. The GM should always
use the Actual Price Table and these modifiers for a character that is
buying or selling large amounts of goods or resources (such as a
merchant or interstellar trader). A character may use his
\emph{economics} skill to determine the price of items or resources
that he wishes to buy or sell on other worlds at a future time (see
\ref{sec:interpersonal-skills}). The base price for all types of
resources in bulk are listed on the World Resource Table.

The basic costs of living are not discussed in detail. As a guideline,
a week's room and board of mediocre quality in a spaceport or city
costs \textbf{300 Mils} (\textbf{50 Mils} per day). A day's travel
rations (tang and spacefood sticks) costs \textbf{5 Mils}. A
seven-course meal in a first class restaurant costs at least
\textbf{100 Mils}.

\begin{table}[htbp]
  \centering
  \fbox{%
    \begin{minipage}{0.95\textwidth}
      \centering
      \caption{Actual Price Table}
      \label{tab:actual-price}

      \medskip

      \parbox{2.5in}{\begin{tabular}{cc}
          Percentile & Actual Price of Goods\\
          Dice Result & (Compared to listed price) \\
          0 or less & 50\% \\
          \rowcolor{grey}
          1-10 & 65\% \\
          11-20 & 75\% \\
          \rowcolor{grey}
          21-30 & 85\% \\
          31-40 & 90\% \\
          \rowcolor{grey}
          41-50 & 95\% \\
          51-60 & 100\% \\
          \rowcolor{grey}
          61-70 & 110\% \\
          71-80 & 125\% \\
          \rowcolor{grey}
          81-90 & 145\% \\
          91-100 & 170\% \\
          \rowcolor{grey}
          101 or more & 200\%
        \end{tabular}}
      \parbox{3in}{\noindent MODIFIERS: \\
        \begin{itemize}
        \item[--30] if good is abundant on world or is exported from world.
        \item[+30] if good does no exist or is not manufactured on
          world.  (See the World Resource Table, pages
          \ref{tab:world-resource} and
          \ref{tab:world-resource-explanation}, for explanation of
          above modifiers.)
        \item[+50] if item is one Civ Level higher than that of the world. 
        \item Multiply percentage by \textbf{10} if item is two Civ Levels higher 
          that that of the world. 
        \item Multiply percentage by the Law Level of the world if the 
          item is illegal (or restricted, if the purchaser does not 
          have a permit). 
        \item Variable modifier may be applied when using trading 
          skill. 
        \end{itemize}}
    \end{minipage}}
\end{table}

\section{Weapons}
\label{sec:weapons}

Personal weapons range from a simple dagger to the arc gun, a
duo-laser beam military assault weapon. The federation prohibits the
ownership of military weapons by the general public. Other weapons are
available with permit, for scientific use and personal protection (if
cause can be shown to authorities). Archaic personal weapons may be
owned by anyone. The following list of weapons is divided into three
categories (open, restricted, and military). All information necessary
for use of each weapon is provided on the \textbf{Weapon Chart} (Table
\vref{tab:weapon}).  Some of this information includes:

\begin{enumerate}
\item \textbf{Rate of Fire.} The maximum number of fires that may be conducted
  per Action Round (see \ref{sec:firing-weapons}). Also represents the number
  of fires in a single clip (one load) for the weapon.
\item \textbf{Hit Strength.} The force with which a single fire from the weapon
  strikes a target.
\item \textbf{Base Hit Chance.} The percentage chance of a single fire from the
  weapon hitting a target. This chance varies depending on how many
  hexes away the target is (\emph{range in hexes}). A hex equals \textbf{5
  meters}.  The Terrain Value of the area is multiplied by the
  \emph{Terrain Multiplier} (which varies with the range); this
  product also affects the Base Hit Chance. See also \ref{sec:fire-modifier-chart}.
\end{enumerate}

No cost or availability is given for ammunition used in these weapons.
Aside from the time required to reload during Action Rounds (see
\ref{sec:firing-weapons}), it is assumed that a character with a
weapon has plenty of ammunition. If the GM wishes to keep track of
ammunition, he should have the players note the expenditure of one
clip each time they fire a number of shots equal to the weapon's
\emph{Fire Rate} (cumulative from Action Round to Action Round). The
GM is responsible for determining the price, weight, and availability
of each type of clip if he is keeping track of ammunition.

\medskip

\noindent\textbf{OPEN}

Musket, short bow, long bow, crossbow, dagger, and sword. 


\medskip

\noindent\textbf{RESTRICTED}
\begin{description}
\item[\hypertarget{tag:cl6-paintgun}{\textbf{Paint Gun} (Civ Level 6)}.] A rifle-shaped laser weapon using a
  low intensity laser as an aiming device. Once the target is spotted,
  the intensity is increased. This version is very bulky due to the
  need for a power pack. 
\item[\hypertarget{tag:cl6-laser}{\textbf{Laser Pistol} (Civ Level 6)}.] A single shot laser weapon.
  Requires reloading after each shot. 
\item[\textbf{Stun Pistol} (Civ Level 7 \& 8).] A pistol shaped weapon emitting
  a subsonic pulse temporarily deadening the target's nervous system,
  stunning it (see \ref{sec:stun-pistol}). Does not work in a vacuum. 
\item[\textbf{Needle Rifle} (Civ Level 5 \& 7).] A longarm that fires needles;
  the needles do no damage, but any substance they contain may (see
  \ref{sec:toxin-effects-matrix}).
\end{description}

Also rifles, smoke grenades, illumination grenades, and gas grenades.

\medskip

\noindent\textbf{MILITARY}
\begin{description}
\item[\hypertarget{tag:cl8-paintgun}{\textbf{Paint Gun} (Civ Level 8)}.] Same as Civ Level 6 version except
  with improved rate of fire and much less bulky power source. 
\item[\hypertarget{tag:arcgun}{\textbf{Arc Gun} (Civ Level 8)}.] A rifle shaped twin beam weapon using a
  sighting mechanism similar to a single lens reflex camera. The
  energy beams are harmless except at the point of intersection, which
  is adjusted by a slider on the barrel. The beams are not activated
  until the trigger is depressed. Ideal weapon for fighting inside
  spaceships due to its exact targeting. Requires a great deal of
  skill to use. 
\item[\textbf{\hypertarget{tag:plastic-pistol}{Plastic Pistol}} (Civ Level 6).] Made of a high-density synthetic
  resin, this weapon is constructed entirely of non-metal parts and
  will not register on a metal detector if unloaded.
\item[\textbf{Needle Pistol} (Civ Level 8).] Similar to the needle rifle, but
  in a pistol configuration.
\item[\hypertarget{tag:cl8-laser}{\textbf{Laser Pistol} (Civ Level 8)}.] These power packs will allow three
  fires.
\item[\textbf{Fragmentation Grenade} (Civ Level 4).] The hit strength is
  applied to the target hex and every surrounding hex.
\end{description}

Also pistols, machine guns, submachine guns, and carbines.

\begin{table}[htbp]
  \small
  \centering
  \fbox{%
    \begin{minipage}{0.95\textwidth}
      \centering
      \caption{Weapon Chart}
      \label{tab:weapon}

      \medskip

%    \renewcommand{\thefootnote}{\alph{footnote}}
\label{tab:weapon-change-table}
\begin{tabular}{ccccccclccccc}
      \rotate{CIV LEVEL} & \rotate{AVAILABLILITY} & \rotate{WEIGHT IN
        KG} & \rotate{\parbox{1in}{PRICE\\(TRANS)}} & \rotate{RATE OF FIRE} &
      \rotate{HIT STRENGTH} & \rotate{\parbox{1in}{BASE REPAIR\\TIME (HOURS)}} &
      TERRAIN MULTIPLIER &  0 & 2 & 4 & 6 & 8 \\
      &&&&&&& RANGE IN HEXES & 0 & 1--4 & 5--13 & 14--25 & 26--40\\ 
\rowcolor{grey}
6 & R & 12 & 7 & 2 & 8 & 1 & Paint Gun & 35 & 25 & 10 & --10 & --25\\
8 & M & 3 & 9 & 6 & 8 & 2 & Paint Gun & 35 & 25 & 10 & --10 & --25\\
\rowcolor{grey}
8 & M & 2 & 12 & 4 & 10 & 2 & Arc Gun & 20 & 15 & 5 & --10 & --25\\
5 & M & 2 & 0.08 & 4 & 5 & 1 & Pistol \footnotemark[10] & 60 & 40 & 0 & P & P\\
\rowcolor{grey}
5 & M & 1 & 0.2 & 4 & 5 & 1 & Plastic Pistol \footnotemark[9]{\tiny $^,$}\footnotemark[10] & 60 & 40 & 0 & P & P\\
8 & M & 2 & 2 & 2 & 1 & 2 & Needle Pistol \footnotemark[3]{\tiny $^,$}\footnotemark[10] & 55 & 35 & --10 & P & P\\
\rowcolor{grey}
6 & R & 4 & 1 & 1 & 6 & 1 & Laser Pistol & 60 & 40 & 20 & 0 & --20\\
8 & M & 2 & 2 & 3 & 6 & 2 & Laser Pistol & 60 & 40 & 20 & 0 & --20\\
\rowcolor{grey}
7 & R & 2 & 1 & 1 & 8 & 2 & Stun Pistol \footnotemark[4] & 65 & 30 & P & P & P\\
8 & R & 1 & 2 & 2 & 9 & 2 & Stun Pistol \footnotemark[4] & 65 & 30 & P & P & P\\
\rowcolor{grey}
5 & M & 12 & 3 & 15 & 7 & 1 & Machine Gun \footnotemark[1] & P & 40 & 20 & 0 & --20\\
5 & M & 5 & 2 & 10 & 5 & 1 & Submachine Gun\footnotemark[10] & 40 & 35 & 10 & 0 & P\\
\rowcolor{grey}
2 & O & 5 & 0.3 & 2 & 4 & 1 & Musket \footnotemark[10] & P & 0 & --20 & --40 & P\\
5 & R & 3 & 0.8 & 6 & 7 & 1 & Rifle\footnotemark[10] & 30 & 35 & 20 & 0 & --15\\
\rowcolor{grey}
5 & M & 3 & 0.6 & 4 & 6 & 1 & Carbine \footnotemark[10] & 45 & 35 & 15 & --5 & --25\\
5 & R & 5 & 2 & 1 & 1 & 1 & Needle Rifle \footnotemark[3]{\tiny $^,$}\footnotemark[10] & 40 & 30 & 10 & --20 & P\\
\rowcolor{grey}
7 & R & 3 & 3 & 2 & 1 & 2 & Needle Rifle \footnotemark[3]{\tiny $^,$}\footnotemark[10] & 40 & 30 & 10 & --20 & P\\
1 & O & 1 & 0.2 & 3 & 4 & 1 & Short Bow \footnotemark[2]{\tiny $^,$}\footnotemark[10] & 15 & 25 & 0 \footnotemark[10] & P & P\\
\rowcolor{grey}
1 & O & 3 & 0.6 & 2 & 6 & 1 & Long Bow \footnotemark[2]{\tiny $^,$}\footnotemark[10] & P & 20 & 10 & 5 \footnotemark[10] & --55 \footnotemark[10]\\
3 & O & 4 & 0.5 & 1 & 1 & 1 & Crossbow & 0 & 30 & 10 & --15 & P\\
\rowcolor{grey}
4 & M & 0.5 & 0.1 & 1 & 25 & 1 & Fragmentation Grenade & P & 50 & --10 & --60 \footnotemark[10] & P\\
4 & R & 0.5 & 0.1 & 1 & 4 & 1 & Smoke Grenade \footnotemark[7] & P & 50 & --10 & --60 \footnotemark[10] & P\\
\rowcolor{grey}
4 & R & 0.5 & 0.1 & 1 & 4 & 1 & Illumination Grenade \footnotemark[8] & P & 50 & --10 & --60 \footnotemark[10] & P\\
4 & R & 0.5 & 0.1 & 1 & 4 & 1 & Gas Grenade \footnotemark[5] & P & 50 & --10 &
      --60 \footnotemark[10] & P\\
\rowcolor{grey}
1 & O & 0.2 & 0.01 & 1 & 6 & 1 & Dagger\footnotemark[6] & P & 10 & --30 \footnotemark[10] & P & P\\
1 & O & 2 & 0.3 & na & 10 & 1 & Sword \footnotemark[6] & P & P & P & P & P\\
  \end{tabular}

  \medskip

  \parbox{4in}{NOTES: AVAILABILITY: M: Military, R: Restricted, O: Open. \\
  P: Prohibited.}
\renewcommand{\thempfootnote}{\arabic{mpfootnote}}
\footnotetext[1]{Must be emplaced before first fire.}
\footnotetext[2]{May not be fired by a prone character.}
\footnotetext[3]{The target may be 
affected by substance in needle; see \ref{sec:toxin-effects-matrix}.}
\footnotetext[4]{The strength of weapon used to check for stun only; see 
\ref{sec:stun-pistol} (Terrain Value considered \textbf{0}).}
\footnotetext[5]{Target hex and all adjacent hexes are filled with gas for six Action 
Rounds; all beings there which do not have their own air supply suffer the effects of the appropriate 
poison (See \ref{sec:toxin-effects-matrix}).}
\footnotetext[6]{Hit Strength is used in close combat; Hit Strength is \textbf{3} if thrown.}
\footnotetext[7]{Target hex and 
all adjacent hexes are treated as if it were dark for six Action Rounds.}
\footnotetext[8]{Target hex and all adjacent 
hexes are treated as if it were light for six Action Rounds.}
\footnotetext[9]{Will not be detected by metal detector.}
\footnotetext[10]{Weapon produces recoil with each shot (See \ref{sec:fire-modifier-chart}).}
\footnotetext[11]{Strength Rating of firing (or throwing) character 
added to hit percentage.}
\end{minipage}}
\end{table}


\section{Protective Equipment}
\label{sec:protective-equipment}

The Protective Attire Chart (see page \pageref{tab:protective-attire})
summarizes the attributes of 
suits, body armor, and other devices the characters may wear to
protect themselves when in a hostile environment and/or when involved
in combat. Some of the attributes listed on this chart are:

\begin{itemize}
\item \textbf{Price.} Certain prices include a waiting period between
  the time the item is purchased and the time it is ready for use. The
  item is being contoured to fit the wearer's body during this time.
\item \textbf{Air Supply.} After the listed time, the oxygen supply
  must be replenished. This may be done from any vehicle or structure
  with its own oxygen supply. Oxygen produced by a Chemsynthesizer may
  replenish the supply at a rate of \textbf{1} hour's oxygen for every
  \textbf{10} minutes of use.
\item \textbf{Encumbrance.} A quantification of how much the item
  impedes the wearer's movement; the higher the rating, the slower the
  wearer may move. An \emph{Encumbrance Rating} may be overcome by the
  wearer's EVA or Body Armor Skill Level. Certain types of body armor
  are augmented; the outfit is powered and may increase the strength
  and mobility of a wearer with body armor skill. The effects of
  encumbrance and augmentation are detailed in
  \ref{sec:foot-travel}, \ref{sec:carrying-equipment}, and \ref{sec:action-round-movement}.
\item \textbf{Projectile and Beam Armor Defense Ratings.} Used in
  combat to defend against enemy fire, as explained in
  \ref{sec:armor-defense}. The Projectile Defense Rating may also protect the
  wearer from blows and accidental impact.
\item \textbf{Hit Strength.} Used to attack or defend in close combat
  (see \ref{sec:close-combat}).
\end{itemize}



The following material features additional information about the items
on the chart and other items not listed.

\begin{description}
\item[\textbf{\hypertarget{tag:armor-vest}{Armor Vest}.}] Protects the chest, trunk, and groin like body
  armor. When wearer is hit, check \emph{Hit Table} die roll. If the
  \emph{unmodified} roll is \textbf{4}, \textbf{5}, \textbf{6}, or
  \textbf{7}, the Armor Defense Ratings of the vest are used.
\item[\textbf{Battle Sleeve.}] Protects and increases the strength of the
  wearer's hand, arm, and shoulder.  Its Hit Strength is used in close
  combat like body armor.  When wearer is hit by fire, check the
  \emph{Hit Table} die roll. If the \emph{unmodified} roll is
  \textbf{1} or \textbf{2}, the Armor Defense Ratings of the sleeve
  are used.
\item[\textbf{\hypertarget{tag:body-armor}{Body Armor}.}] Suit constructed of high-impact metals and
  ceramics. Includes all attributes of an expedition suit (see
  following). May withstand most corrosive atmospheres. Various
  production methods make some body armor more effective against
  projectiles (impact armor) and others more effective against beams
  (reflective armor). Much practice is required to use augmented
  armor, which electronically accentuates all movement by the wearer.
\item[\hypertarget{tag:expedsuit}{Expedition Suit}.] Provides complete protection from a
  vacuum, water, or contaminated or poison atmosphere. Contains
  two-way headset radio in helmet. Audio system allows wearer to hear
  and speak normally.
\item[Filter Mask.] Filter covering the mouth and nose, providing
  complete protection from contaminated atmospheres. Must be replaced
  after 10 days' use. Civ Level: \textbf{5}. Weight: Negligible. Price:
  \textbf{40 Mils}.
\item[Force Field.] Wire rig worn about the shoulders. When activated,
  an energy field surrounds the wearer providing excellent protection
  from projectiles and any other solid objects. Does not stop light.
  Wearer may not attack or be attacked in close combat. Energy pack
  allows one hour of continuous protection. Pack costs \textbf{1 Tran}
  to replace.
\item[\hypertarget{tag:respirator}{Respirator}.] Filter mask connected to oxygen tank (worn on the
  back). Allows character to function underwater, in a contaminated
  atmosphere, and in some poisonous atmospheres. Civ Level:
  \textbf{6}.  Weight: \textbf{3 kg}. Price: \textbf{1 Tran}. Air
  Supply: \textbf{12} hours.
\item[\hypertarget{tag:respirator-helmet}{Respirator Helmet}.] Covers the head, shoulders, and upper chest.
  Includes all attributes of a respirator (see preceding).  Contains
  two-way headset radio. Audio system allows wearer to hear and speak.
  When wearer is hit by fire, check \emph{Hit Table} die roll. If the
  \emph{unmodified} roll is an \textbf{8}, \textbf{9}, or \textbf{10},
  the helmet's Armor Defense Ratings are used.
\end{description}

\begin{table}[htbp]
  \centering
  \fbox{%
    \begin{minipage}{5.25in}
      \centering
      \caption{Protective Attire Chart}
      \label{tab:protective-attire}

      \medskip

      \begin{tabular}{lcrrrrrrrr}
        TYPE & \rotate{CIV LEVEL} & \rotate{WEIGHT (kg)} & \rotate{\parbox{1in}{PRICE\\(TRANS)}} & \rotate{\parbox{1in}{AIR SUPPLY\\(Hours)}} & \rotate{ENCUMBRANCE} &
        \rotate{\parbox{1.5in}{PROJECTILE\\DEFENSE}} & \rotate{\parbox{1.5in}{BEAM\\DEFENSE}} & \rotate{\parbox{1.5in}{BASE
          REPAIR\\TIME (HOURS)}} & \rotate{HIT STRENGTH}\\
        ARMOR\\
        \rowcolor{grey}
        Reflective & 6 & 11 & 15w & 12 & 2 & 1 & 3 & 2 & 2\\
        Reflective & 8 & 9 & 28w & 24 & 1 & 2 & 5 & 2 & 5\\
        \rowcolor{grey}
        Impact & 5 & 22 & 9 & None & 5 & 2 & 1 & 2 & 9\\
        Armor Vest & 6 & 2 & 1.5 & None & None & 2p & 2p & 2 & None\\
        \rowcolor{grey}
        Battle Sleeve & 8 & 4 & 15w & None & None & 5p & 5p & 2 & 25\\
        \\
        POWERED ARMOR\\
        \rowcolor{grey}
        Impact & 7 & 18 & 31w & 18 & 3a & 4 & 2 & 3 & 18\\
        Reflective / Impact & 7 & 20 & 31w & 18 & 4a & 3 & 4 & 3 & 11\\
        \rowcolor{grey}
        Reflective / Impact & 8 & 16 & 46w & 24 & 2a & 5 & 5 & 3 & 20\\
        \\
        RESPIRATORS\\
        \rowcolor{grey}
        Respirator & 6 & 3 & 1 & 12 & None & 0 & 0 & 1 & None\\
        Respirator Helmet & 6 & 6 & 4 & 12 & None & 2p & 2p & 1 & None\\
        \rowcolor{grey}
        Respirator Helmet & 8 & 4 & 10 & 24 & None & 4p & 4p & 1 & None\\
        Filter Mask & 5 & na & 0.04 & 0 & None & 0 & 0 & 1 & None\\
        \\
        FORCE FIELD\\
        \rowcolor{grey}
        Force Field On & 8 & 2 & 55 & None & Imm & 6 & 0 & 2 & Imm\\
        Force Field Off & 8 & 2 & 55 & None & None & 0 & 0 & 2 & None\\
        \\
        EXPEDITION SUIT\\
        \rowcolor{grey}
        Expedition Suit & 6 & 11 & 6 & 12 & 2 & 1 & 1 & 2 & None\\
        Expedition Suit & 7 & 7 & 11w & 24 & 1 & 1 & 1 & 2 & None\\
        \rowcolor{grey}
        Expedition Suit & 8 & 5 & 11w & 30 & 1 & 1 & 1 & 2 & None\\
      \end{tabular}
     
      \medskip

      \parbox{\textwidth}{\begin{description}
        \item[w:] Waiting period of one week required upon purchase.
          \emph{Exception:} Waiting period for Battle Sleeve is two days.
        \item[a:] Movement of wearer may be augmented, depending on Body
          Armor Skill Level.
        \item[p:] Provides partial protection only.
        \item[Imm:] Wearer may not move at all. He may not attack or fire
          any type of weapon except a beam weapon.
        \end{description}}
    \end{minipage}}
\end{table}

\section{Vehicles}\label{sec:vehicles}

In addition to traveling between worlds and stars, characters will
need to traverse the surface of worlds. On land, above or beneath the
surface, vehicles are needed for transportation. The vehicles listed
on the charts and those detailed in this Section do not represent all
the vehicles of the universe. They are intended as a guide from which
the GM should borrow to invent others.

The vehicle charts detail important data for each individual vehicle
type, including the following information:

\begin{itemize}
\item \textbf{Range in 100 km.} Maximum range vehicle can travel
  before refueling. Refueling costs: (Range in \textbf{100 km}) $\times$
  Mils.
\item \textbf{Performance Modifier.} A quantification of the vehicle's
  responsiveness and structural integrity. Applied to the chance to
  avoid an accident, in accordance with \ref{sec:vehaccidents}.
\item \textbf{Repair Time in Days (or Hours).} Base time required to
  repair the vehicle (see \ref{sec:repair}).
\end{itemize}

Every vehicle Civ Level 6 or higher which carries more than one person
comes equipped with a planetary short wave radio. All vehicles with an
air lock come with an oxygen supply equal to twice the vehicle's Civ
Level in days. An airlock holds one or two people at a time (GM's
discretion). Civ Level 8 vehicles may, at the GM's discretion, come
equipped with an oxygen synthesizer which will recycle breathable air
for a period of three months, at which time the system must be cleaned
and recharged at a cost of \textbf{1 Tran}.

\subsection[Land Vehicles]{Land vehicles are used to traverse the
  solid surface of a world.}
\label{sec:land-vehicles}

 The \textbf{Land Vehicle Chart} (page \pageref{tab:land-vehicle}) lists
various land vehicle types. Listed for each is a \emph{Terrain Value Limit}
and a \emph{Terrain Value Modifier}. The Terrain Value Limit indicates the
maximum Terrain Value the vehicle may traverse (see Terrain Effects
Chart, table \vref{tab:terrain-effects}). The Terrain Value Modifier
affects the actual speed with which 
a vehicle may travel through an allowable terrain. To calculate this
speed, add the Terrain Value Modifier to the Terrain Value of the
environ. If the sum is less than \textbf{1}, treat it as \textbf{1}. Divide the listed
speed in km/hour or hexes per Action Round by this sum (round down).
This is the maximum number of km or hexes the vehicle may move in that
terrain. Treat a road as having a Terrain Value of \textbf{1}; a trail as
having a Terrain Value of \textbf{2} (when in a higher Terrain Value). The car,
horse cart, and tractor all are similar to vehicles the players and GM
are familiar with. The others are detailed below:

\begin{description}
\item[Scout Car.]  A small non-enclosed vehicle using eight
  multi-celled tires for movement; petroleum powered. No air lock.
\item[Crawler.] An enclosed ATV using either balloon tires or
  telescoping legs for movement; nuclear powered. Has an air lock.
  Three Action Rounds (45 seconds) are required to change from wheels
  to legs.
\item[Rover.] A non-enclosed ATV with balloon tires and dynamic
  suspension for ease of movement; nuclear powered. No air lock.
\item[Mobile Lab.] A three-sectioned ATV with normal tires for
  movement. Each section is independent of the others and connected
  (like train cars) by accordion bags. The first section is for the
  driver and crew; the second is the lab; the third is for holding
  creatures and can adapt its atmosphere to match those suitable to
  its occupants. The whole lab is enclosed and air tight; nuclear
  powered.  An air lock connects sections 2 and 3; another connects
  section 1 with the outside.
\item[Amphibian.] Truly an all-terrain vehicle, this ATV can travel on
  land, on the surface of water, or submerged. Uses semi-balloon tires
  for land movement and is propeller-driven in water. Enclosed and
  airtight; has an air lock. Applicable driver's skill depends on mode
  of operation. Maximum depth in water of \textbf{75} meters.
\item[Van Rig.] A large truck with sleeping quarters for the driving
  team; designed for continuous travel. Enclosed and air tight;
  nuclear powered. Has an air lock.
\item[AutoSled.] Uses skis and treads to propel it through snow and
  over ice; petroleum powered and not enclosed.
\item[Half-Track.] Uses treads in rear and wheels in front; not
  enclosed. Petroleum powered.
\item[Loricate.] An armored personnel carrier using treads for
  movement. Enclosed, air tight, and nuclear powered. Has an air lock.
\item[Armored ATV.] An armored ATV using treads for movement; armed
  with a machine gun. Enclosed, air tight, and petroleum powered. Has
  an air lock.
\item[Tank.] Similar to tanks the players know except uses
  sophisticated laser aiming to achieve accurate fire of either
  anti-personnel or armor-piercing shells. Petroleum powered, not air
  tight.
\item[Self-Propelled Artillery.] An artillery gun on treads; not
  enclosed; petroleum powered.
\end{description}

\begin{table}[htbp]
  \scriptsize
  \centering
  \fbox{%
    \begin{minipage}{0.95\textwidth}
      \centering
      \caption{Land Vehicle Chart}
      \label{tab:land-vehicle}

      \medskip

      \begin{tabular}{lcrrrrrrrrrrrr}
        VEHICLE & \rotate{CIV LEVEL} &
        \rotate{\parbox{1.5in}{PRICE\\(TRANS)}} & \rotate{\parbox{1.5in}{RANGE\\(100 x Kilometers)}} &
        \rotate{\parbox{1in}{SPEED\\(km/h)}} & \rotate{\parbox{1in}{SPEED\\(Hexes/AR)}} & \rotate{\parbox{1.5in}{TERRAIN
          VALUE\\LIMIT}} & \rotate{\parbox{1.5in}{TERRAIN VALUE\\MODIFIER}} & \rotate{PASSENGERS}
        & \rotate{\parbox{1.5in}{CARGO\\CAPACITY\\($\times 100$ kg)}} & \rotate{\parbox{1.5in}{PERFORMANCE\\MODIFIER}} &
        \rotate{\parbox{1.5in}{BASE REPAIR\\TIME (Hours)}} & \rotate{\parbox{1.5in}{ARMOR\\(Projectile/Beam)}} &
        \rotate{\parbox{1.5in}{LENGTH\\(Meters)}}\\
        {\footnotesize ASSOCIATED SUB-SKILL}\\
        \rowcolor{grey}
        Scout Car&&&&&&&&&&&&&\\
        \rowcolor{grey}
        {\footnotesize ATV} & 5 & 35 & 12.5 & 25 & 15 & 3 & 0 & 5 & 4 &
        --10 & 6 & 2/2 & 5\\
        \hypertarget{tag:crawler}{Crawler}&&&&&&&&&&&&&\\
        {\footnotesize ATV}\\Wheeled & 
        8 & 200 & 200 & 45 & 25 & 3 & --1 & 8 & 5 & +25 & 24 & 4/3 & 7\\
        \rowcolor{grey}
        Legs & 8 & 200 & 100 & 10 & 5 & N & N & 8 & 5 & +25 & 24 & 4/3 & 7\\
        \hypertarget{tag:rover}{Rover} & 8 & 150 & 300 & 75 & 45 & 4 & --2 & 2 & 1 & +25 & 24 & 3/3 & 5 \\
        \rowcolor{grey}
        Mobile Lab&&&&&&&&&&&&&\\
        \rowcolor{grey}
        {\footnotesize ATV} & 8 & 250 & 100 & 50 & 30 & 2 & +1 & 9 &
        25 & +15 & 24 & 4/3 & 10\\ 
        Amphibian\\
        {\footnotesize TRUCK} & 6 & 180 & 20 & 25 & 15 & 3 & +1 & 4 &
        3 & --15 & 24 & 3/2 & 5\\ 
        \rowcolor{grey}
        Van Rig&&&&&&&&&&&&&\\
        \rowcolor{grey}
        {\footnotesize TRUCK} & 6 & 240 & 24 & 100 & 65 & 1 & +2 & 5 &
        150 & --10 & 12 & 3/2 & 8 \\ 
        Car\\
%        {\hyperlink{tag:auto}{\footnotesize AUTOMOBILE} & 5 & 40 & 10 & 200 & 120 & 1.5 & 0 & 3 &
%        1 & --25 & 6 & 1/1 & 5 \\ 
        {\footnotesize AUTOMOBILE} & 5 & 40 & 10 & 200 & 120 & 1.5 & 0 & 3 &
        1 & --25 & 6 & 1/1 & 5 \\ 
        \rowcolor{grey}
        Horse Cart&&&&&&&&&&&&&\\
        \rowcolor{grey}
        {\footnotesize ANIMAL DRAWN} & 1 & 1 & U & 10a & 15a & 2 & +2 & 6 & 2 & --15 & 2 & 1/1 & 4 \\
        Auto Sled & 6 & 60 & 15 & 25 & 15 & 3 & 0 & 2 & 1 & 0 & 6 & 2/2 & 4\\
        \rowcolor{grey}
        Tractor & 5 & 35 & 5 & 15 & 10 & 4 & --1 & 1 & 100 & --20 & 6 & 1/1 & 5 \\
        Half-Track & 7 & 1000 & 50 & 25 & 15 & 2 & --2 & 8 & 4 & 0 & 24 & 4/4 & 7 \\
        \rowcolor{grey}
        Loricate&&&&&&&&&&&&&\\
        \rowcolor{grey}
        {\footnotesize APC} & 6 & 700 & 50 & 50 & 30 & 1.5 & +1 & 12 & 6 & +5 & 24 & 5/4 & 8\\
        Armored ATV & 6 & 200 & 15 & 75 & 45 & 3 & --1 & 6 & 1 & +15 & 24 & 5/5 & 5 \\
        \rowcolor{grey}
        Tank & 6 & 2000 & 50 & 50 & 30 & 3 & 0 & 4 & 2 & +20 & 24 & 6/6 & 6 \\
        Self-Propelled Artillery & 5 & 2000 & 10 & 25 & 15 & 2 & +1 & 3 & 1 & 0 & 24 & 6/5 & 7 
      \end{tabular}

      \medskip

      \parbox{\textwidth}{\textbf{a:} Speed in km/h represents a pace which
          can be  maintained 
          over a long period of time.\\
        \textbf{SPEED (Hexes / AR)} represents the maximum speed which can
          be maintained for a short period of time.\\
        \textbf{m:} See Marine Vehicle Chart for further details. \\
        \textbf{U:} Unlimited range. \\
        \textbf{N:} None.}
    \end{minipage}}
\end{table}

\subsection[Marine Vehicles]{Marine vehicles are used to travel
  either on or under the surface of water.}
\label{sec:marine-vehicles}

A sailing ship and all other types of 
%% Original:
%% wind or man powered water craft
%% New:
wind- or man-powered water craft\label{sec:marine-vehicles-hyphens-1}
%% End
are familiar to the players and the GM, and are among those summarized
on the \textbf{Marine Vehicle Chart} (see page
\pageref{tab:marine-vehicles}). Others include:

\begin{description}
\item[Mini Submarine.] A propeller driven submersible with a maximum
  depth of 100 meters. Petroleum powered; has an air lock.
\item[Flexi craft.] Traverses the surface with an eel-like motion;
  uses no propellers. Enclosed with an air lock; nuclear powered.
\item[Reef Walker.] Much as its name implies, this submersible has a
  set of telescoping legs for travel along the bottom of the ocean.
  Nuclear powered, it has a maximum depth of two kilometers.
\item[Water Scooter.] A propeller driven one-seater; not enclosed.
  Runs on electric batteries similar to a robot's.
\item[Amphibian.] See land vehicles (\ref{sec:land-vehicles}) for
  details on this vehicle.
\end{description}

\renewcommand{\thefootnote}{\Alph{footnote}}
\begin{table}[htbp]
  \centering
  \fbox{%
    \begin{minipage}{5.75in}
      \centering
      \caption{Marine Vehicle Chart}
      \label{tab:marine-vehicles}

      \medskip

      \begin{tabular}{lcrrrrrrrrr}
        VEHICLE &
        \rotate{CIV LEVEL} &
        \rotate{\parbox{1.5in}{PRICE\\(TRANS)}} &
        \rotate{\parbox{1.5in}{RANGE\\(km)}} &
        \rotate{\parbox{1.5in}{SPEED\\(km/h)}} &
        \rotate{\parbox{1.5in}{SPEED\\(Hexes/AR)}} &
        \rotate{PASSENGERS} &
        \rotate{\parbox{1.5in}{CARGO CAPACITY\\($\times 100$ kg)}} &
        \rotate{\parbox{1.5in}{BASE REPAIR\\TIME (Days)}} &
        \rotate{\parbox{1.5in}{ARMOR\\(Projectile/Beam)}} &
        \rotate{\parbox{1.5in}{LENGTH\\(Meters)}}\\
        {\footnotesize ASSOCIATED SUB-SKILL}\\
        \rowcolor{grey}
        Sailing Ship & 6 & 90 & U & 10 & 5 & 8 & 4 & 1 & 3/2 & 12\\ 
        Mini Submarine\\
        {\footnotesize SUBMARINE} & 8 & 300 & 15 & 25 & 15 & 3 & 1 & 1 & 3/2 & 4\\
        \rowcolor{grey}
        Flexicraft &&&&&&&&&&\\
        \rowcolor{grey}
        {\footnotesize MOTORBOAT} & 5 & 1000 & 150 & 100 & 60 & 20 & 50 & 6 & 2/4 & 40 \\
        Reef Walker \\
        {\footnotesize SUBMARINE} & 7 & 600 & 20 & 20 & 10 & 6 & 15 & 2 & 3/3 & 18 \\
        \rowcolor{grey}
        Water Scooter &&&&&&&&&&\\
        \rowcolor{grey}
        {\footnotesize MOTOR BOAT} & 7 & 5 & 1 & 15 & 10 & 1 & 0.2 & 1 & 1/1 & 2 \\
        \rowcolor{grey}
        {\footnotesize SUBMARINE} &&&&&&&&&&\\
        Amphibian\footnotemark[12]\\ 
        {\footnotesize MOTORBOAT} & 8 & 180 & 20 & 25 & 15 & 4 & 3 & 1 & 3/2 & 5\\
        {\footnotesize SUBMARINE}
      \end{tabular}

      \medskip

      \parbox{\textwidth}{\begin{description}
        \item[L:] See Land Vehicle Chart for further details. 
        \item[U:] Unlimited range. 
        \end{description}
        
        Where more than one associated sub-skill is listed, use the one
        appropriate to the current use of the vehicle.}
    \end{minipage}}
\end{table}

\subsection[Air Vehicles]{Air vehicles are used to fly above the
  surface of a world.}
\label{sec:air-vehicles}

All propeller driven and jet planes are familiar to the players and
the GM, and are among those summarized on the \textbf{Air Vehicle
  Chart} (page \pageref{tab:air-vehicles}). Others include:

\begin{description}
\item[Dragonfly.] A small man-powered craft with wings which are
  driven by the arms and legs; very light and fragile.  Weighs 20 kg.
\item[Ornithopter.] A large flying vessel using the actual beating of
  its wings for lift. Enclosed, air tight, and with an air lock, the
  example given is the largest and most powerful of its type. Nuclear
  powered.
\item[Air Car.] The common man's transportation; powered by an
  electrical turbine which uses the atmosphere as a propellant.
  Powered by batteries similar to a robot's.
\item[Levitator.] An anti-gravity craft using magnetic monopoles to
  repel the gravity field; maximum altitude is one kilometer. Enclosed
  with an air lock.
\item[Floater.] Similar to a levitator with a maximum altitude of 500
  meters.
\item[Skimmer.] Similar to levitator and floater except it has no air
  lock and cannot rise above 5 meters in altitude.
\item[Shuttle.] Similar to the shuttle of the 1980's except it needs
  no rocket boosters (lifts off like an airplane) and is nuclear
  powered.
%% New:
The example given is the largest of its
type.\label{sec:air-vehicles-change-largest}
%% End change
\item[\hypertarget{tag:jetpack}{Jetpack}.] Rocket powered backpack (solid fuel propellant)
  allowing wearer to fly in any atmosphere (or vacuum). Wearer uses
  jet-pack skill (not air vehicle skill). Civ Level: \textbf{7}. Cost:
  \textbf{75 Trans}. Range: \textbf{200 km}. Speed (km/hr):
  \textbf{200}. Repair Time: \textbf{12 hours}.  Weight: \textbf{40
    kg}.
\end{description}

\begin{table}[htbp]
  \footnotesize
  \centering
  \fbox{%
    \begin{minipage}{0.95\textwidth}
      \centering
      \caption{Air Vehicle Chart}
      \label{tab:air-vehicles}

      \medskip

      \begin{tabular}{lcrrrrrrrrrrrr}
        VEHICLE &
        \rotate{CIV LEVEL}  &
        \rotate{\parbox{1.5in}{PRICE\\(TRANS)}}  &
        \rotate{\parbox{1.5in}{RANGE\\($\times 100$ km)}}  &
        \rotate{\parbox{1.5in}{SPEED\\($\times 100$ km/h)}}  &
        \rotate{TRACE OR NONE*}  &
        \rotate{THIN*}  &
        \rotate{OTHERS*}  &
        \rotate{PASSENGERS}  &
        \rotate{\parbox{1.5in}{CARGO CAPACITY\\($\times 100$ kg)}}  &
        \rotate{\parbox{1.5in}{PERFORMANCE\\MODIFIER}}  &
        \rotate{\parbox{1.5in}{BASE REPAIR\\TIME (Days)}}  &
        \rotate{\parbox{1.5in}{ARMOR\\(Projectile/Beam)}}  &
        \rotate{\parbox{1.5in}{LENGTH\\(Meters)}} \\
        {\footnotesize ASSOCIATED SUB-SKILL}\\
        \rowcolor{grey}
        Dragonfly &&&&&&&&&&&&&\\
        \rowcolor{grey}
        {\footnotesize GLIDER} & 6 & 5 & 1 & 0.1 & P & 3 & 5 & 1 & .1 & --25 & 0.5 & 1/1 & 2 \\
        Ornithopter &&&&&&&&&&&&&\\
        {\footnotesize GLIDER} & 8 & 300 & 200 & 1.5 & P & 4 & 7 & 12 & 12 & 0 & 1 & 1/1 & 20 \\
        \rowcolor{grey}
        Propeller Plane & 5 & 200 & 40 & 5 & P & 3 & 7 & 5 & 25 & --5 & 0.5 & 2/2 & 8 \\
        Jet Plane & 7 & 1000 & 600 & 9 & P & 3 & 9 & 8 & 50 & +25 & 2 & 3/3 & 25 \\
        \rowcolor{grey}
        Air Car &&&&&&&&&&&&&\\
        \rowcolor{grey}
        {\footnotesize JET} & 8 & 1300 & 10 & 7 & P & 6 & 9 & 3 & 2 & +20 & 1 & 3/2 & 5 \\
        Levitator &&&&&&&&&&&&&\\
        {\footnotesize DIRECT LIFT} & 7 & 4000 & U & 0.3 & 6m & 6m & 6m & 100 & 100 & 0 & 10 & 4/4 & 40\\
        \rowcolor{grey}
        Floater &&&&&&&&&&&&&\\
        \rowcolor{grey}
        {\footnotesize DIRECT LIFT} & 8 & 400 & U & 0.4 & 6m & 6m & 6m & 10 & 5 & +25 & 6 & 1/1 & 8 \\
        Skimmer &&&&&&&&&&&&&\\
        {\footnotesize DIRECT LIFT} & 8 & 500 & U & 0.6 & 7m & 7m & 7m & 5 & 3.5 & +10 & 7 & 1/1 & 5 \\
        \rowcolor{grey}
        Helicopter & 6 & 150 & 20 & 1 & P & 3 & 6 & 6 & 5 & --20 & 1 & 2/3 & 5 \\
        Shuttle &&&&&&&&&&&&&\\
        {\footnotesize SHUTTLE or PILOT} & 7 & 8000 & 640 & 20 & 8 & 8 & 8 & 40 & 200 & +10 & 5 & 4/4 & 120 
      \end{tabular}

      \medskip

      \parbox{\textwidth}{\begin{description}
        \item[*:] MAXIMUM WORLD SIZE the Air Vehicle may in operate by
          ATMOSPHERE 
        \item[P:] Prohibited. 
        \item[U:] Unlimited Range; overhaul required once per year at a cost
          of 100 Credits. 
        \item[m:] Cannot be used on a world Size 3 or less. 
        \end{description}}
    \end{minipage}}
\end{table}

\section{Personal Equipment}
\label{sec:personal-equipment}

Representative personal scientific equipment, tech kits,
communications gear, and miscellaneous equipment is summarized on the
\textbf{Personal Equipment Charts} (pages
\pageref{tab:scientific-equipment}, \pageref{tab:repair-equipment},
\pageref{tab:comm-equip}, \pageref{tab:misc-equip}).\label{sec:personal-equipment-split-tables}

\subsection[Scientific Equipment]{Scientific equipment enables a
  character to undertake specific scientific tasks.}
\label{sec:scientific-equip}

Three kinds of equipment are available: \emph{labs}, which require a
piece of the item to be analyzed; \emph{scanners}, which allow remote
analysis up to a certain range; and \emph{synthesizers}, which require
input of raw material to produce usable substances.

\begin{description}
\item[Geo Lab.] Allows geology \emph{analysis} tasks only (see
  \ref{sec:skill-geology}). 
\item[\hypertarget{tag:geoscanner}{Geo Scanner}.] Allows any geology task; has a range of
  \textbf{50} times its Civ Level in kilometers. Will give direction
  of item being scanned if task is successful; gives distance only if
  item is within \textbf{50 km}.
\item[Mediscanner.] Allows diagnosis and treatment tasks. Must touch
  the being to be activated.
\item[\hypertarget{tag:bioscanner}{Bio Scanner}.] See biology skill, \ref{sec:skill-biology}.  Must
  be within \textbf{5} meters of the entity being scanned in order to
  operate.
\item[\hypertarget{tag:neuroscanner}{Neuroscanner}.] Can detect the electromagnetic waves emitted by
  any living thing within \textbf{300} meters. Will reveal the number
  and type of every living thing within range and also their general
  direction. If the population is very dense, this scanner is
  virtually useless. No skill required for use.  See \ref{sec:awareness}.
\item[\hypertarget{tag:chemlab}{Chem Lab}.] Allows chemistry \emph{analyze sample} tasks (see
  \ref{sec:skill-geology}).
\item[\hypertarget{tag:chemsynth}{Chem Synthesizer}.] Allows chemistry \emph{synthesis} tasks.
  Amount synthesized will be small (one kilo for simple compounds, one
  gram for complex compounds). Synthesis of a gas (oxygen, for
  example) will yield abundant amounts.
\item[Energy Scanner.] Used for all physics tasks except energy
  tapping. Range is \textbf{100} meters.
\end{description}


\begin{table}[htbp]
  \centering
  \fbox{%
    \begin{minipage}{4in}
      \centering
      \caption{Scientific Equipment Chart}
      \label{tab:scientific-equipment}
      
      \medskip
      
      \begin{tabular}{lcrrrrr}
        SCIENTIFIC &
        \rotate{CIV LEVEL} &
        \rotate{\parbox{1.5in}{WEIGHT\\(Kilograms)}} &
        \rotate{\parbox{1.5in}{PRICE\\(TRANS)}} &
        \rotate{SKILL BONUS} &
        \rotate{\parbox{1.5in}{TIME NEEDED\\FOR USE}} &
        \rotate{\parbox{1.5in}{BASE REPAIR\\TIME (Hours)}}\\
        EQUIPMENT &&&&&&\\
        \rowcolor{grey}
        Geo Lab & 5 & 15 & 1 & 0 & 2 & 1\\
        Geoscanner I & 6 & 25 & 6 & 0 & 1 & 1\\
        \rowcolor{grey}
        Geoscanner II & 8 & 7 & 15 & +2 & 0.5 & 2\\
        \hypertarget{tag:first-aid-kit}{First Aid Kit} & 4 & 5 & 1 & 0 & 1 & 1\\
        \rowcolor{grey}
        \hypertarget{tag:cl6-mediscanner}{Mediscanner I} & 6 & 10 & 12 & +1 & 0.5 & 1\\
        \hypertarget{tag:cl8-mediscanner}{Mediscanner II} & 8 & 5 & 10 & +2 & 0.3 & 2\\
        \rowcolor{grey}
        Bioscanner & 7 & 7 & 8 & 0 & (a) & 1\\
        Bioscanner & 8 & 2 & 12 & +1 & (a) & 2\\
        \rowcolor{grey}
        Neuroscanner & 8 & 5 & 18 & na & None & 2\\
        Chem Lab I & 5 & 20 & 2 & 0 & 1 & 1\\
        \rowcolor{grey}
        Chem Lab II & 7 & 15 & 7 & +1 & 0.5 & 1\\
        Chem Synthesizer I & 6 & 25 & 2 & 0 & 4 & 1\\
        \rowcolor{grey}
        Chem Synthesizer II & 8 & 22 & 6 & +2 & 2 & 2\\
        Energy Scanner & 7 & 5 & 10 & 0 & (a) & 1\\
      \end{tabular}

      \medskip

      \parbox{\textwidth}{\begin{description}
        \item[na:] Not Applicable. 
        \item[(a):] See \ref{sec:scientific-skills} for explanation. 
        \end{description}}
    \end{minipage}}
  \end{table}

\subsection[Tech Kits]{Tech kits enable a character to attempt to
  repair various types of damage to specific equipment.}
\label{sec:tech-kits}

The kits include replacement parts, troubleshooting equipment, tools,
and other items of repair. Unless otherwise stated, a kit may be used
to repair any type of damage to the appropriate item (even if the item
is partially destroyed).

\begin{description}
\item[\hypertarget{tag:basicrepait}{Basic Repair Kit}.] Allows repair of superficial damage to any
  item smaller than a large vehicle, and of light damage to any item
  weighing less than \textbf{15 kg}.
\item[Suit Kit.] Allows repair of environment suits and respirators.
  Also allows repair of punctures to body armor.
\item[Armor Kit.] Allows repair of environment suits, body armor, and
  all other items listed on the Protective Attire Chart, except the
  force field. May repair projectile and beam armor on body armor,
  vehicles, and armored structures. The Civ Level 7 kit does not allow
  repair of a partially destroyed item.
\item[Weapon Kit.] Allows repair of all weapons except those mounted
  on a spaceship. The Civ Level 6 kit does not allow repair of a
  partially destroyed item.
\item[\hypertarget{tag:electrokit}{Electro Kit}.] Allows repair of all types of small electronic
  equipment (scanners, labs, etc.). Partially destroyed equipment
  weighing more than \textbf{15 kg} may be repaired only with the Civ
  Level 8 version of the kit.
\item[\hyperlink{tag:robotkit}{Robot Kit}.] Allows repair of all types of damage to all robots. 
\item[\hypertarget{tag:vehiclekit}{Vehicle Kit}.] Allows repair to all types of vehicles. Does not
  repair any damage done to the armor of these vehicles. The Civ Level
  6 version of this kit will not repair a partially destroyed vehicle.
\item[\hypertarget{tag:spaceshipkit}{Spaceship Kit}.] Allows repair to all facets of spaceships, one
  compartment at a time. No spaceship kit will repair anything that is
  partially destroyed.
\end{description}

\begin{table}[htbp]
  \centering
  \fbox{%
    \begin{minipage}{3in}
      \centering
      \caption{Repair Equipment Chart}
      \label{tab:repair-equipment}

      \medskip

      \begin{tabular}{lcrrr}
        TECH KIT &
        \rotate{CIV LEVEL} & 
        \rotate{\parbox{1.5in}{WEIGHT\\(Kilograms)}} & 
        \rotate{\parbox{1.5in}{PRICE\\(TRANS)}} & 
        \rotate{SKILL BONUS}\\
        \rowcolor{grey}
        Basic Repair Kit & 5 & 5 & 1 & 0 \\
        Suit Kit & 7 & 2 & 2 & 0 \\
        \rowcolor{grey}
        Armor Kit I & 7 & 15 & 20 & +1 \\
        Armor Kit II & 8 & 15 & 30 & +2 \\
        \rowcolor{grey}
        Weapon Kit I & 6 & 7 & 2 & +1 \\
        Weapon Kit II & 8 & 7 & 4 & +2 \\
        \rowcolor{grey}
        Electro Kit I & 7 & 5 & 11 & +1 \\
        Electro Kit II & 8 & 5 & 15 & +2 \\
        \rowcolor{grey}
        Robot Kit & 8 & 15 & 30 & +1 \\
        Vehicle Kit I & 6 & 30 & 25 & +1 \\
        \rowcolor{grey}
        Vehicle Kit II & 8 & 30 & 35 & +2 \\
        Spacecraft Kit I & 6 & 100 & 60 & 0 \\
        \rowcolor{grey}
        Spacecraft Kit II & 7 & 100 & 75 & +1 \\
        Spacecraft Kit III & 8 & 100 & 100 & +2 \\
      \end{tabular}

      \medskip

      \parbox{\textwidth}{See \ref{sec:repair} for explanation.}
    \end{minipage}}
\end{table}

\subsection[Communications Gear]{Communications gear allows
  characters to converse over distances or translate speech.}
\label{sec:comm-gear}

Any communications device relying on the transmissions of radio waves
will be subject to the normal vagaries of that medium. The GM should
feel free to allow occurrences such as solar flares, magnetic
deviations, and signal jamming to interfere with the characters'
transmission. Amongst other communication devices, there are:

\begin{description}
\item[Headset Radio.] Installed inside a respirator helmet, this has a
  range of \textbf{2 km}.
\item[Planetary Shortwave.] Allows communication anywhere on any world
  and also into orbit, if the two radios are on the same side of the
  world.
\item[Interplanetary Radio Dish.] Allows communication anywhere within
  a star system.
\item[\hypertarget{tag:commlink}{Interstellar Commlink}.] Psionic
  device available to the public 
  allowing instantaneous communication between star systems that
  contain a psionic institute (see \ref{sec:psiinstitute}.  Puts
  character in contact with psion who will transmit short message
  anywhere in settled space.  In addition to purchase cost, there is a
  charge for each use of \textbf{50 Mils} plus \textbf{5 Mils} per
  light year message is sent.  \emph{Exception:} No charge for thinker or
  psionic navigator.
\item[\hypertarget{tag:translator}{Translator}.] Aids in communication
  with beings who speak a 
  foreign language. The Civ Level 7 version simply translates what the
  other being is saying; the Civ Level 8 version may allow the
  character to speak in that language also. In this version, it will
  yield a phonetic version of the phrases to be uttered.
\end{description}

\begin{table}[htbp]
  \centering
  \fbox{%
    \begin{minipage}{4in}
      \centering
      \caption{Communications Equipment Chart}
      \label{tab:comm-equip}

      \medskip

      \begin{tabular}{p{1.5in}ccccc}
        COMMUNICATIONS EQUIPMENT &
        \rotate{CIV LEVEL} &
        \rotate{\parbox{1.5in}{WEIGHT\\(Kilograms)}} &
        \rotate{\parbox{1.5in}{PRICE\\(TRANS)}} &
        \rotate{SKILL BONUS} &
        \rotate{\parbox{1.5in}{BASE REPAIR\\TIME (Hours)}} \\
        \rowcolor{grey}
        Headset Radio & 5 & 1 & 0.2 & +1 & 1 \\
        Planetary Short-wave & 5 & 20 & 0.4 & 0 & 1 \\
        \rowcolor{grey}
        Interplanetary Radio & 7 & 5 & 2 & 0 & 1 \\
        Planetary Headset & 8 & 1 & 2 & +1 & 2 \\
        \rowcolor{grey}
        Interstellar Commlink & 8 & 1 & 2 & na & 2 \\
        Translator & 7 & 2 & 16 & +1  & 1 \\
        \rowcolor{grey}
        Translator & 8 & 1 & 48 & +2  & 2 
      \end{tabular}

      \medskip

      \parbox{\textwidth}{\begin{description}
        \item[na:] Not Applicable. 
        \end{description}}
    \end{minipage}}
\end{table}
  
\subsection[Miscellaneous Equipment]{Miscellaneous equipment
  includes recording equipment, cages, vision equipment, and any
  other devices the characters may use.}
\label{sec:misc-equip}

\begin{description}
\item[\hypertarget{tag:superoid}{Superoid Camera}.] Produces a holographic still image; all
  developing is done within the camera. Price of the film (20 images)
  is \textbf{20 Mils}.
\item[\hypertarget{tag:Holographer}{Holographer}.] Produces a holographic still or moving picture;
  developing is done in the camera. Price of the film (one hour of
  images) is \textbf{50 Mils}.
\item[Force Cage.] A cage used to hold creatures utilizing a force
  field instead of bars. One force cage may hold a large creature.
  Force cages may be joined together to hold larger entities.
\item[Night Glasses.] Allow vision in the dark through infrared
  emissions. The item resembles binoculars.
\item[Rocket Flare.] A flare which, when shot from any pistol, rises
  to a height of \textbf{1 km} and explodes, sustaining the light for
  \textbf{\textonehalf} hour, floating on a small set of rockets.
\item[Super Cage.] A non-metallic, entirely sealed container. Buffered
  walls protect fragile life-forms.
\item[\hypertarget{tag:rock-blaster}{Rock Blaster}.] Shock absorbing jackhammer/compressor. May hold
  ore it extracts. Increases mining and asteroid mining Skill Levels
  by \textbf{1}. Price: \textbf{15 Trans}. Weight: \textbf{25 kg}. Civ
  Level: \textbf{6}.
\item[Anaerobic Torch.] Hand-held, electrical/chemical light source.
  Provides daylight in \textbf{20} meter radius for \textbf{100}
  hours.  Works in any atmosphere or a vacuum.
\item[Computer Time.] Available to characters in any urban environment
  of Civ Level 5 or above. \emph{Library data} time will be found at a
  Data Center, and most unclassified information concerning the world,
  star system, and federation will be easily found. It is very rare
  that information of a classified nature will be obtainable (even
  through use of programming skill) at a terminal of this type. One
  hour of time costs \textbf{10} times the Civ Level of the world in
  Mils. The sophistication of the information obtained increases with
  the Civ Level of the computer used.
  
  \emph{Programming time} is obtained differently. This includes all
  official commercial use of computers by the characters. All small
  business and record keeping procedures the characters may wish to
  carry out can be accomplished on a small portable business computer.
  Often, however, the GM will decide the task which the characters
  need to accomplish will require a larger machine. One hour of time
  is available through large corporations at a cost of \textbf{20}
  times the Civ Level of the world in Mils.
\item[\hypertarget{tag:audio-sealed-case}{Audio Sealed Case}.] An attach\'e case with the voice print of
  the owner encoded within the locking mechanism. Also has projectile
  Armor Rating of \textbf{2}. Price: \textbf{15 Trans}. Weight:
  \textbf{10 kg}. Civ Level: \textbf{7}.
\item[\hypertarget{tag:false-identity}{False Identity}.] A full set of false ID papers, photos,
  fingerprints, retinal patterns, etc. Price: \textbf{10 Trans $\times$
  Law Level of world}. Weight:
  Negligible. Civ Level: \textbf{6}.
\item[\hypertarget{tag:buscomp}{Business Computer}.] A small portable computer capable of
  performing many menial calculations and record keeping. Comes in
  many different sizes and I/O configurations. Price: \textbf{5
    Trans}. Weight: Varies; approximately 10 kg. Civ Level:
  \textbf{6}.
\item[Plastic Explosives.] Produces explosions of varying intensities;
  the GM must adjudicate their effects. Price: \textbf{1 Tran}.
  Weight: \textbf{1 kg}.  Civ Level: \textbf{5}.
\end{description}

This list by no means exhausts all the equipment the characters will
wish to purchase. The GM must create these items within the framework
of the guidelines shown in this Section; he should feel free to be as
creative as he desires. This particular area will personalize a GM's
campaign more than any other.

\begin{table}[htbp]
  \centering
  \fbox{%
    \begin{minipage}{5in}
      \centering
      \caption{Miscellaneous Equipment Chart}
      \label{tab:misc-equip}

      \medskip

      \begin{tabular}{lcccccc}
        MISCELLANEOUS &
        \rotate{CIV LEVEL} &
        \rotate{\parbox{1.5in}{WEIGHT\\(Kilograms)}} &
        \rotate{\parbox{1.5in}{PRICE\\(TRANS)}} &
        \rotate{SKILL BONUS} &
        \rotate{\parbox{1.5in}{TIME NEEDED\\FOR USE}} &
        \rotate{\parbox{1.5in}{BASE REPAIR\\TIME (Hours)}}\\
        EQUIPMENT&&&&&&\\
        RECORDING &&&&&&\\
        \rowcolor{grey}
        Still Camera & 4 & 5 & 0.1 & na & AR & 1 \\
        Superoid Camera & 5 & 7 & 0.4 & na & AR & 1 \\
        \rowcolor{grey}
        Holographer & 7 & 25 & 3 & na & AR & 1 \\
        VISION &&&&&&\\
        \rowcolor{grey}
        Night Glasses & 7 & 0.3 & 3 & na & na & 1 \\
        Rocket Flare & 5 & 0.5 & 0.1 & na & AR & na \\
        \rowcolor{grey}
        Anaerobic Torch & 6 & 0.5 & 0.1 & na & AR & na\\ 
        Flash Light & 4 & 0.5 & 0.1 & na & AR & 1 \\
        \rowcolor{grey}
        Binoculars & 5 & 1 & 0.1 & na & na & 1 \\
        AUDIO &&&&&&\\
        \rowcolor{grey}
        Audio Recorder & 5 & 2 & 0.1 & na & AR & 1 \\
        RESTRAINT &&&&&&\\
        \rowcolor{grey}
        Metal Cage 0.5m$^3$ & 2 & 2 & 0.1 & na & na & 1 \\
        Metal Cage 8.0m$^3$ & 2 & 50 & 0.5 & na & na & 1 \\
        \rowcolor{grey}
        Super Cage 0.5m$^3$ & 5 & 1 & 0.2 & na & na & 1\\ 
        Super Cage 0.5m$^3$ & 5 & 30 & 1 & na & na & 1 \\
        \rowcolor{grey}
        Force Cage & 8 & 3 & 20 & na & AR & 2 \\
        MINING &&&&&&\\
        \rowcolor{grey}
        Rock Blaster & 6 & 25 & 15 & +1 Mining &&\\
        \rowcolor{grey}
        &&&& +1 Asteroid Mining &&
      \end{tabular}

      \medskip

      \parbox{\textwidth}{\begin{description}
        \item[na:] Not Applicable. 
        \item[AR:] Activation takes One Action Round. 
        \end{description}}
    \end{minipage}}
\end{table}


%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "gm_guide"
%%% End: 
