%%
%% $Id: robots.tex,v 1.2 2004/10/08 19:42:17 goltz20707 Exp $
%%
\chapter{Robots}
\label{cha:robots}

%% ZZYZZ checked tables to here Wed -09-29 1341
%% ZZYZZ indexed to here Wed -09-29 1627

Robots may function as either mechanical aides and/or thinking
companions for characters. Robots may be used to increase characters'
Skill Levels when performing a given task, and sometimes they can
replace a character for minor jobs. They will often be able to provide
information to the characters
that would otherwise be unavailable.

Within society, robots are regarded with little overt concern, yet
doubts exist over the degree of independent thought which they should
be given. Therefore, some legal restrictions have been placed on the
actions and available systems of all robots, to avoid any problems
with these thought systems. Robots with artificial intelligence
occasionally go ``insane'' and the GM may implement an incident of
this kind when an accident is generated on the Encounter Table (see
\ref{tab:encounter}). The specific descriptions of robot thought systems
detail the likelihood of breakdowns such as these.

\section{Robot Attributes}
\label{sec:robot-attributes}

Robots are controlled by their owners, but others may illegally gain
control through use of programming and compu/\linebreak[0]robot tech
skills.  When purchased, a robot consists of a \emph{chassis}, a
variable number of \emph{systems} of the owner's choice, and two
\emph{controllers}: a headset and a hand-held version.

The GM and the players should record each robot's characteristics and
systems on an index card for use during play.  Robots may be purchased
with a one-week waiting period at the prices listed on a world of a
Civ Level equal to or higher than the Civ Level of the robot. Upon
placing the order for the robot, a character must declare which
systems he wants.

If attempting to purchase a Civ Level 8 robot on a Civ Level 7 world,
the price is doubled; on a Civ Level 6 world, the price is multiplied
by \textbf{10}. Below Civ Level 6, Level 8 robots would be very scarce and only
available through the black market. These price relationships hold
true for purchasing a Civ Level 7 robot on lower Civ Level worlds. 

The
names given on the Robot Chassis Chart are model names only; the
characters should invent individual names for their robots.

\subsection[Robot Chassis]{A robot chassis contains its central
  processor and any appendages necessary for full function.}
\label{sec:robot-chassis}

A robot chassis with no systems may move and pick up and carry
objects. It can see and hear as well as a human, but cannot synthesize
the information it receives, and will reveal this information in an
absolutely non-analytical manner. It will remember nothing, and will
only do something when commanded. It will attempt to do what it cannot
possibly do without any hesitation. Each chassis comes complete with
whatever systems the owner wishes and which can fit into the unit. The
chassis are rated for both Hardware Points and Software Capacity, and
the systems are given point and size ratings. A chassis cannot contain
more hardware or software than its capacity in either area.

The owner controls a robot through vocal commands or through a
controller (see \ref{sec:robot-controllers}). These controllers are to
every robot as keys are to an automobile; thus, the controller for
robot A cannot be used to command robot B. Through these controllers
the owner's voiceprint is placed in the robot's central processor, and
thereafter the robot will only answer its owner's voice. Civ Level 8
robots may have three other voices encoded in addition to the owner,
but will always give the owner's voice priority. This programming of
the additional voices is accomplished via the controller itself.

When in audible range, the robot may be commanded verbally without the
use of the controller. The robot will answer and speak to the owner in
a normal vocal range. The volume can be altered with the controller.

Robots will never object to being turned off by anyone, which is a
simple matter of flipping a switch on the controller. A chassis may
operate for 2 weeks before exhausting its energy. The batteries then
need replacement or recharging. Any vehicle with a range of 2,000 km
or greater may be used to recharge a robot's battery, and any city
will have recharging facilities. A normal fee for this service is 100
Mils. Extra batteries may be purchased for 500 Mils per battery.

A chassis has a standard vocabulary, and understands the meaning of
all distance and time measurements.

\subsection[Robot Controllers]{A controller is a device used to
  operate a robot.}
\label{sec:robot-controllers}

This device allows communication with a robot's central processor.
Controllers come in three configurations: hand-held (pocket calculator
size), headset (fits inside a respirator helmet), and wrist mount
(twice the size of a normal watch). The hand-held, headset, and wrist
mount versions all have a TV screen, transceiver, and --- except the
headset --- a keypad. The TV screen allows the user to see exactly
what the robot sees while the transceiver allows verbal two-way
communication. The keypad is used to transmit coded commands to the
robot and can at any time be used to operate the robot without verbal
communication. The keys are not marked, and only the owner knows which
keys issue which commands. This information can be shared with anyone
the owner wishes, and this is  the only way anyone whose voiceprint
is not in the robot's memory may operate it. Anyone with a programming
skill may-attempt to gain access to a robot through the controller's
keypad. If successful, the robot may be given new voiceprints and the
keypad new coding.

A controller's range is 10 km, and subject to normal radio
interference. The handheld and headset versions (both Civ Level 7)
come with the robot, while the wrist mount (Civ Level 8 robots only)
cost 5 Trans extra.

\subsection[Robot Combat]{A robot may be used to engage in combat.}
\label{sec:robot-combat}

A robot with a weapon may fire it at a specified target by receiving
an order to do so. It must have a weapon target system \emph{and} a weapon
system to engage in fire combat. The robot may fire as many times in
an Action Round as a human could. If a robot is firing a recoil
weapon, it does not suffer adverse modifications due to the recoil. A
robot will not fire a weapon (unless malfunctioning) withouta specific
command.

To engage in close combat, the robot must have a close attack system
and be ordered to engage.  \textbf{Exception}: A robot with a creative
thought system may engage in close combat by its own choice.

Movement of robots in an Action Round is detailed in
\ref{sec:action-round-movement}.  A robot may travel a number of
kilometers per hour equal to its \index{agility!and robot movement}Agility Rating, and it is affected by
terrain as detailed in \ref{sec:foot-travel}. \textbf{Note}: This
means a robot with an Agility Rating less than \textbf{5} cannot keep
up with humans travelling on foot.

\subsection[Robot Chassis Chart]{The Robot Chassis Chart lists the
  various types of robots and their characteristics.}
\label{sec:robot-chassis-chart}

See Table \vref{tab:robot-chassis}.

% Type             Avail ST   DX   AY  Port   Armor   Points  Wt Repair Cost
%                                      Cap.   Pj Bm   Hw Sw  (kg) Time
%
% CL 7
% Brummagen I        R   17    5    1   350   2   2   2   1   800   3    75
% Brummagen II       O    8    2    3    35   1   1   3   1   150   3   110
% Manner 31b         R   20    3    5   200   4   4   1   1   520   3    70
% Manner 36sd        O    9    8    8    90   2   1   2   2   300   3   125
% Manner 38sdf       O    5    1   10     8   2   3   1   2   120   3    90
% Frazette Blue      O    1    6   11     3   0   0   1   1    45   3    50
%
% CL 8
% Brummagen II       R   38   12    2   450   5   4   5   6   580   6   600
% Manner 44          R   33    4    5   900   5   5   5   4   1020  6   450
% Manner 50sd        R   20   12    6    60   4   4   2   4   140   6   200
% Manner 51sdf       O   18    7    9    15   4   2   5   7   140   6   770
% Frazette Green     R   22    9   10   250   4   5   6   4   370   6   600
% Frazette Amber     R   18   14   12   150   3   3   4   4   240   6   280
% Soidistant RAR     R   21   11   13   550   2   1   6   6   660   6   850
% Soidistant V-201   O    8   13   14     5   2   3   3   6    85   6   400
% Soidistant V-202   M   32   15   15   700   5   5   6   7   850   6  1200


\begin{table}[htbp]
  \centering
  \fbox{%
    \begin{minipage}{6in}
      \centering
      \caption{Robot Chassis Chart}
      \label{tab:robot-chassis}

      \medskip

      \begin{tabular}{lcrrrrrrrrrrr}
        \textbf{ROBOT TYPE}\\[-1in]
        \textbf{Civ Level 7} & \rotate{AVAILABILITY} & \rotate{STRENGTH} &
        \rotate{DEXTERITY} & \rotate{AGILITY} & \rotate{PORT CAPACITY} &
        \multicolumn{1}{c}{\rotate{\parbox{1in}{PROJECTILE\\ARMOR}}} &
        \rotate{\parbox{1in}{BEAM\\ARMOR}} & \rotate{\parbox{1in}{HARDWARE\\POINTS}} & \rotate{\parbox{1in}{SOFTWARE\\POINTS}} &
        \rotate{WEIGHT} & \rotate{\parbox{2in}{BASE REPAIR\\TIME (HOURS)}} & \rotate{COST
          (TRANS)}\\ 
        \rowcolor{grey}
        \hypertarget{tag:bot-brummagen-i}{Brummagen I}      & R & 17 &  5 &  1 & 350 & 2 & 2 & 2 & 1 & 800 & 3 &  75\\
        \hypertarget{tag:bot-brummagen-ii}{Brummagen II}     & O &  8 &  2 &  3 &  35 & 1 & 1 & 3 & 1 & 150 & 3 & 110\\
        \rowcolor{grey}
        Manner 31b       & R & 20 &  3 &  5 & 200 & 4 & 4 & 1 & 1 & 520 & 3 &  70\\
        \hypertarget{tag:bot-manner-36sd}{Manner 36sd}      & O &  9 &  8 &  8 &  90 & 2 & 1 & 2 & 2 & 300 & 3 & 125\\
        \rowcolor{grey}
        \hypertarget{tag:bot-manner-38sdf}{Manner 38sdf}     & O &  5 &  1 & 10 &   8 & 2 & 3 & 1 & 2 & 120 & 3 &  90\\
        \hypertarget{tag:bot-frazette-blue}{Frazette Blue}    & O &  1 &  6 & 11 &   3 & 0 & 0 & 1 & 1 &  45 & 3 &  50\\
        \multicolumn{13}{l}{\textbf{Civ Level 8}}\\
        \rowcolor{grey}
        Brummagen II     & R & 38 & 12 &  2 & 450 & 5 & 4 & 5 & 6 & 580 & 6 & 600\\
        \hypertarget{tag:bot-manner-44}{Manner 44}        & R & 33 &  4 &  5 & 900 & 5 & 5 & 5 & 4 & 1020& 6 & 450\\
        \rowcolor{grey}
        \hypertarget{tag:bot-manner-50sd}{Manner 50sd}      & R & 20 & 12 &  6 &  60 & 4 & 4 & 2 & 4 & 140 & 6 & 200\\
        Manner 51sdf     & O & 18 &  7 &  9 &  15 & 4 & 2 & 5 & 7 & 140 & 6 & 770\\
        \rowcolor{grey}
        Frazette Green   & R & 22 &  9 & 10 & 250 & 4 & 5 & 6 & 4 & 370 & 6 & 600\\
        \hypertarget{tag:bot-frazette-amber}{Frazette Amber}   & R & 18 & 14 & 12 & 150 & 3 & 3 & 4 & 4 & 240 & 6 & 280\\
        \rowcolor{grey}
        Soidistant RAR   & R & 21 & 11 & 13 & 550 & 2 & 1 & 6 & 6 & 660 & 6 & 850\\
        \hypertarget{tag:bot-soidistant-v-201}{Soidistant V-201} & O &  8 & 13 & 14 &   5 & 2 & 3 & 3 & 6 &  85 & 6 & 400\\
        \rowcolor{grey}
        Soidistant V-202 & M & 32 & 15 & 15 & 700 & 5 & 5 & 6 & 7 & 850 & 6 &1200\\
      \end{tabular}

      \medskip
      
      \parbox{\textwidth}{\textbf{Port Capacity} is in kilograms and
        represents the weight that may be carried.
        
        \textbf{Cost} (in Trans) includes cost of the listed hardware
        and software systems.
        
        \textbf{Weight} is the estimated weight in kilograms of the
        robot before additional hardware.
        
        \textbf{Availability} is the estimated restrictions of
        purchase and the requirement of permits based on the robot's
        potential danger to the public in strength and armor.}
    \end{minipage}}
\end{table}

\section{Robot Systems}

Robot systems are self-contained units, which augment the physical or
logical capacities of the robot. Each system details what abilities if
confers on the robot, how much it costs (if purchased independent of a
robot), and how many Hardware Points or how much Software Space it
occupies. Hardware adds to the physical capabilities of a robot;
Software contributes logic systems, which allow various thought
processes. Certain systems are available to Civ Level 8 robots only.

When deciding upon systems for a robot, the total Hardware Points
and/or Software Sizes may not exceed the capacity of the robot.  Extra
systems may be purchased, but may only be interchanged by a skilled
individual who has access to and can operate the robot's controller.
An individual with programming skill can interchange one Software
Point in half an hour; an individual with compu/robot tech may
interchange one Hardware Point in half an hour.

All systems that include a weapon are bound by the same legal
restrictions governing ownership of firearms by humans (see \ref{sec:weapons}).

The following list of systems includes the abilities the system
confers, their cost, and their Hardware Point total and Software Size.
Availability, Civ Level and new Systems. For clarification a Skill
Bonus is noted, for systems that have capabilities the Skill bonus is
not available (i.e. ``\ldots repair capabilities of Civ 8 Vehicle
Kit'' does not yield a +2 Vehicle Tech bonus. \emph{(Additional: Ian
  Taylor)}

\newcommand{\botsys}[2]{%
\bigskip%
%
\noindent\textbf{\hypertarget{tag:bot-#1}{#2}}}

\label{sec:robot-systems-change-1}
\botsys{aerodynamic}{Aerodynamic}\\
Allows flight via a modified jetpack as per 29.3. Flies in all
respects as a character with Skill Level 4.\\
\textbf{Availability:} Open \textbf{Civ Level:} 6 \textbf{Hardware:} 2 \textbf{Software:} 1
\textbf{Price:} 25

\botsys{agriculture}{Agriculture} \\
Functions just like a character with agriculture Skill Level 5. \\
\textbf{Availability:} Open \textbf{Civ Level:} 6 \textbf{Hardware:} 1
\textbf{Software:} 1 \textbf{Price:} 15 


\botsys{anti-bio}{Anti-Bio}\\
Includes all attributes of the \hyperlink{tag:bio}{bio} system (Aids a character with
biology skill; +2 to Biology Skill Level). Includes a catalog of all
known creatures, a Civ Level 8 Bioscanner, and a Civ Level 8
Neuroscanner), plus stun pistol, needle pistol, and a full supply of
drugs and poison for use against creatures. Weapon target system for
those weapons only also included.\\
\textbf{Availability:} Restricted
\textbf{Civ Level:} 8 \textbf{Hardware:} 2 \textbf{Software:} 1
\textbf{Price:} 40

\botsys{aquatic}{Aquatic}\\
Allows robot to swim above and beneath water. Robot's speed in liquid
in an Action Round is twice its Agility in hexes; long distance
movement is twice Agility in kilometers per
hour. Maximum depth is 50 meters.\\
\textbf{Availability:} Open \textbf{Civ Level:} 6 \textbf{Hardware:} 2
\textbf{Software:} 0 \textbf{Price:} 35

\botsys{arc-gun}{Arc Gun}\\
Includes arc gun and weapon target system
that may be used only for that weapon.\\
\textbf{Availability:} Military \textbf{Civ Level:} 8 \textbf{Hardware:}
1 \textbf{Software:} 1 \textbf{Price:} 30

\botsys{bio}{Bio}\\
Aids a character with biology skill; +2 to Biology Skill Level.
Includes a catalog of all known creatures, a Civ Level 8 Bioscanner,
and  a Civ Level 8 Neuroscanner. \\
\textbf{Availability:} Open \textbf{Civ Level:} 8 \textbf{HARDWARE
  POINTS:} 1 \textbf{Software:} 1 \textbf{Price:} 40  

\botsys{chemical}{Chemical}\\
Aids a character with chemistry skill; +2 to his Chemical Skill
Level. Includes a Civ Level 7 Chemlab and a Civ Level 8  
Chemsynthesizer. \\
\textbf{Availability:} Open \textbf{Civ Level:} 8 \textbf{Hardware:} 1 \textbf{Software:} 1 \textbf{Price:} 25 

\botsys{close-attack}{Close Attack}\\
Allows robot to engage in close combat during an Action Round. The
robot's Close Combat Strength is determined by adding together its
Strength, Dexterity, and Agility Ratings. If any of these is greater
than its Strength, that rating is considered \emph{equal} to its Strength
when calculating this sum. A robot that does not have this system
may never \emph{attack} in close combat and, if defending, uses its Strength
Rating only. Unless a robot possesses the creative thought system, it
will never attack unless commanded to do so.\\
\textbf{Availability:} Restricted \textbf{Civ Level:} 8
\textbf{Hardware:} 0 \textbf{Software:} 1
\textbf{Price:} 50  

\botsys{compu-robot-tech}{Compu/Robot Tech}\\
Federation prohibits robots with this system. \\
\textbf{Availability:} Forbidden \textbf{Civ Level:} Forbidden \textbf{Hardware:} Forbidden \textbf{Software:} Forbidden \textbf{Price:} 
Forbidden 

\botsys{construction}{Construction}\\ 
Functions in all respects as a character with a construction Skill Level of 4. 
\\
\textbf{Availability:} Open \textbf{Civ Level:} 6 \textbf{Hardware:} 1 \textbf{Software:} 1 \textbf{Price:} 15 

\botsys{creative-thought}{Creative Thought}\\
 Includes all features of self-activation, information, language, and
learning systems. The robot can interact with humans in all manner of
conversation and intellectual activity. An always operating logic
sub-system allows the robot to make conclusions and connections
concerning any data if has in its memory. The robot may make
observations the origins of which may elude the characters. The robot
always speaks in deference to humans, but does not like being shut
down. Movement is greatly improved by this system; if a robot's
movement in an Action Round is 0 or less, it may always move one hex
(like a character).
\textbf{Note:} Robots with this system may suffer mental malfunction
due to a common, rare, or unique accident. This system may only be
mounted on Civ Level 8 robots.\\
\textbf{Availability:} Restricted \textbf{Civ Level:} 8
 \textbf{Hardware:} 2 \textbf{Software:} 5
 \textbf{Price:} 100  

\botsys{driver}{Driver}\\
Functions in all respects as a character with a Vehicle Skill Level of
4; generic vehicle type (military, ground, air, or marine) must be  
chosen by the purchaser who must also allocate 10 Sub-Skill Points to
specific vehicle types.  
\\
\textbf{Availability:} Open \textbf{Civ Level:} 6 \textbf{Hardware:} 0 \textbf{Software:} 1 \textbf{Price:} 30  

\botsys{electro-tech}{Electro Tech}\\
Functions in all respects as a character with an Electro tech Skill
Level of 5 with repair capabilities of a Civ Level 8 Electrokit.  
\\
\textbf{Availability:} Open \textbf{Civ Level:} 8 \textbf{Hardware:} 1 \textbf{Software:} 1 \textbf{Price:} 25  

\botsys{eva}{EVA}\\
Allows robot to conduct operate in Zero Gravity. Equipped with Zero G
propulsion devices, grapple fittings and tethers for working in  
that environment. Functions in all respects as a character with an EVA
Skill Level of 6.  
\\
\textbf{Availability:} Open \textbf{Civ Level:} 6 \textbf{Hardware:} 1 \textbf{Software:} 1 \textbf{Price:} 15  \\
\emph{(Additional - Ian Taylor)}

\botsys{force-field}{Force Field}\\ 
Robot has a force field. Unless robot also has a self-activation
system, it must be commanded to turn on the field. Any one character  
standing immediately next to the robot is also included in the field;
both robot and character are subject to normal force field  
restrictions. 
\\
\textbf{Availability:} Restricted \textbf{Civ Level:} 8
\textbf{Hardware:} 2 \textbf{Software:} 0
\textbf{Price:} 65  

\botsys{geo}{Geo}\\
Includes all attributes of the miner system and has a Civ Level 8 Geoscanner. Aids character with miner skill, + 2 to his Geo Skill Level. 
Includes mining laser (equivalent to Civ Level 6 laser pistol), scooper/digger, rock blaster, and ore refiner. This system is required to 
refine ore. 
\\
\textbf{Availability:} Open \textbf{Civ Level:} 8 \textbf{Hardware:} 2 \textbf{Software:} 1 \textbf{Price:} 30 

\botsys{grenade-launcher}{Grenade Launcher}\\
Functions in all respects as a character with grenades Skill Level of 5. Requires weapon target system to fire. Includes two dozen 
grenades of any kind; refills must be provided by owner 
\\
\textbf{Availability:} Military \textbf{Civ Level:} 7 \textbf{Hardware:} 1 \textbf{Software:} 0 \textbf{Price:} 15 

\botsys{gunnery}{Gunnery}\\
Functions in all respects as a character with a Gunnery Skill Level of 6. 
\\
\textbf{Availability:} Military \textbf{Civ Level:} 7 \textbf{Hardware:} 0 \textbf{Software:} 2 \textbf{Price:} 50 

\botsys{heightened-vision}{Heightened Vision}\\
Robot can see twice as far as a human and may see in the dark (at normal range) as if it were light. 
\\
\textbf{Availability:} Open \textbf{Civ Level:} 7 \textbf{Hardware:} 0 \textbf{Software:} 1 \textbf{Price:} 15 

\botsys{holographer}{Holographer}\\
Equips robot with Holographer. If robot also has self-activation system, this does not take up any Software Space. 
\\
\textbf{Availability:} Open \textbf{Civ Level:} 7 \textbf{Hardware:} 1 \textbf{Software:} 1 \textbf{Price:} 15 


\botsys{information}{Information}\\
Enables robot to remember all data that it is exposed to.  The robot
will not analyse or draw any conclusions from this data; will discuss
anything it knows impartially and uncreatively. The robot's memory is
an open book; it will build a library of knowledge once activated.
\\
\textbf{Availability:} Open \textbf{Civ Level:} 6 \textbf{HARDWARE
  POINTS:} 0 \textbf{Software:} 1 \textbf{Price:} 30

\botsys{language}{Language}\\
Allows the robot to interpret and communicate fluently with languages besides Universal. Civ Level 7 robots have provision for one 
additional language while in Civ Level 8 robots may have four additional languages. The modified onboard Translator aids in 
communication between human beings who speak a foreign language with the robot acting as interpreter. If the robot has the language 
installed it will converse with an individual using the language it was addressed with. The system can also read and translates written 
languages. Language packages must be specified. 
\\
\textbf{Availability:} Open \textbf{Civ Level:} 7 \textbf{Hardware:} 0 \textbf{Software:} 1 \textbf{Price:} 30 



 
\botsys{learning}{Learning}\\ 
Enables robot to gain Experience Points for using any of its systems, which involve a Skill Level. A robot may never exceed the 
maximum Skill Level obtainable by a character. This system includes all features of the information system. 

\textbf{Note:} Robots with this system may suffer mental malfunction
due to a \emph{unique} accident. This system may be mounted only on
Civ Level 8 robots.
\\
\textbf{Availability:} Open \textbf{Civ Level:} 8 \textbf{Hardware:} 0
\textbf{Software:} 2 \textbf{Price:} 75

\botsys{medical}{Medical}\\
Aids a character with diagnosis or treatment skills. +3 to his Diagnosis and Treatment Skill Levels and includes attributes of a Civ Level 
8 Mediscanner with a full complement of drugs and medicines. 
\\
\textbf{Availability:} Restricted \textbf{Civ Level:} 8 \textbf{Hardware:} 1 \textbf{Software:} 2 \textbf{Price:} 50 

\botsys{miner}{Miner}\\
Aids character with Miner skill, +2 to his Miner Skill Level. Includes mining laser (equivalent to Civ Level 6 laser pistol), scooper/digger, 
rock blaster, and ore refiner. This system is required to refine ore. 
\\
\textbf{Availability:} Open \textbf{Civ Level:} 6 \textbf{Hardware:} 2 \textbf{Software:} 0 \textbf{Price:} 25 

\botsys{missile-guidance}{Missile Guidance}\\
Functions in all respects as a character with a Missile Guidance Skill Level of 6. 
\\
\textbf{Availability:} Restricted \textbf{Civ Level:} 7 \textbf{Hardware:} 0 \textbf{Software:} 2 \textbf{Price:} 50 

\botsys{pilot}{Pilot}\\
Functions in all respects as a character with a pilot Skill Level of 6. 
\\
\textbf{Availability:} Open \textbf{Civ Level:} 7 \textbf{Hardware:} 0 \textbf{Software:} 2 \textbf{Price:} 55 

\botsys{recorder}{Recorder}\\
Equips robot with a Visual / Audio Recording unit. Enables it to make an audio/visual recording of whatever it sees. 
\\
\textbf{Availability:} Open \textbf{Civ Level:} 6 \textbf{Hardware:} 1 \textbf{Software:} 0 \textbf{Price:} 10 

\botsys{self-activation}{Self-Activation}\\
Enables robot to take actions without commands. The robot will use any of its abilities and equipment as the owner wishes without 
specifically commanding it to. 

\textbf{Note:} Robots with this system may suffer mental malfunction
due to a \emph{rare} or \emph{unique} accident. This system may be
mounted only on Civ Level 8 robots.
\\
\textbf{Availability:} Open \textbf{Civ Level:} 8 \textbf{Hardware:} 0
\textbf{Software:} 2 \textbf{Price:} 75

\botsys{spacecraft-tech}{Spacecraft Tech}\\
Functions in all respects as a character with a Spaceship Tech Skill Level of 7 with repair capabilities of a Civ Level 8 Spacecraft kit. 
\\
\textbf{Availability:} Open \textbf{Civ Level:} 8 \textbf{Hardware:} 3 \textbf{Software:} 1 \textbf{Price:} 110 

\botsys{suit-tech.}{Suit Tech.}\\
Functions in all respects as a character with a Suit Tech Skill Level of 7 with repair capabilities of a Civ Level 7 Armor Kit. 
\\
\textbf{Availability:} Open \textbf{Civ Level:} 7 \textbf{Hardware:} 1 \textbf{Software:} 1 \textbf{Price:} 30 

\botsys{valet-secretary}{Valet/Secretary}\\
Enables robot to perform all functions of a gentleman's gentleman, combined with a live-in maid. 
\\
\textbf{Availability:} Open \textbf{Civ Level:} 7 \textbf{Hardware:} 0 \textbf{Software:} 1 \textbf{Price:} 30 

\botsys{vehicle-tech}{Vehicle Tech}\\
Functions in all respects as a character with a Vehicle Tech Skill Level of 5 with the repair capabilities of a Civ Level 8 Vehicle Kit 
\\
\textbf{Availability:} Open \textbf{Civ Level:} 8 \textbf{Hardware:} 2 \textbf{Software:} 1 \textbf{Price:} 35 


\botsys{weapon}{Weapon}\\
Functions in all respects as a character with a weapon Skill Level two
less than the maximum allowed for the weapon with which the robot is
equipped. Weapon choices include paint gun, laser pistol, rifle,
submachine gun, machine gun (owner must choose which and add the
weapon cost). Requires weapon target system to fire. A Civ Level 7
robot takes as much time to reload as a human; a Civ Level 8 robot can
reload in the same Action Round it fires. The system includes the
chosen weapon; a robot may be equipped with more than one weapon
system.
\\
\textbf{Availability:} Restricted \textbf{Civ Level:} 7
\textbf{Hardware:} 1 \textbf{Software:} 0 \textbf{Price:} 10

\botsys{weapon-target}{Weapon Target}\\
Required to fire weapon. May target any number of weapons, one at a time. 
\\
\textbf{Availability:} Restricted \textbf{Civ Level:} 7 \textbf{Hardware:} 0 \textbf{Software:} 1 \textbf{Price:} 20 

\botsys{weapon-tech}{Weapon Tech}\\
Functions in all respects as a character with a Weapon Tech Skill Level of 6 with the repair capabilities of a Civ Level 8 Weapon Kit. 
\\
\textbf{Availability:} Restricted \textbf{Civ Level:} 8 \textbf{Hardware:} 1 \textbf{Software:} 1 \textbf{Price:} 20 

\label{sec:robot-systems-change-2}
\begin{table}[htbp]
  \setlength{\tabcolsep}{0.5\tabcolsep}
  %\small
  \centering
  \fbox{%
    \begin{minipage}{\textwidth}
      \centering
      \caption{Robot Systems Table}
      \label{tab:robot-systems}

      \medskip

      \begin{small}
        \begin{tabular}{llccclllrc}
          \multicolumn{1}{c}{\textbf{System}} &
          \multicolumn{1}{c}{\textbf{AV}} & CIV & HP & 
          SP & \multicolumn{1}{c}{\textbf{Skill}} &
          \multicolumn{1}{c}{\textbf{Equipment}} & 
          \multicolumn{1}{c}{\textbf{Skill+}} & \textbf{Price} & \textbf{Notes}\\
                                % System    & Avail.     & C & H & S & Skill
                                % & Equipment        & Skill +      & $$ & Notes\\
          \rowcolor{grey}
          Aerodynamic & Open       & 6 & 1 & 1 & Jetpack 4
          & Modified Jetpack &              & 25 &\\
          Agriculture & Open       & 6 & 1 & 1 & Agriculture 5
          &                  &              & 15 &\\
          \rowcolor{grey}
          Anti-Bio    & Restricted & 8 & 2 & 1 & 
          & Civ 8 Bioscanner & +2 Biology & 40 &\\
          \rowcolor{grey}
          &            &  &  &  & 
          & Civ 8 Neuroscanner &            &    &\\
          \rowcolor{grey}
          &            &  &  &  & 
          & Catalog of Lifeforms &        &    &\\
          \rowcolor{grey}
          &            &  &  &  & Handgun 3
          & Stun Pistol            &        &    &\\
          \rowcolor{grey}
          &               &  &  &  & Handgun 3
          & Needle Pistol          &        &    &\\
          Aquatic  & Open          & 6 & 2 & 0
          &                        &        & &35 & 7\\
          \rowcolor{grey}
          Arc Gun  & Military   & 8 & 1 & 1 & Arc Gun 5 & Arc Gun & & 30 & 4\\
          Bio      & Open       & 8 & 1 & 1 & & Civ 8 Bioscanner & +2 Biology & 40\\
          &            &   &   &   & & Civ 8 Neuroscanner\\
          &            &   &   &   & & Catalog of Lifeforms\\
          \rowcolor{grey}
          Chemical & Open       & 8 & 1 & 1 & & Civ 7 Chemlab & +2
          Chemistry & 25 &\\
          \rowcolor{grey}
          &            &   &   &   & & Civ 8 Chemsynthesizer & & &\\
          Close Attack & Restricted & 8&0&1& & & & 50\\
          \rowcolor{grey}
          Construction & Open   & 6 & 1 & 1 & Construction 4 & & & 15 &\\
          Creative Thought & Restricted & 8&2&5 & & & & 100 & 1\\
          \rowcolor{grey}
          Driver   & Open       & 6 & 0 & 1 & Vehicle 4 & & & 30 &\\
          Electro Tech & Open   & 8 & 1 & 1 & Electro Tech 5 & Civ 8
          Electrokit & & 25\\
          \rowcolor{grey}
          EVA & Open & 6 & 1 & 1 & EVA 6 & Zero G Propulsion & & 15 &\\
          Force Field & Restricted & 8 & 2 & 0 & &  & 65\\
          \rowcolor{grey}
          Geo & Open & 8 & 2 & 1 & & Civ Level 8 Geoscanner & +2 Geology &
          30 &\\
          \rowcolor{grey}
          & & & & & & Mining Laser & +2 Miner & &\\
          \rowcolor{grey}
          &&&&&&Scooper Digger&&&\\
          \rowcolor{grey}
          &&&&&&Rock Blaster&&&\\
          \rowcolor{grey}
          &&&&&&Ore Refiner&&&\\
          Grenade Launcher & Military & 7 & 1 & 0 & Grenade 5 & Grenade
          Launcher & & 15 & 3\\
          &&&&&&24 Grenades\\
          \rowcolor{grey}
          Gunnery & Military & 7 & 0 & 2 & Gunnery 6 & & & 50 &\\
          Heightened Vision & Open & 7 & 0 & 1 & &&&15\\
          \rowcolor{grey}
          Holographer & Open & 7 & 1 & 1 & &Holographer & & 15 & 2\\
          Information & Open & 6 & 0 & 1 & &&&30\\
          \rowcolor{grey}
          Language & Open & 7 & 0 & 1 & &Translator & &30 & 5\\
          Learning & Open & 8 & 0 & 2 & &&&75 & 1\\
          \rowcolor{grey}
          Medical & Restricted & 8 & 1 & 2 & &Civ 8 Mediscanner & +3 Diagnosis & 40&\\
          \rowcolor{grey}
          &&&&&&&+3 Treatment&&\\
          Miner & Open & 6 & 2 & 0 & &Mining Laser & +2 Miner & 25\\
          &&&&&&Scooper Digger\\
          &&&&&&Rock Blaster\\
          &&&&&&Ore Refiner\\
          \rowcolor{grey}
          Missile Guidance & Restricted & 7 & 0 & 2 & Missile Guidance 6 & &&50&\\
          Pilot & Open & 7 & 0 & 2 & Pilot 6 & &&55\\
          \rowcolor{grey}
          Recorder & Open & 6 & 1 & 0 & &Recorder & &10&\\
          Self Activation & Open & 8 & 0 & 2 & &&&75 & 1\\
          \rowcolor{grey}
          Spacecraft Tech & Open & 8 & 3 & 1 & Spacecraft Tech 7 & Civ 8 Spacecraft Kit & &110&\\
          Suit Tech & Open & 7 & 1 & 1 & Suit Tech 7 & Civ 7 Armor Kit & &30\\
          \rowcolor{grey}
          Valet/Secretary & Open & 7 & 0 & 1 & &&&30&\\
          Vehicle Tech & Open & 8 & 2 & 1 & Vehicle 5 & Civ 8 Vehicle Kit & &35\\
          \rowcolor{grey}
          Weapon & Restricted & 7 & 1 & 0 & See note & &&10 & 3\\
          Weapon Target & Restricted & 7 & 0 & 1 & &&&20 & 6\\
          \rowcolor{grey}
          Weapon Tech & Restricted & 8 & 1 & 1 & Weapon Tech 6 & Civ 8 Weapon Kit & &20&
        \end{tabular}

        \medskip
        
        \parbox{\textwidth}{\textbf{AV}: Availability. \textbf{CIV}:
          Civ Level \textbf{HP}: Hardware Points \textbf{SP}: Software
          Points \textbf{Skill}: Inherent Skill \textbf{Skill +}:
          Skill Bonus

          See page \ref{tab:robot-systems-notes} for notes.}
      \end{small}
    \end{minipage}}
\end{table}


\begin{table}[htbp]
  \centering
  \fbox{%
    \begin{minipage}{0.95\textwidth}
      \centering
      \caption{Robot Systems Table (notes)}
      \label{tab:robot-systems-notes}
      
      \medskip
      
      \parbox{\textwidth}{\textbf{NOTES:}

        \begin{enumerate}
        \item Robots with this system may suffer mental malfunction due to a
          rare or unique accident. This system may be mounted only on Civ
          Level 8 robots.
        \item Software Point cost is 0 if the robot has the Self Activation
          System.
        \item Requires the Weapon Target System to be installed also. A Civ
          Level 7 robot takes as much time to reload as a human; a Civ Level
          8 robot can reload in the same Action Round it fires. The final
          weapon Skill level is 2 less than the maximum skill level
          permitted for the weapon.
        \item Has the Weapon Target System for use with the specified
          weapons only.
        \item Additional onboard languages; Robot is Civ 7: 1, Civ 8: 4.
        \item Required to fire any and all onboard weapons.
        \item Robot's speed in liquid in an Action Round is twice its
          Agility\index{agility!and Aquatic robot system} in hexes; long distance movement is twice Agility in
          kilometers per hour. Maximum depth is 50 meters.
        \end{enumerate}}
    \end{minipage}}
\end{table}

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "gm_guide"
%%% End: 
